### 2022-09-28 samplepaper.tex
Initial commit

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2018/03/10
%
\documentclass[runningheads]{llncs}

\usepackage[T1]{fontenc}
\def\doi#1{\href{https://doi.org/\detokenize{#1}}{\url{https://doi.org/\detokenize{#1}}}}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
%
\usepackage{listings}
\lstset{language=Pascal}
% Please use the

\begin{document}
%
\title{Contribution Title\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

Program listings or code snippets in the text like \lstinline!var i:integer;!
should be set in typewriter font. The ``listings'' macro package can be used
for reader-friendly syntax highlighting as in the following Pascal sample code:\footnote{For
typesetting pseudocode, the ``algorithmic'', ``algorithm2e'', ``algorithmicx'',
and ``program'' packages are also worth considering.}
\begin{lstlisting}
for i:=maxint to 0 do
begin
    { do nothing }
end;
Write('end of sample code');
\end{lstlisting}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.

\subsubsection{Acknowledgements} Please place your acknowledgments at
the end of the paper, preceded by an unnumbered run-in heading (i.e.
3rd-level heading).

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/123456789_0}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

### 2022-09-28 paper.tex
chore: simplify the latex structure

--- 2022-09-28 samplepaper.tex
@@
 %
-\title{Contribution Title\thanks{Supported by organization x.}}
+\title{Tracking Spatio-Temporal Phenomena through GNN and Aggregate Computing} %% Todo improve
 %
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
+\author{Gianluca Aguzzi\inst{1}\orcidID{todox} \and
+Roberto Casadei\inst{1}\orcidID{todox} \and
+Lukas Esterle \inst{2} \orcidID{todox} \and
+Mirko Viroli \inst{1}\orcidID{todox}}
 %
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
+\authorrunning{G. Aguzzi et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
+\institute{ALMA MATER STUDIORUM --- Università di Bologna, Cesena, Italy \and
+Aarhus University, Aarhus, Denmark}
+
 %
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+Todo
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Aggregate Computing  \and Graph Neural Network \and Cyber-Physical Swarms.}
 \end{abstract}
 %
-%
-%
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+\section{Introduction}
 
-Subsequent paragraphs, however, are indented.
+\section{Background}
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+%% Todo find a good name for the main contribution
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+\section{Simulation Setup}
 
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
+\section{Evaluation}
 
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
-\end{figure}
-
-Program listings or code snippets in the text like \lstinline!var i:integer;!
-should be set in typewriter font. The ``listings'' macro package can be used
-for reader-friendly syntax highlighting as in the following Pascal sample code:\footnote{For
-typesetting pseudocode, the ``algorithmic'', ``algorithm2e'', ``algorithmicx'',
-and ``program'' packages are also worth considering.}
-\begin{lstlisting}
-for i:=maxint to 0 do
-begin
-    { do nothing }
-end;
-Write('end of sample code');
-\end{lstlisting}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-
-\subsubsection{Acknowledgements} Please place your acknowledgments at
-the end of the paper, preceded by an unnumbered run-in heading (i.e.
-3rd-level heading).
-
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/123456789_0}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
+\section{Conclusion}
 \end{document}

### 2023-04-17 paper.tex
chore: add todos and add description
draft for title and abstract

@@
 %
-\title{Tracking Spatio-Temporal Phenomena through GNN and Aggregate Computing} %% Todo improve
+\title{Field-informed Reinforcement Learning for Learnining Large Scale Collective Tasks} %% Todo improve
+% or: Field-Informed Reinforcement Learning: A Scalable and Effective Method for Collective Intelligence
+% or: A Novel Approach to Reinforcement Learning for Adaptive Collective Systems
 %
 \author{Gianluca Aguzzi\inst{1}\orcidID{todox} \and
-Roberto Casadei\inst{1}\orcidID{todox} \and
-Lukas Esterle \inst{2} \orcidID{todox} \and
-Mirko Viroli \inst{1}\orcidID{todox}}
+%Roberto Casadei\inst{1}\orcidID{todox} \and
+Mirko Viroli \inst{1}\orcidID{todox} \and
+Lukas Esterle \inst{2} \orcidID{todox} 
+}
 %
@@
 \begin{abstract}
-Todo
+Coordinatinating a group of smart agents in multi-agent systems 
+ is a research problem that has been addressed for a long time, 
+ due to the challenges posed by distribution and of the definition of distributed intelligence. 
+%
+The problem is even more evident in collective adaptive systems, 
+ where the scale of the systems considered makes the definition of collective behaviors even more challenging.
 
+In this paper we consider a new reinforcement learning approach 
+  for the definition of intelligent agents called \emph{Field-Informed reinforcement learning}, 
+  where we use computational fields to manage the interaction between them in a stigmergic way 
+  and Graph Neural Networks to learn the intelligence necessary to solve collective tasks. 
+%
+This allows us to create distributed controllers informed by a collective knowledge 
+ distilled during learning but that use only local information at runtime.
+% 
+We demonstrate the effectiveness of this new approach in several case studies 
+ where coordination tasks are successfully solved.
 \keywords{Aggregate Computing  \and Graph Neural Network \and Cyber-Physical Swarms.}
@@
 
-\section{Simulation Setup}
-
 \section{Evaluation}
+\subsection{Simulation setup}
 

### 2023-04-17 paper.tex
chore: move to acsos template

@@
 %
-\documentclass[runningheads]{llncs}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
 
@@
 %
-\author{Gianluca Aguzzi\inst{1}\orcidID{todox} \and
-%Roberto Casadei\inst{1}\orcidID{todox} \and
-Mirko Viroli \inst{1}\orcidID{todox} \and
-Lukas Esterle \inst{2} \orcidID{todox} 
+
+\author{
+\IEEEauthorblockN{Gianluca Aguzzi}
+\IEEEauthorblockA{
+%\textit{Department of Computer Science and Engineering} \\
+\textit{%Alma Mater Studiorum--
+University of Bologna}\\
+Cesena, Italy\\
+gianluca.aguzzi@unibo.it
 }
-%
-\authorrunning{G. Aguzzi et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{ALMA MATER STUDIORUM --- Università di Bologna, Cesena, Italy \and
-Aarhus University, Aarhus, Denmark}
+
+\and
+%\linebreakand %\and
+\IEEEauthorblockN{Mirko Viroli}
+\IEEEauthorblockA{
+%\textit{Department of Computer Science and Engineering} \\
+\textit{%Alma Mater Studiorum--
+University of Bologna}\\
+Cesena, Italy\\
+mirko.viroli@unibo.it % 0000−0003−2702−5702
+}
+\and
+\IEEEauthorblockN{Lukas Esterle}
+\IEEEauthorblockA{
+%\textit{Department of Computer Science and Engineering} \\
+\textit{%Alma Mater Studiorum--
+Aarhus University}\\
+Aarhus, Denmark\\
+lukas.esterle@ece.au.dk}
+}
 
@@
  where coordination tasks are successfully solved.
-\keywords{Aggregate Computing  \and Graph Neural Network \and Cyber-Physical Swarms.}
 \end{abstract}
+
+\begin{IEEEkeywords}
+aggregate computing, Graph Neural Network, Cyber-Physical Swarms.
+\end{IEEEkeywords}
 %

### 2023-04-19 paper.tex
some intro

@@
 \usepackage[T1]{fontenc}
+\usepackage{xcolor}
 \def\doi#1{\href{https://doi.org/\detokenize{#1}}{\url{https://doi.org/\detokenize{#1}}}}
@@
 % Please use the
+
+\newcommand{\lukas}[1]{{\color{purple}\textbf{Lukas:} #1}}
 
@@
 We demonstrate the effectiveness of this new approach in several case studies 
- where coordination tasks are successfully solved.
+ where coordination tasks are successfully solved. \lukas{do we really have multiple case studies?} 
 \end{abstract}
@@
 \section{Introduction}
+Various phenomena in the real world are not evenly distributed across the environment such as watershed, wild fires, traffic, human crowd movement, or wild life movement from insects, fish, birds, and mammals.
+To acquire appropriate information about such phenomena, we also need to distribute the sensors accordingly to cover all aspects sufficiently.
+While manual deployment will mitigate the problem, not all phenomena are known exactly \emph{a priori} and would require maintenance and adjustment during runtime.
+Even worse, with phenomena able to change their location, size, and distribution---as it is the case with crowds, wild life, or wild fires---an adaptation of the collective sensors is required in order to keep the required information.
+
+This leads to several questions 
+\begin{enumerate}
+	\item How to distribute sensors to ensure good attainment of information?
+	\item How to maintain knowledge about the phenomena and unlearn irrelevant information?
+	\item How to move sensors when the phenomena moves to ensure good coverage?
+\end{enumerate} \lukas{refine and rework these questions}
+
+In this paper, we propose a novel approach combining aggregate computing~\cite{Beal2015Computer} with Graph Neural Networks to focus on the relevant areas.
+We further combine this with reinforcement learning techniques to respond to changing conditions in the environment.
+Specifically, we will show... \lukas{more info...}
+
+The remainder of this paper is structured as follows.
+First we introduce the relevant background and problem formulation.
+Afterwards we introduce our approach in Section~\ref{sec:approach}.
+Section~\ref{sec:eval} outlines the performed experiments and discusses the obtained results.
+Finally, we will present our conclusions in Section~\ref{sec:conclusion}.
 
 \section{Background}
+\label{sec:background}
 
@@
 
+\section{Approach}
+\label{sec:approach}
+
 \section{Evaluation}
+\label{sec:eval}
 \subsection{Simulation setup}
@@
 \section{Conclusion}
+\label{sec:conclusion}
+
+\bibliographystyle{IEEEtran}
+\bibliography{bibliography}
+
 \end{document}

### 2023-04-19 paper.tex
ref

@@
 
-In this paper, we propose a novel approach combining aggregate computing~\cite{Beal2015Computer} with Graph Neural Networks to focus on the relevant areas.
+In this paper, we propose a novel approach combining aggregate computing~\cite{Beal2015Computer} with Graph Neural Networks~\cite{Zhou2020AIOpen} to focus on the relevant areas.
 We further combine this with reinforcement learning techniques to respond to changing conditions in the environment.

### 2023-04-19 paper.tex
chore: merge with lukas intro

--- 2023-04-19 paper.tex
@@
 \lstset{language=Pascal}
-% Please use the
-
-\newcommand{\lukas}[1]{{\color{purple}\textbf{Lukas:} #1}}
-
 \begin{document}
@@
 }
-
 %
@@
 %
+
+%%% Comment command, to be removed before submission
+\newcommand{\meta}[3]{\textcolor{#1}{\textbf{#2}: #3}}
+\newcommand{\ga}[1]{\meta{red}{GA}{#1}}
+\newcommand{\lukas}[1]{\meta{purple}{Lukas}{#1}}
+\newcommand{\mv}[1]{\meta{green}{MV}{#1}}
+\ga{Page limit: 10 pages (include refs)}
+
 \begin{abstract}
-Coordinatinating a group of smart agents in multi-agent systems 
+Coordinatinating a group of intelligent agents in multi-agent systems 
  is a research problem that has been addressed for a long time, 
@@
   for the definition of intelligent agents called \emph{Field-Informed reinforcement learning}, 
-  where we use computational fields to manage the interaction between them in a stigmergic way 
-  and Graph Neural Networks to learn the intelligence necessary to solve collective tasks. 
+  where we use computational \emph{computational fields} to manage the interaction between agents 
+  in stigmergically and Graph Neural Networks to learn a local behaviour necessary to solve collective tasks. 
 %
@@
 
-In this paper, we propose a novel approach combining aggregate computing~\cite{Beal2015Computer} with Graph Neural Networks to focus on the relevant areas.
-We further combine this with reinforcement learning techniques to respond to changing conditions in the environment.
+In this paper, we propose a novel approach combining aggregate computing
+~\cite{Beal2015Computer} with Graph Neural Networks
+~\cite{Zhou2020AIOpen} to focus on the relevant areas. We further combine this with reinforcement learning techniques to respond to changing conditions in the environment.
 Specifically, we will show... \lukas{more info...}
@@
 \label{sec:background}
+\ga{Page budget: 1 page}
 
-%% Todo find a good name for the main contribution
+\section{Background and Motivation}
+\ga{Page budget: 1 page}
+\ga{Plan: discussed the problem of coordination in multi-agent systems 
+and the need for a scalable approach. In doing this, we will discuss some of the existing approaches}
+\subsection{Field Coordination Approaches}
+\subsection{Many-Agent Reinforcement Learning}
+\subsection{Graph Neural Network}
+
+\section{Field-informed Reinforcement Learning}
 
@@
 \subsection{Simulation setup}
+\subsection{Scenario}
 
-\section{Conclusion}
+\paragraph*{Variant A}
+\paragraph*{Variant B}
+\paragraph*{Variant C}
+
+\section{Conclusion and Future Work}
 \label{sec:conclusion}

### 2023-04-19 paper.tex

@@
 \usepackage{xcolor}
+\usepackage{cleveref}
 \def\doi#1{\href{https://doi.org/\detokenize{#1}}{\url{https://doi.org/\detokenize{#1}}}}
@@
 The remainder of this paper is structured as follows.
-First we introduce the relevant background and problem formulation.
+First we introduce the relevant background and problem formulation in \Cref{sec:background}.
 Afterwards we introduce our approach in Section~\ref{sec:approach}.
@@
 
-\section{Background}
+\section{Background and Motivation}
 \label{sec:background}
-\ga{Page budget: 1 page}
-
-\section{Background and Motivation}
-\ga{Page budget: 1 page}
-\ga{Plan: discussed the problem of coordination in multi-agent systems 
-and the need for a scalable approach. In doing this, we will discuss some of the existing approaches}
+\ga{Page budget: 1/2 page \\}
+\ga{Plan: discussion about the problem of coordination in multi-agent systems 
+and the need for a scalable approach. In doing this, we will discuss some of the existing approaches (declarative, RL and )}
 \subsection{Field Coordination Approaches}
+\ga{discussion about field used as a coordination mechanism in multi-agent systems. Introduction to AC}
+\subsection{Graph Neural Networks for Multi-Agent Systems}
 \subsection{Many-Agent Reinforcement Learning}
-\subsection{Graph Neural Network}
+\ga{plan: we could discuss briefly about standard appraoch (few agents), mean field approach (many agents).
+This will led to our approach, which is a combination of DQN (or any Deep RL approach) and GNN.}
 
 \section{Field-informed Reinforcement Learning}
-
-\section{Approach}
 \label{sec:approach}
-
+\ga{Page budget: 2/3 pages \\}
+\ga{Plan: we should discuss the approach in detail, including the system model, the learning dynamics.
+Particularly, I would like to highlight how the each node in the graph is a local controller,
+but it could be see as a global evolution, therefore we can use global information to inform the local controller.}
+\subsection{System model}
+\ga{Plan: 
+  discussion about typical AC evalaution model, specifying how GNN (with 1-hop information diffusion)
+  can be easily integrate with this distributed model
+}
+\subsection{Learning dynamics}
+\ga{
+  Plan: discuss about the choosing approach, that is Centralised Training and Distributed Execution.
+  Particularly, we use DQN in which the neural network is a GNN. So, the dataset consist in a set of graphs,
+  where each graph is a snapshot of the system. 
+  The target is the Q-value of the action taken in the current state.
+  Explain how this can be then use for distributed controller.
+}
 \section{Evaluation}
 \label{sec:eval}
-\subsection{Simulation setup}
+\ga{Page budget: 2 pages \\}
+\ga{Plan: we should discuss the robot aggregation scenario, specifiying the state space, the action space, the reward function.
+I dunno if it makes sense to discuss the variants (i.e., fixed area, two ares and moving area). We will see.
+}
 \subsection{Scenario}
@@
 
+\subsection{Discussion and Results}
 \section{Conclusion and Future Work}
+\ga{Page budget: 0.5}
 \label{sec:conclusion}

### 2023-04-24 paper.tex
chore: on MARL

@@
 \usepackage{cleveref}
+\usepackage{acronym}
 \def\doi#1{\href{https://doi.org/\detokenize{#1}}{\url{https://doi.org/\detokenize{#1}}}}
@@
 
+\acrodef{DecPOMDP}[DecPOMDP]{decentralized partially observable Markov decision processes}
+
 \begin{abstract}
@@
 First we introduce the relevant background and problem formulation in \Cref{sec:background}.
-Afterwards we introduce our approach in Section~\ref{sec:approach}.
-Section~\ref{sec:eval} outlines the performed experiments and discusses the obtained results.
-Finally, we will present our conclusions in Section~\ref{sec:conclusion}.
+Afterwards we introduce our approach in Section~\Cref{sec:approach}.
+Section~\Cref{sec:eval} outlines the performed experiments and discusses the obtained results.
+Finally, we will present our conclusions in Section~\Cref{sec:conclusion}.
 
@@
 
+Reinforcement learning (RL) has gained a lot of interest recently, 
+ thanks to its successful application in various scenarios, 
+ ranging from video games (such as Alpha Go and Atari) 
+ to chatbots (like ChatGPT). 
+% 
+In RL, an agent (i.e., a smart entity capable of making decisions) 
+ performs actions in an environment (i.e., everything outside the agent) according to a policy, 
+ with the goal of maximizing long-term rewards.
+
+One interesting application of RL is when there are multiple learning agents involved. 
+ Such scenarios are referred to as multi-agent reinforcement learning (MARL). 
+ While standard RL approaches use Markov decision processes to model the world, 
+ in MARL for distributed systems, the notion of 
+ \acp{DecPOMDP} is used.
+
+A \ac{DecPOMDP} represents a sequential game, 
+ in which each agent receives a local observation about the world they 
+ inhabit and receives a joint immediate reward in response to the actions they take. 
+% 
+\ga{Improve, if necessary, the following paragraph.}
+Formally, a DecPOMDP is defined as a tuple $<N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma>$, 
+ where $N$ is the total number of agents that live in the environment, 
+ $S$ is the state set of the environment, 
+ $\{A_i\}$ is the set of actions for agent $i$ and $A$ is the joint action space, 
+ $P$ is the probability transition function, 
+ $R$ is the immediate joint reward function, and 
+ $\{\Omega_i\}$ is the set observation function for the agent $i$,
+ $O$ is the observation space, and finally $\gamma$ is the discount factor.
+%
+On top of this model, each agent try to learn a policy $\pi_i$ in order 
+ to maximise the long term join reward signal.
+
+In base of some of this concept, 
+ we have several different MARL, like cooperative MARL (i.e., the reward function incetivize collective outcomes)
+ heterogenous MARL (i.e., the agents have different action space and policy) and
+ homogenous MARL (i.e., the agents have the same action space and policy).
+%  
+In particular, in this work we consider homogenous many agent RL, 
+ where, N >> 2 and action space of each agent and the policy is the same (i.e., the agents are interchangeable)
 \section{Field-informed Reinforcement Learning}

### 2023-04-24 paper.tex
fixed brackets

@@
 \ga{Improve, if necessary, the following paragraph.}
-Formally, a DecPOMDP is defined as a tuple $<N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma>$, 
+Formally, a DecPOMDP is defined as a tuple $\langle N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma \rangle$, 
  where $N$ is the total number of agents that live in the environment, 

### 2023-04-24 paper.tex
fixed greater-equal

@@
 In particular, in this work we consider homogenous many agent RL, 
- where, N >> 2 and action space of each agent and the policy is the same (i.e., the agents are interchangeable)
+ where, $N \geq 2$ and action space of each agent and the policy is the same (i.e., the agents are interchangeable)
 \section{Field-informed Reinforcement Learning}

### 2023-04-26 paper.tex
chore: more on many agent rl

--- 2023-04-24 paper.tex
@@
 \acrodef{DecPOMDP}[DecPOMDP]{decentralized partially observable Markov decision processes}
-
+\acrodef{RL}[RL]{reinforcement learning}
+\acrodef{MARL}[MARL]{multi-agent reinforcement learning}
 \begin{abstract}
@@
 
-Reinforcement learning (RL) has gained a lot of interest recently, 
+\Ac{RL} has gained a lot of interest recently, 
  thanks to its successful application in various scenarios, 
- ranging from video games (such as Alpha Go and Atari) 
- to chatbots (like ChatGPT). 
+ ranging from video games (such as Alpha Go~\cite{Silver2016Go} and Atari~\cite{Atari2016DQN}) 
+ to chatbots (like ChatGPT~\cite{ChatGPT2023}). 
 % 
-In RL, an agent (i.e., a smart entity capable of making decisions) 
+In \ac{RL}, an agent (i.e., a smart entity capable of making decisions) 
  performs actions in an environment (i.e., everything outside the agent) according to a policy, 
@@
 
-One interesting application of RL is when there are multiple learning agents involved. 
- Such scenarios are referred to as multi-agent reinforcement learning (MARL). 
- While standard RL approaches use Markov decision processes to model the world, 
+One interesting application of \ac{RL} is when there are multiple learning agents involved. 
+ Such scenarios are referred to as multi-agent reinforcement learning (MARL)~\cite{zhang2019marl}. 
+ While standard \ac{RL} approaches use Markov decision processes to model the world, 
  in MARL for distributed systems, the notion of 
- \acp{DecPOMDP} is used.
+ \acp{DecPOMDP}~\cite{Decpomdp2000} is used.
 
@@
 \ga{Improve, if necessary, the following paragraph.}
-Formally, a DecPOMDP is defined as a tuple $<N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma>$, 
+Formally, a \ac{DecPOMDP} is defined as a tuple $\langle N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma \rangle$, 
  where $N$ is the total number of agents that live in the environment, 
@@
 On top of this model, each agent try to learn a policy $\pi_i$ in order 
- to maximise the long term join reward signal.
+ to maximise the long term joint reward signal.
+In base of some of these concepts, 
+ there are several approaches/classifications in literature, 
+ like cooperative (i.e., the reward function incetivize collective outcomes)
+ heterogenous (i.e., the agents have different action space and policy) and
+ homogenous (i.e., the agents have the same action space and policy) settings.
+%  
+In particular, in this work we consider homogenous many agent RL~\cite{yang2021many}, 
+ where $N \gg 2$ and each agent is interchangeable and indistinguishable.
+%
+This research area is relevant in the context of in large-scale systems 
+ when the incentive of the individual is reduced compared to the collective behavior of the system, 
+ which is found in collective adaptive systems such as swarm robotics.
 
-In base of some of this concept, 
- we have several different MARL, like cooperative MARL (i.e., the reward function incetivize collective outcomes)
- heterogenous MARL (i.e., the agents have different action space and policy) and
- homogenous MARL (i.e., the agents have the same action space and policy).
+In fact, the first works in this direction were discussed precisely in the last field, 
+ exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating 
+ a policy equal for the entire system. 
+Modern approaches, however, 
+ have started to consider the use of deep learning approaches to try 
+ to synthesize robust controllers capable of generalizing to new tasks. 
+In this context, mean-field reinforcement learning~\cite{pmlr-v80-yang18d} is certainly noteworthy. 
+ Mean-field RL is a technique for learning approximate inference in large-scale systems, 
+ where the interactions between the agents can be decoupled into a 
+ single conditional probability distribution function for each agent, 
+ thus simplifying the learning process and reducing the computational cost.
 %
-In particular, in this work we consider homogenous many agent RL, 
- where, N >> 2 and action space of each agent and the policy is the same (i.e., the agents are interchangeable)
+Some known approaches using mean-field reinforcement learning include Q-mean, 
+ which is an extension of Q-learning to mean-field settings, 
+ and actor-critic mean-field, which combines actor-critic algorithms with mean-field approximations. 
+%
+These approaches have shown promising results in various domains, such as multi-agent coordination 
+ and decentralized control, and are actively being researched and developed for further applications
+
 \section{Field-informed Reinforcement Learning}

### 2023-05-03 paper.tex
feat: on GNN

@@
 \usepackage{xcolor}
-\usepackage{cleveref}
 \usepackage{acronym}
+\usepackage{amsmath}
 \def\doi#1{\href{https://doi.org/\detokenize{#1}}{\url{https://doi.org/\detokenize{#1}}}}
@@
 \usepackage{graphicx}
+
 % Used for displaying a sample figure. If possible, figure files should
@@
 \usepackage{listings}
+
+\usepackage{cleveref}
 \lstset{language=Pascal}
@@
 \ga{Page limit: 10 pages (include refs)}
-
 \acrodef{DecPOMDP}[DecPOMDP]{decentralized partially observable Markov decision processes}
@@
 %
+\ga{Paper levels:
+\begin{itemize}
+  \item Distributed collective intelligence
+  \item GNN and field coordination
+  \item Spatial tracking
+\end{itemize}
+}
 \section{Introduction}
@@
 \ga{discussion about field used as a coordination mechanism in multi-agent systems. Introduction to AC}
-\subsection{Graph Neural Networks for Multi-Agent Systems}
+\subsection{Graph Neural Networks}
+Over the years, several different neural networks have been proposed to solve specific tasks. 
+%
+Whereas Convolutional Neural Networks are designed to be used with image and spatial-like data (e.g., audio),
+ Graph Neural Networks (GNNs) are a novel model used to process graph-structured data. 
+%
+Let $ G = (V, E) $ be a graph, where $ V $ is the set of nodes and $ E $ is the set of edges. 
+%
+Given $X_v$ a feature vector associated to each node $v \in V$, 
+ the goal of a GNN is to learn the node embedding $X'_v$ for each node $v \in V$.
+ In modern GNNs, the node embedding $X'_v$ is computed by aggregating information from the node's neighbors $N(v)$,
+  and then combining it with the node's current embedding $X_v$.
+% 
+Formally, a GNN can be defined by two phases:
+\begin{equation}
+a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}(v)\right\}\right)   
+\end{equation}
+\begin{equation}
+h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
+\end{equation}
+where $h_{v}^{k}$ is the embedding of node $v$ at the $k$-th layer, 
+ $\mathcal{N}(v)$ is the set of neighbors of node $v$ computed from $E$.
+%
+The differential part come into play in the $\operatorname{COMBINE}$ function, 
+ which is usually a differentiable function such as a neural network layer.
+%
+$\operatorname{AGGREGATE} $ instea is a function that aggregates the information from the neighbors of a node $v$
+  and it could be a simple sum or a more complex function such as a neural network layer.
+%
+This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbors.
+
+GNN are used in several fields such as social network analysis, chemistry, and physics.
+In this paper, we use GNNs to learn a local behavior for each agent in a multi-agent system (more details in Section~\ref{sec:approach}).
 \subsection{Many-Agent Reinforcement Learning}

### 2023-05-03 paper.tex
chore: minors and citations

--- 2023-04-26 paper.tex
@@
 \newcommand{\mv}[1]{\meta{green}{MV}{#1}}
-\ga{Page limit: 10 pages (include refs)}
-
+%% ACRONYMS
 \acrodef{DecPOMDP}[DecPOMDP]{decentralized partially observable Markov decision processes}
@@
 \acrodef{MARL}[MARL]{multi-agent reinforcement learning}
+\acrodef{MLP}[MLP]{multi-layer perceptron}
+\acrodef{GNN}[GNN]{graph neural network}
+\acrodef{CNN}[CNN]{convolutional neural network}
+\ga{Page limit: 10 pages (include refs)}
+
 \begin{abstract}
@@
  is a research problem that has been addressed for a long time, 
- due to the challenges posed by distribution and of the definition of distributed intelligence. 
+ due to the challenges posed by distribution and the definition of distributed intelligence. 
 %
 The problem is even more evident in collective adaptive systems, 
- where the scale of the systems considered makes the definition of collective behaviors even more challenging.
+ where the scale of the systems considered makes the definition of collective behaviours even more challenging.
 
-In this paper we consider a new reinforcement learning approach 
+In this paper, we consider a new reinforcement learning approach 
   for the definition of intelligent agents called \emph{Field-Informed reinforcement learning}, 
@@
 The remainder of this paper is structured as follows.
-First we introduce the relevant background and problem formulation in \Cref{sec:background}.
-Afterwards we introduce our approach in Section~\Cref{sec:approach}.
+First, we introduce the relevant background and problem formulation in \Cref{sec:background}.
+Afterwards, we introduce our approach in Section~\Cref{sec:approach}.
 Section~\Cref{sec:eval} outlines the performed experiments and discusses the obtained results.
@@
 \ga{discussion about field used as a coordination mechanism in multi-agent systems. Introduction to AC}
-\subsection{Graph Neural Networks for Multi-Agent Systems}
+\subsection{Graph Neural Networks}
+Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
+  simple \ac{MLP} to \ac{CNN}~\cite{oshea2015introduction}.
+%
+Whereas \ac{CNN} is designed to be used with image and spatial-like data (e.g., audio),
+ \ac{GNN} is a novel model used to process graph-structured data. 
+%
+Let $ G = (V, E) $ be a graph, where $ V $ is the set of nodes and $ E $ is the set of edges. 
+%
+Given $X_v$ a feature vector associated to each node $v \in V$, 
+ the goal of a GNN is to learn the node embedding $X'_v$ for each node $v \in V$.
+ In modern GNNs, the node embedding $X'_v$ is computed by aggregating information from the node's neighbours $N(v)$,
+  and then combining it with the node's current embedding $X_v$.
+% 
+Formally, a GNN can be defined by two phases:
+\begin{equation}
+a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}(v)\right\}\right)   
+\end{equation}
+\begin{equation}
+h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
+\end{equation}
+where $h_{v}^{k}$ is the embedding of node $v$ at the $k$-th layer, 
+ $\mathcal{N}(v)$ is the set of neighbors of node $v$ computed from $E$.
+%
+The differential part come into play in the $\operatorname{COMBINE}$ function, 
+ which is usually a differentiable function such as a neural network layer.
+%
+$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
+  and it could be a simple sum or a more complex function such as a neural network layer.
+%
+This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
+
+\acp{GNN} are used in several fields such as social network analysis, chemistry, and physics.
+In this paper, we use GNNs to learn a local behaviour for each agent in a multi-agent system (more details in Section~\ref{sec:approach}).
 \subsection{Many-Agent Reinforcement Learning}
-\ga{plan: we could discuss briefly about standard appraoch (few agents), mean field approach (many agents).
-This will led to our approach, which is a combination of DQN (or any Deep RL approach) and GNN.}
+\ga{plan: we could discuss briefly about standard approach (few agents), and mean-field approach (many agents). 
+This will lead to our approach, which is a combination of DQN (or any Deep RL approach) and GNN.}
 
@@
  performs actions in an environment (i.e., everything outside the agent) according to a policy, 
- with the goal of maximizing long-term rewards.
+ to maximize long-term rewards.
 
@@
 %
-On top of this model, each agent try to learn a policy $\pi_i$ in order 
- to maximise the long term joint reward signal.
-In base of some of these concepts, 
- there are several approaches/classifications in literature, 
- like cooperative (i.e., the reward function incetivize collective outcomes)
- heterogenous (i.e., the agents have different action space and policy) and
+On top of this model, each agent tries to learn a policy $\pi_i$ in order 
+ to maximise the long-term joint reward signal.
+In the case of some of these concepts, 
+ there are several approaches/classifications in the literature, 
+ like cooperative (i.e., the reward function incentivizes collective outcomes)
+ heterogeneous (i.e., the agents have different action spaces and policies) and
  homogenous (i.e., the agents have the same action space and policy) settings.
 %  
-In particular, in this work we consider homogenous many agent RL~\cite{yang2021many}, 
+In particular, in this work, we consider homogenous many agent RL~\cite{yang2021many}, 
  where $N \gg 2$ and each agent is interchangeable and indistinguishable.
 %
-This research area is relevant in the context of in large-scale systems 
- when the incentive of the individual is reduced compared to the collective behavior of the system, 
+This research area is relevant in the context of large-scale systems 
+ when the incentive of the individual is reduced compared to the collective behaviour of the system, 
  which is found in collective adaptive systems such as swarm robotics.
@@
  exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating 
- a policy equal for the entire system. 
+ a policy equal to the entire system. 
 Modern approaches, however, 
@@
  Mean-field RL is a technique for learning approximate inference in large-scale systems, 
- where the interactions between the agents can be decoupled into a 
- single conditional probability distribution function for each agent, 
+ where the interactions between the agents can be decoupled into a single conditional probability distribution function for each agent, 
  thus simplifying the learning process and reducing the computational cost.
@@
 \ga{Page budget: 2/3 pages \\}
-\ga{Plan: we should discuss the approach in detail, including the system model, the learning dynamics.
-Particularly, I would like to highlight how the each node in the graph is a local controller,
-but it could be see as a global evolution, therefore we can use global information to inform the local controller.}
+\ga{Plan: we should discuss the approach in detail, including the system model, and the learning dynamics.
+Particularly, I would like to highlight how each node in the graph is a local controller,
+but it could be seen as a global evolution, therefore we can use global information to inform the local controller.}
 \subsection{System model}
 \ga{Plan: 
-  discussion about typical AC evalaution model, specifying how GNN (with 1-hop information diffusion)
-  can be easily integrate with this distributed model
+  discussion about typical AC evaluation model, specifying how GNN (with 1-hop information diffusion)
+  can be easily integrated with this distributed model
 }
@@
 \ga{
-  Plan: discuss about the choosing approach, that is Centralised Training and Distributed Execution.
-  Particularly, we use DQN in which the neural network is a GNN. So, the dataset consist in a set of graphs,
+  Plan: discuss the chosen approach, that is Centralised Training and Distributed Execution.
+  Particularly, we use DQN in which the neural network is a GNN. So, the dataset consists in a set of graphs,
   where each graph is a snapshot of the system. 
   The target is the Q-value of the action taken in the current state.
-  Explain how this can be then use for distributed controller.
+  Explain how this can be then used for distributed controllers.
 }
@@
 \ga{Page budget: 2 pages \\}
-\ga{Plan: we should discuss the robot aggregation scenario, specifiying the state space, the action space, the reward function.
-I dunno if it makes sense to discuss the variants (i.e., fixed area, two ares and moving area). We will see.
+\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
+I dunno if it makes sense to discuss the variants (i.e., fixed area, two areas and moving area). We will see.
 }

### 2023-05-03 paper.tex
chore: minors

@@
 \ga{Plan: 
-  discussion about typical AC evaluation model, specifying how GNN (with 1-hop information diffusion)
+  discussion about the typical AC evaluation model, specifying how GNN (with 1-hop information diffusion)
   can be easily integrated with this distributed model

### 2023-05-03 paper.tex
chore: minor on title
chore: on intro

@@
 %
-\title{Field-informed Reinforcement Learning for Learnining Large Scale Collective Tasks} %% Todo improve
+\title{Field-informed Reinforcement Learning for Learning Large Scale Collective Tasks} %% Todo improve
 % or: Field-Informed Reinforcement Learning: A Scalable and Effective Method for Collective Intelligence
@@
 \acrodef{MARL}[MARL]{multi-agent reinforcement learning}
+\acrodef{MAARL}[ManyRL]{many-agent reinforcement learning}
+
 \acrodef{MLP}[MLP]{multi-layer perceptron}
@@
 
-In this paper, we propose a novel approach combining aggregate computing
-~\cite{Beal2015Computer} with Graph Neural Networks
-~\cite{Zhou2020AIOpen} to focus on the relevant areas. We further combine this with reinforcement learning techniques to respond to changing conditions in the environment.
-Specifically, we will show... \lukas{more info...}
-
+In this paper, we propose a novel approach combining aggregate computing~\cite{Beal2015Computer} with Graph Neural Networks~\cite{Zhou2020AIOpen} to focus on the relevant areas.
+We further combine this with reinforcement learning techniques to respond to changing conditions in the environment.
+%Specifically, we will show... \lukas{more info...}
+Specifically, 
+ our approach utilizes aggregate computing to disseminate information by manipulating a computational field, 
+ which is a distributed data structure that maps devices to information. 
+ This field functions as a dynamic environment in which information is continuously updated and diffused throughout the system. 
+ By leveraging this layer, we can build collective distributed intelligence using \ac{GNN},
+  which uses the information in the field to determine the best action for a given task.
+\ga{first draft, need a refinement}
 The remainder of this paper is structured as follows.
@@
 \section{Background and Motivation}
+Field-based RL draws upon research areas that were not previously considered in traditional RL. 
+%
+This section provides an overview of three key concepts that underpin this approach, 
+ namely 
+ field-based coordination --i.e., coordinating multi-agent systems through computational fields--, 
+ \ac{GNN} --i.e., a novel type of neural network architecture designed to process graph-structured data--, 
+ and \ac{MAARL} ---i.e., a large number of agents learning a collective policy to achieve a global goal.
 \label{sec:background}

### 2023-05-03 paper.tex
chore: remove section written twice

@@
 First, we introduce the relevant background and problem formulation in \Cref{sec:background}.
-Afterwards, we introduce our approach in Section~\Cref{sec:approach}.
-Section~\Cref{sec:eval} outlines the performed experiments and discusses the obtained results.
-Finally, we will present our conclusions in Section~\Cref{sec:conclusion}.
+Afterwards, we introduce our approach in \Cref{sec:approach}.
+\Cref{sec:eval} outlines the performed experiments and discusses the obtained results.
+Finally, we will present our conclusions in \Cref{sec:conclusion}.
 

### 2023-05-04 paper.tex
chore: rewrite GNN section and add a problem definition (probably to general, we need to think about it)

@@
 \section{Background and Motivation}
-Field-based RL draws upon research areas that were not previously considered in traditional RL. 
+\label{sec:background}
+In this section, we aim to provide a concise overview of the system we intend to address, 
+ starting with a system definition. 
 %
-This section provides an overview of three key concepts that underpin this approach, 
- namely 
- field-based coordination --i.e., coordinating multi-agent systems through computational fields--, 
- \ac{GNN} --i.e., a novel type of neural network architecture designed to process graph-structured data--, 
- and \ac{MAARL} ---i.e., a large number of agents learning a collective policy to achieve a global goal.
-\label{sec:background}
+We then proceed to discuss relevant research areas that inform our approach. 
+%
+Finally, we present a formalization of the problem and a discussion of the motivations behind our proposed work.
+%Field-based RL draws upon research areas that were not previously considered in traditional RL. 
+%
+%This section provides an overview of three key concepts that underpin this approach, 
+% namely 
+% field-based coordination --i.e., coordinating multi-agent systems through computational fields--, 
+% \ac{GNN} --i.e., a novel type of neural network architecture designed to process graph-structured data--, 
+% and \ac{MAARL} ---i.e., a large number of agents learning a collective policy to achieve a global goal.
+
 \ga{Page budget: 1/2 page \\}
@@
 and the need for a scalable approach. In doing this, we will discuss some of the existing approaches (declarative, RL and )}
+\subsection{Problem definition}
+This article focuses on the definition of decentralized collective intelligence within the context of large-scale distributed systems. 
+%
+Specifically, we consider the case where agents can interact through 1-hop neighbourhood connections and have a local perception of the environment, 
+ which is typical in systems such as swarm robotics or large-scale IoT.
+
+We model the system at a given time instance as a graph, $G = (V, E)$, 
+ where $E$ defines the neighbourhood relations for each participating node, 
+ and $V$ identifies the nodes present in the graph at that time. 
+% 
+Each node $v$ is associated with an observation (or feature set), $o_v$, 
+ which can come from sensors installed on it and messages received from the neighbourhood in the previous step.
+
+Our goal is to find a homogeneously distributed controller $\pi$ that, 
+ starting from only local configurations, leads 
+ the system to achieve a graph $G_T$ that satisfies certain collective requirements through cooperation, such as spatial area coverage, phenomena tracking, and robot aggregation.
 \subsection{Field Coordination Approaches}
@@
 %
-Let $ G = (V, E) $ be a graph, where $ V $ is the set of nodes and $ E $ is the set of edges. 
+Let $ G $ be the above-defined graph.
 %
-Given $X_v$ a feature vector associated to each node $v \in V$, 
- the goal of a GNN is to learn the node embedding $X'_v$ for each node $v \in V$.
- In modern GNNs, the node embedding $X'_v$ is computed by aggregating information from the node's neighbours $N(v)$,
-  and then combining it with the node's current embedding $X_v$.
+Given $o_v$ a feature vector associated with each node $v \in V$, 
+ the goal of a GNN is to learn the node embedding $h_v$ for each node $v \in V$.
+ In modern GNNs, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $N(v)$,
+  and then combining it with the node's current embedding $h_v$.
 % 
@@
 \end{equation}
-where $h_{v}^{k}$ is the embedding of node $v$ at the $k$-th layer, 
+where $h_{v}^{k}$ is the embedding of node $v$ at the $k$-th layer ($h_v^0 = o_v$), 
  $\mathcal{N}(v)$ is the set of neighbors of node $v$ computed from $E$.

### 2023-05-05 paper.tex
feat: add system defintion and problem formalisation. MARL section need to be shrinked

@@
 \usepackage[T1]{fontenc}
+\usepackage{amsfonts}
 \usepackage{xcolor}
@@
 \acrodef{MAARL}[ManyRL]{many-agent reinforcement learning}
-
+\acrodef{MDP}[MDP]{Markov Decision Process}
 \acrodef{MLP}[MLP]{multi-layer perceptron}
@@
 \subsection{Problem definition}
-This article focuses on the definition of decentralized collective intelligence within the context of large-scale distributed systems. 
+This article focuses on the definition of \emph{decentralized collective intelligence} within the context of \emph{large-scale} distributed systems. 
 %
-Specifically, we consider the case where agents can interact through 1-hop neighbourhood connections and have a local perception of the environment, 
- which is typical in systems such as swarm robotics or large-scale IoT.
-
-We model the system at a given time instance as a graph, $G = (V, E)$, 
- where $E$ defines the neighbourhood relations for each participating node, 
- and $V$ identifies the nodes present in the graph at that time. 
+Specifically, we consider systems of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator}
+ each of which can only interact with the \emph{1-hop} neighbourhood that can 
+ be logical (i.e., a communication network) or physical (i.e., a spatial neighbourhood).
 %
-Each node $v$ is associated with an observation (or feature set), $o_v$, 
- which can come from sensors installed on it and messages received from the neighbourhood in the previous step.
-
-Our goal is to find a homogeneously distributed controller $\pi$ that, 
- starting from only local configurations, leads 
- the system to achieve a graph $G_T$ that satisfies certain collective requirements through cooperation, such as spatial area coverage, phenomena tracking, and robot aggregation.
+Each node executes a \emph{local} controller that can only access the information available to its \emph{context}
+ that is the information available in the neighbourhood and the information available from its sensors.
+%
+After each program evaluation, 
+ the node can send messages to its neighbourhood about its current evaluation and update its internal state.
+%
+This general model can be adapted to several different scenarios, 
+ such as a network of sensors, a network of robots (e.g., swarm robotics), or a network of mobile phones.
+%
+Our goal is to find a \emph{homogeneously} distributed controller $\pi$ (i.e., the same controller for the whole system) that, 
+ starting from \emph{only} local configurations, leads 
+ the system to achieve a certain \emph{collective} requirements through \emph{cooperation}, 
+ such as spatial area coverage, phenomena tracking, and robot aggregation.
+%
+The homogeneity of the controller is a key requirement to ensure the \emph{scalability} of the approach, 
+ as it allows us to avoid the need for a centralised controller that would be a bottleneck for the system,
+ and it is the typical choice in swarm-like systems~\cite{brambilla2013swarm,yang2021many,pmlr-v80-yang18d,DBLP:conf/aaai/ZhengYCZZWY18}.
+%
 \subsection{Field Coordination Approaches}
@@
 %
-Let $ G $ be the above-defined graph.
+Let $G = (V, E)$ a graph  
+where $E$ defines the neighbourhood relations for each participating node, 
+and $V$ identifies the nodes present in the graph at that time. 
+% 
+Each node $v$ is associated with an observation (or feature set), $o_v$, 
 %
 Given $o_v$ a feature vector associated with each node $v \in V$, 
- the goal of a GNN is to learn the node embedding $h_v$ for each node $v \in V$.
- In modern GNNs, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $N(v)$,
+ the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$.
+ In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $N(v)$,
   and then combining it with the node's current embedding $h_v$.
 % 
-Formally, a GNN can be defined by two phases:
+Formally, a \ac{GNN} can be defined by two phases:
 \begin{equation}
@@
  and decentralized control, and are actively being researched and developed for further applications
+\subsection{Problem formalisation}
+Given the homogeneity and large scale considered and the \emph{locality} (i,e., each agent can only observe its neighbours), 
+ the problem can be modelled through the SwarMDP model~\cite{DBLP:conf/atal/SosicKZK17} -- 
+ an extension of the \ac{DecPOMDP}~\cite{Decpomdp2000} model to swarm-like systems. \ga{todo: define swarm-like systems}
+% 
+A SwarMDP is characterized by a \emph{swarming agent} ($\mathbb{A}$) and the dynamics of the environment ($\mathbb{E}$).
+Specifically, $\mathbb{A}$ is a tuple ($\mathcal{S}, \mathcal{O}, \mathcal{A}, \mathcal{R}, \pi$) where:
+\begin{itemize}
+  \item $\mathcal{S, O, A}$ are the set of local states, observations (or features), and actions, respectively;
+  \item $\mathcal{R}: \mathcal{S} \rightarrow \mathbb{R}$ is the reward function, which is influenced by the environment;
+  \item $\pi: \mathcal{O} \rightarrow \mathcal{A}$ is the policy function, which maps the observations to the actions. It could be deterministic or stochastic.
+\end{itemize}
+Starting from this definition, the environment $\mathbb{E}$ is defined as a tuple ($\mathcal{P}, \mathbb{A}, \mathcal{T}, \xi$), where:
+\begin{itemize}
+  \item $\mathcal{P}$ is the total number of agents in the systems (the agent population), which is assumed to be fixed;
+  \item $\mathbb{A}$ is the above defined agent prototype that rules each agent $v \in P$;
+  \item $\mathcal{T}: \mathcal{S}^P \times \mathcal{A}^P \times \mathcal{S}^P \rightarrow \mathbb{R}$ is the transition  global function, which is influenced by the actions of the agents and returns a collective reward -- this is typically not known by the swarming agents;
+  \item $\xi: \mathcal{S^P} \rightarrow \mathcal{O^P}$ is the global observation model of the systems.
+\end{itemize}
+In swarMDP, the neighbourhood is not directly defined, but it is implicitly defined by the observation model $\xi$.
+Since in our case, the agents can only interact with 1-hop neighbourhoods and they aren't directly influenced by other 
+  agent observations, we restrict the observation model as follows:
+\begin{equation}
+\xi(v): \{s_j, j \in \mathcal{N}^v\} \rightarrow O
+\end{equation}
+\begin{equation}
+\xi = \{\xi(v), v \in \mathcal{P}\}
+\end{equation}
+where $\mathcal{N}^v$ is the set of neighbours of $v$.
 
+This model can be used then to express the evolution of the system in time.
+Specifically, starting from a global state $\mathcal{S}^P_t$, the next state $\mathcal{S}^P_{t+1}$ is defined as:
+\begin{equation}
+\mathcal{A}^P_t = \pi(\xi(S^P_t))
+\end{equation}
+\begin{equation}
+\mathcal{S}^P_{t+1} = \mathcal{T}(\mathcal{S}^P_t, \mathcal{A}^P_t)
+\end{equation}
+Given a time $t$, the system can be also represented as a graph $G_t = (V_t, E_t)$, 
+  where $E_t$ is built from $\mathcal{N}$.
+  Each node is then decorated with the local observation $o_v$. 
+  Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022}.
 \section{Field-informed Reinforcement Learning}

### 2023-05-05 paper.tex
chore: shrink marl section removing DecPOMPD discussion

@@
 and the need for a scalable approach. In doing this, we will discuss some of the existing approaches (declarative, RL and )}
-\subsection{Problem definition}
+\subsection{System definition}
 This article focuses on the definition of \emph{decentralized collective intelligence} within the context of \emph{large-scale} distributed systems. 
@@
  Such scenarios are referred to as multi-agent reinforcement learning (MARL)~\cite{zhang2019marl}. 
- While standard \ac{RL} approaches use Markov decision processes to model the world, 
- in MARL for distributed systems, the notion of 
- \acp{DecPOMDP}~\cite{Decpomdp2000} is used.
-
-A \ac{DecPOMDP} represents a sequential game, 
- in which each agent receives a local observation about the world they 
- inhabit and receives a joint immediate reward in response to the actions they take. 
+% While standard \ac{RL} approaches use Markov decision processes to model the world, 
+% in MARL for distributed systems, the notion of 
+% \acp{DecPOMDP}~\cite{Decpomdp2000} is used.
 %
-\ga{Improve, if necessary, the following paragraph.}
-Formally, a \ac{DecPOMDP} is defined as a tuple $\langle N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma \rangle$, 
- where $N$ is the total number of agents that live in the environment, 
- $S$ is the state set of the environment, 
- $\{A_i\}$ is the set of actions for agent $i$ and $A$ is the joint action space, 
- $P$ is the probability transition function, 
- $R$ is the immediate joint reward function, and 
- $\{\Omega_i\}$ is the set observation function for the agent $i$,
- $O$ is the observation space, and finally $\gamma$ is the discount factor.
+%A \ac{DecPOMDP} represents a sequential game, 
+% in which each agent receives a local observation about the world they 
+% inhabit and receives a joint immediate reward in response to the actions they take. 
 % 
-On top of this model, each agent tries to learn a policy $\pi_i$ in order 
- to maximise the long-term joint reward signal.
-In the case of some of these concepts, 
- there are several approaches/classifications in the literature, 
- like cooperative (i.e., the reward function incentivizes collective outcomes)
- heterogeneous (i.e., the agents have different action spaces and policies) and
- homogenous (i.e., the agents have the same action space and policy) settings.
+%\ga{Improve, if necessary, the following paragraph.}
+%Formally, a \ac{DecPOMDP} is defined as a tuple $\langle N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma \rangle$, 
+% where $N$ is the total number of agents that live in the environment, 
+% $S$ is the state set of the environment, 
+% $\{A_i\}$ is the set of actions for agent $i$ and $A$ is the joint action space, 
+% $P$ is the probability transition function, 
+% $R$ is the immediate joint reward function, and 
+% $\{\Omega_i\}$ is the set observation function for the agent $i$,
+% $O$ is the observation space, and finally $\gamma$ is the discount factor.
+%
+%On top of this model, each agent tries to learn a policy $\pi_i$ in order 
+% to maximise the long-term joint reward signal.
+%In the case of some of these concepts, 
+% there are several approaches/classifications in the literature, 
+% like cooperative (i.e., the reward function incentivizes collective outcomes)
+% heterogeneous (i.e., the agents have different action spaces and policies) and
+% homogenous (i.e., the agents have the same action space and policy) settings.
 %  

### 2023-05-05 paper.tex
chore: add motivation, background draft done

@@
   where we use computational \emph{computational fields} to manage the interaction between agents 
-  in stigmergically and Graph Neural Networks to learn a local behaviour necessary to solve collective tasks. 
+  and Graph Neural Networks to learn a local behaviour necessary to solve collective tasks. 
 %
@@
 % 
-We demonstrate the effectiveness of this new approach in several case studies 
+We demonstrate the effectiveness of this new approach in a simulated use case 
  where coordination tasks are successfully solved. \lukas{do we really have multiple case studies?} 
@@
 %  
-In particular, in this work, we consider homogenous many agent RL~\cite{yang2021many}, 
+In particular, in this work, we consider homogenous \ac{MAARL}~\cite{yang2021many}, 
  where $N \gg 2$ and each agent is interchangeable and indistinguishable.
@@
   Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022}.
+\subsection{Motivation}
+The definition of distributed controllers for intelligent systems can be divided into two major approaches: 
+ \emph{manual design}, where the application is conceived and described through a programming language, 
+ and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesize distributed controllers. 
+ Our work lies between these two approaches, particularly in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
+
+Compared to earlier works, 
+ our approach builds on the concept of co-fields, 
+ where agents exploit the field as a stigmergic medium to receive system-wide information and apply reasoning to this data.
+%  
+We present a subsequent approach where agent intelligence is synthesized through \ac{MAARL} and the use of \ac{GNN} to learn a local representation from the neighbourhood.
+%
+By learning a smart policy directly in the environment, the agent becomes capable of adapting its behaviour to new situations.
+
+The use of \acp{GNN} as part of a distributed controller has been explored in previous literature, 
+ where it was shown that they could be used to break down the evaluation of local and distributed systems. 
+%
+However, in these works, communication was left entirely to the neural network, 
+ making the learning process potentially more complex and unstable. 
+ In our approach, the \ac{GNN} is \emph{informed} by computational fields that collect the necessary information to compute a certain task, 
+ limiting learning to only the specific task defined by a collective reward function.
+ This will speed up the learning process and make it more stable.
 \section{Field-informed Reinforcement Learning}

### 2023-05-05 paper.tex
chore: define swarm-like systems

@@
 Finally, we present a formalization of the problem and a discussion of the motivations behind our proposed work.
-%Field-based RL draws upon research areas that were not previously considered in traditional RL. 
-%
-%This section provides an overview of three key concepts that underpin this approach, 
-% namely 
-% field-based coordination --i.e., coordinating multi-agent systems through computational fields--, 
-% \ac{GNN} --i.e., a novel type of neural network architecture designed to process graph-structured data--, 
-% and \ac{MAARL} ---i.e., a large number of agents learning a collective policy to achieve a global goal.
-
 \ga{Page budget: 1/2 page \\}
@@
 \subsection{System definition}
-This article focuses on the definition of \emph{decentralized collective intelligence} within the context of \emph{large-scale} distributed systems. 
-%
-Specifically, we consider systems of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator}
+This article focuses on the definition of \emph{decentralized collective intelligence} within the context of \emph{large-scale} distributed systems, 
+ specifically what we call \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}
+ that are systems of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator}
  each of which can only interact with the \emph{1-hop} neighbourhood that can 
@@
  Such scenarios are referred to as multi-agent reinforcement learning (MARL)~\cite{zhang2019marl}. 
-% While standard \ac{RL} approaches use Markov decision processes to model the world, 
-% in MARL for distributed systems, the notion of 
-% \acp{DecPOMDP}~\cite{Decpomdp2000} is used.
-%
-%A \ac{DecPOMDP} represents a sequential game, 
-% in which each agent receives a local observation about the world they 
-% inhabit and receives a joint immediate reward in response to the actions they take. 
-% 
-%\ga{Improve, if necessary, the following paragraph.}
-%Formally, a \ac{DecPOMDP} is defined as a tuple $\langle N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma \rangle$, 
-% where $N$ is the total number of agents that live in the environment, 
-% $S$ is the state set of the environment, 
-% $\{A_i\}$ is the set of actions for agent $i$ and $A$ is the joint action space, 
-% $P$ is the probability transition function, 
-% $R$ is the immediate joint reward function, and 
-% $\{\Omega_i\}$ is the set observation function for the agent $i$,
-% $O$ is the observation space, and finally $\gamma$ is the discount factor.
-%
-%On top of this model, each agent tries to learn a policy $\pi_i$ in order 
-% to maximise the long-term joint reward signal.
-%In the case of some of these concepts, 
-% there are several approaches/classifications in the literature, 
-% like cooperative (i.e., the reward function incentivizes collective outcomes)
-% heterogeneous (i.e., the agents have different action spaces and policies) and
-% homogenous (i.e., the agents have the same action space and policy) settings.
 %
@@
  the problem can be modelled through the SwarMDP model~\cite{DBLP:conf/atal/SosicKZK17} -- 
- an extension of the \ac{DecPOMDP}~\cite{Decpomdp2000} model to swarm-like systems. \ga{todo: define swarm-like systems}
+ an extension of the \ac{DecPOMDP}~\cite{Decpomdp2000} model to swarm-like systems.
 % 
@@
 Compared to earlier works, 
- our approach builds on the concept of co-fields, 
+ our approach builds on the concept of co-fields~\cite{DBLP:journals/pervasive/MameiZL04}, 
  where agents exploit the field as a stigmergic medium to receive system-wide information and apply reasoning to this data.

### 2023-05-06 paper.tex
incipit on field coordination approach

@@
 \ga{discussion about field used as a coordination mechanism in multi-agent systems. Introduction to AC}
+The field-based coordination approaches utilize a concept of \emph{computational fields} 
+ (or simply \emph{fields}), 
+ which are \emph{distributed} data structures that associate locations with values and evolve over time. 
+%
+A modern field-coordination approach called \emph{aggregate computing}~\cite{Beal2015Computer}
+ is rooted in earlier work on \emph{artificial potential fields}~\cite{DBLP:conf/icra/Warren89} and \emph{co-fields}~\cite{DBLP:journals/pervasive/MameiZL04}. 
+% 
+ In this \emph{macro-programming}~\cite{DBLP:journals/corr/abs-2201-03473} paradigm is possible to devise collective and self-organizing behaviour through a composition of functions operating on fields.
+%
+The fields map a set of individual agents to computational values, 
+ allowing them to associate what they sense, process, and actuate. 
+%
+Fields are computed locally but subject to a global viewpoint, 
+ enabling emergent collective behaviour through the interplay of the system model (similar to the above described) 
+ and the \emph{programming model} known as field calculus.
 \subsection{Graph Neural Networks}

### 2023-05-07 paper.tex
chore: minors

@@
 We further combine this with reinforcement learning techniques to respond to changing conditions in the environment.
+%
+This follow a novel trends in which high-level declarative programming language are mixed with machine learning technique in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
 %Specifically, we will show... \lukas{more info...}
@@
  enabling emergent collective behaviour through the interplay of the system model (similar to the above described) 
- and the \emph{programming model} known as field calculus.
+ and the \emph{programming model} known as field calculus~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}.
 \subsection{Graph Neural Networks}
 Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
-  simple \ac{MLP} to \ac{CNN}~\cite{oshea2015introduction}.
+  simple \ac{MLP}~\cite{werbos1975beyond} to \ac{CNN}~\cite{oshea2015introduction}.
 %
@@
  and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesize distributed controllers. 
- Our work lies between these two approaches, particularly in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
+%
+Our work lies between these two approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
 
-Compared to earlier works, 
+Compared to earlier works on field-based coordination, 
  our approach builds on the concept of co-fields~\cite{DBLP:journals/pervasive/MameiZL04}, 
@@
 
-The use of \acp{GNN} as part of a distributed controller has been explored in previous literature, 
+The use of \acp{GNN} as part of a distributed controller has been explored in previous literature~\cite{DBLP:conf/icra/GosrichMLPYR022,DBLP:conf/corl/TolstayaGPP0R19}, 
  where it was shown that they could be used to break down the evaluation of local and distributed systems. 

### 2023-05-07 paper.tex
on learning dynamics

@@
 \usepackage{amsmath}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
 \def\doi#1{\href{https://doi.org/\detokenize{#1}}{\url{https://doi.org/\detokenize{#1}}}}
@@
 \acrodef{CNN}[CNN]{convolutional neural network}
+\acrodef{CTDE}[CTDE]{centralized training and decentralized execution}
+\acrodef{DQN}[DQN]{Deep Q-Learning}
 \ga{Page limit: 10 pages (include refs)}
@@
 but it could be seen as a global evolution, therefore we can use global information to inform the local controller.}
-\subsection{System model}
+\begin{figure*}
+  \includegraphics[width=\textwidth]{images/architecture.pdf}
+  \label{fig:architecture}
+  \caption{High-level description of Field-Informed reinforcement learning dynamics}
+\end{figure*}
+In this section, we aim to clarify the components involved in our proposed solution (i.e., architecture), how these components interact with each other to bring the system to perform the collective behavior (i.e dynamics), and finally, we will detail the learning algorithm designed and used to synthesize the policy (learning algorithm).
+\subsection{Architecture and System dynamics}
 \ga{Plan: 
@@
 }
-\subsection{Learning dynamics}
+The proposed solution consists mainly of two parts, 
+\emph{i)} the aggregate program used to create part of the observation and 
+\emph{ii)} the policy $\pi_{gnn}$ learned through \ac{GNN}-based learning. 
+%
+Let $\Gamma$ be the aggregate program that runs against $G_t$. 
+%
+The evaluation of $\Gamma$  produce a field value $\theta^v_t$ for each node $v$ in $G_t$. 
+The feature vector $o_v$ of which is composed the graph $G'_t$ used by the policy $\pi_{gnn}$ is described by:
+ $o_v = (\theta^v_t, \omega)$.
+ where $\omega$ are other local information gather from sensors (like temperature, distances from neighbourhood, etc.). 
+%
+The policy $\pi_{gnn}$ is then evaluated for each node, 
+ producing an action $a_t$ that will then modify the global state of the system.
+ \ga{todo: put pseudo code to express this dynamics}
+%
+Despite appearing centralized, this system proposes a completely decentralized execution. 
+%
+In fact, the program $\Gamma$ is proactively executed at every node, 
+and the \ac{GNN} can be locally evaluated using only neighborhood information. 
+%
+We want to emphasize that, in this case, the \acp{GNN} must be 1-hop; 
+otherwise, they could not have a local interpretation for each node, according to our system model.
+\subsection{Learning algorithm}
 \ga{
@@
 }
+In swarm-like systems, 
+ the typical approach for learning follows a strategy known as \ac{CTDE}. \ga{perhaps it should be introduced in many RL} 
+The idea is to learn a policy at simulation time where there is a collective view of the system, 
+ and then at runtime use the policy found with global information. 
+This approach allows policies that are influenced by global information but only require local information to function at runtime. 
+%
+The typical approach in such cases is based on actor-critic systems~\cite{DBLP:conf/nips/LoweWTHAM17,wu2022more,song2022ctds,song2022centralized},
+  where the \emph{actor} is the distributed policy (with only local information) and the critic is a neural network that takes the overall system state.
+
+In our case, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighborhood. 
+%
+Therefore, we relied on the value-based approach \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications: 
+\begin{enumerate}
+  \item experience replay stores experiences in the form of \emph{graphs},
+  \item the neural network used to compute the Q function is based on a \ac{GNN} with an \ac{MLP} downstream.
+\end{enumerate}
+The first point is a natural extension due to the fact that we work on graphs rather than simple values. 
+%
+For the second point,
+ the use of \acp{GNN} allows us to define policies on a variable neighborhood  that is essential in such systems as this can vary based on the neighborhood policy,
+ and it is known that \acp{GNN} have a certain ability to generalize to new structures and scale with different nodes. 
+% 
+Additionally, 
+ using the overall graph compared to local experiences makes learning more stable as it reduces the non-stationarity of the environment perceived by each node.
+% 
+ During learning, it is as if we have a \emph{single} agent collectively choosing which action to take (i.e., resulting action graph). 
+%
+The algorithm is described in the following pseudocode.
+
 \section{Evaluation}

### 2023-05-07 paper.tex
minor: add the dqn algorithm description

@@
 \usepackage{amsmath}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
+\usepackage[ruled]{algorithm2e}
 \def\doi#1{\href{https://doi.org/\detokenize{#1}}{\url{https://doi.org/\detokenize{#1}}}}
@@
 %
-The algorithm is described in the following pseudocode.
+\ga{todo: describe the algorithm with the modifications}
+\begin{algorithm}
+  \KwIn{Environment $\mathcal{E}$, graph replay buffer $\mathcal{D}$, target network $\theta^{-}$, current network $\theta$, exploration strategy $\epsilon$}
+  \KwOut{Trained DQN model $\theta$}
+
+  Initialize $\mathcal{D}$ with random initial transitions;
+
+  Initialize $\theta$ with random weights;
+
+  Set $\theta^{-} \leftarrow \theta$;
+
+  \While{not done}{
+  Observe current state $s$;
+
+  With probability $\epsilon$ select a random action $a$, otherwise select $a = \arg\max_{a'} Q(s,a';\theta)$;
+
+  Execute action $a$ in the environment $\mathcal{E}$ and observe the reward $r$ and the next state $s'$;
+
+  Store transition $(s,a,r,s')$ in $\mathcal{D}$;
+
+  Sample a batch of transitions $(s_i,a_i,r_i,s'i)$ from $\mathcal{D}$;
+
+  Compute the target Q-value for each transition: $y_i = r_i + \gamma \max{a'} Q(s'i,a';\theta^{-})$;
+
+  Update the current network weights using gradient descent: $\theta \leftarrow \theta - \alpha \nabla{\theta} \frac{1}{N} \sum_i (y_i - Q(s_i,a_i;\theta))^2$;
+
+  Every $C$ steps, update the target network weights: $\theta^{-} \leftarrow \theta$
+  ;
+  }
+  \caption{Deep Q-Network (DQN)}
+  \end{algorithm}
 

### 2023-05-08 paper.tex
chore: add draft for evaluation

@@
 }
+To test the effectiveness of the proposed approach, 
+ we decided to experiment with a case study related to swarm robotics, 
+ specifically the tracking and coverage of a spatiotemporal phenomenon, cf. tracking a wildfire 
+ or monitoring the water levels in a canal.
+
 \subsection{Scenario}
+The proposed simulation consists 
+ of a set of robots placed in a 2D space capable of perceiving the intensity 
+ of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
+ (e.g., camera, temperature sensor, etc.). 
+%
+Additionally, each node has a coverage range $\omega$ 
+ that describes the area it can monitor. 
+ Each drone can only communicate directly with its own neighborhood $\mathcal{N}$, 
+ which in this case depends on a $\mathbb{O}$ range,
+From the neighbourhood, 
+ each node can also perceive the direction of the other nodes.
+%
+Each node moves following a certain action composed of two components ($\alpha$, $\theta$) 
+ which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
+\subsection{Goal}
+ The objective of this scenario is to:
+\begin{itemize}
+\item Maximize the number of drones within the phenomena
+\item Minimize the number of drones without a neighbourhood
+\item Maximize the coverage of the system
+\end{itemize}
+
+As we are modelling a reinforcement learning system, 
+ these three components must be encoded in a reward function 
+ that provides an estimate of the current action taken by a given robot.
+
+ Regarding the first point, the reward function is simple and is defined as:
+
+ \begin{equation*}
+ R^a_{v} = \begin{cases}
+  1 & \text{if } \zeta_v > 0 \\
+  0 & \text{otherwise} 
+ \end{cases}
+ \end{equation*}
+
+This will lead the system to prefer a configuration in which every node is present in the system. 
+%
+The second point is important because if the system breaks into many scattered drones, 
+ they are not able to move in the environment due to the reduced observability of the system. 
+ In this case, the reward is defined as:
+
+ \begin{equation*}
+ R^{\mathcal{N}}_{v} = \begin{cases}
+  1 & \text{if } \mathcal{N} > 0 \\
+  0 & \text{otherwise} 
+ \end{cases}
+ \end{equation*}
+
+ Finally, to maximize the coverage, 
+  we decided to calculate the maximum and minimum distance of the neighbourhood and give more value in case these two values are similar. 
+%
+This means that the average distance will tend towards the one expressed by the viewing range of each node:
+% 
+\begin{equation*}
+R^{C}_{v} = \frac{d_{max} - d_{min}}{\mathbb{O}}
+\end{equation*}
+The final reward function is defined as:
+\begin{equation*}
+R_{v} = R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}
+\end{equation*}
 

### 2023-05-10 paper.tex
chore: remove different variants
chore: update the description of dqn with gnn

@@
 \begin{figure*}
-  \includegraphics[width=\textwidth]{images/architecture.pdf}
+  \includegraphics[width=\textwidth]{imgs/architecture.pdf}
   \label{fig:architecture}
@@
 \begin{algorithm}
-  \KwIn{Environment $\mathcal{E}$, graph replay buffer $\mathcal{D}$, target network $\theta^{-}$, current network $\theta$, exploration strategy $\epsilon$}
+  \KwIn{Environment $\mathbb{E}$, graph replay buffer $\mathcal{D}$, target network $\theta^{-}$, current network $\theta$, exploration strategy $\epsilon$}
   \KwOut{Trained DQN model $\theta$}
@@
   \While{not done}{
-  Observe current state $s$;
+  Observe current graph observations $G_o$;
 
-  With probability $\epsilon$ select a random action $a$, otherwise select $a = \arg\max_{a'} Q(s,a';\theta)$;
+  With probability $\epsilon$ select a random graph action $G_a$, 
+  otherwise select $G_a =$ 
+  $\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;
 
-  Execute action $a$ in the environment $\mathcal{E}$ and observe the reward $r$ and the next state $s'$;
+  Execute the collective action $G_a$ in the environment $\mathbb{E}$ and observe a graph-level reward $G_r$ and the next state $G_o'$;
 
-  Store transition $(s,a,r,s')$ in $\mathcal{D}$;
+  Store transition $(G_o,G_a,G_r, G_o')$ in $\mathcal{D}$;
 
-  Sample a batch of transitions $(s_i,a_i,r_i,s'i)$ from $\mathcal{D}$;
+  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G^i_o)$ 
+   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$;
 
-  Compute the target Q-value for each transition: $y_i = r_i + \gamma \max{a'} Q(s'i,a';\theta^{-})$;
+  Compute the target Q-value for each node in the graph: 
+  $y = G_r + \gamma * \max{a'} Q(s'i,a';\theta^{-})$;
 
-  Update the current network weights using gradient descent: $\theta \leftarrow \theta - \alpha \nabla{\theta} \frac{1}{N} \sum_i (y_i - Q(s_i,a_i;\theta))^2$;
+  Compute the current expetected value for each node in the graph:
+  $y^* = Q(s'i,a';\theta)$;
+
+  Update the current network weights using gradient descent: $\theta \leftarrow \theta - \alpha \nabla{\theta} \frac{1}{|G_o|} (y - y^*)^2$;
 
@@
   }
-  \caption{Deep Q-Network (DQN)}
+  \caption{Deep Q-Network (DQN) with GNN and Graph Replay Buffer}
   \end{algorithm}
@@
 
-\paragraph*{Variant A}
-\paragraph*{Variant B}
-\paragraph*{Variant C}
+%\paragraph*{Variant A}
+%\paragraph*{Variant B}
+%\paragraph*{Variant C}
 

### 2023-05-10 paper.tex
chore: minor in the fig 1 description, add manual vs automatic design in the abstract

@@
  where the scale of the systems considered makes the definition of collective behaviours even more challenging. 
-
-In this paper, we consider a new reinforcement learning approach 
-  for the definition of intelligent agents called \emph{Field-Informed reinforcement learning}, 
+% 
+Thus far, two main macro approaches have been proposed for creating distributed controllers: 
+ \emph{manual design}, in which programmers build the controllers directly, and 
+ \emph{automatic design}, which involves synthesizing programs using machine learning methods.
+%
+In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning}, 
   where we use computational \emph{computational fields} to manage the interaction between agents 
-  and Graph Neural Networks to learn a local behaviour necessary to solve collective tasks. 
+  and \emph{Graph Neural Networks} to learn a local behaviour necessary to solve collective tasks. 
 %
@@
   \label{fig:architecture}
-  \caption{High-level description of Field-Informed reinforcement learning dynamics}
+  \caption{High-level description of Field-Informed reinforcement learning dynamics. 
+  For each time step, a graph is constructed from the environment, and it will associate each node with a local feature (hexagons in the picture). 
+  Using this feature, an aggregate program computes spatio-temporal information that enhances the feature set of each node, depicted as colours in the middle graph. 
+  Finally, utilizing the GNN, actions are computed for each node in the system to be performed against the environment, enabling advancement in the simulations according to swarMDP rules.
+}
 \end{figure*}

### 2023-05-10 paper.tex
some intro, background and BE fix

--- 2023-05-10 paper.tex
@@
 \section{Introduction}
+
+Autonomous agents interacting and operating towards a common task, rely on shared information and consequently emerging knowledge in order to achieve the underlying goals efficiently.
+Coordinating and organising this information exchange within the network of agents is required for improve the collaboration among them further~\cite{jennings1996coordination}.
+Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complimentary competences of the individual agents, or (iii) the collective resources of multiple agents.
+
 Various phenomena in the real world are not evenly distributed across the environment such as watershed, wild fires, traffic, human crowd movement, or wild life movement from insects, fish, birds, and mammals.
@@
 Specifically, 
- our approach utilizes aggregate computing to disseminate information by manipulating a computational field, 
+ our approach utilises aggregate computing to disseminate information by manipulating a computational field, 
  which is a distributed data structure that maps devices to information. 
@@
 \subsection{System definition}
-This article focuses on the definition of \emph{decentralized collective intelligence} within the context of \emph{large-scale} distributed systems, 
- specifically what we call \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}
- that are systems of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator}
- each of which can only interact with the \emph{1-hop} neighbourhood that can 
- be logical (i.e., a communication network) or physical (i.e., a spatial neighbourhood).
+This article focuses on the definition of \emph{decentralised collective intelligence} within the context of \emph{large-scale} distributed systems.
+ Specifically, we focus on \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}, i.e.,  systems consisting of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator} able to interact with the real world.
+ Each of these nodes can interact with its direct neighbours, either based on physical (e.g. communication range) or logical distance (e.g. 1-hop neighbourhood).
+ \lukas{does the article focus on the definition of decentralised collective intelligence?}
+  \lukas{distributed or decentralised?}
+
 %
-Each node executes a \emph{local} controller that can only access the information available to its \emph{context}
- that is the information available in the neighbourhood and the information available from its sensors.
+Each node is governed by an autonomous software agent, executing a \emph{local} control loop.
+In each iteration of this loop, the agent can access the information available within its own \emph{context}.
+This context is comprised by information acquired by the node directly through its sensors and the information received from its neighbours.
 %
-After each program evaluation, 
- the node can send messages to its neighbourhood about its current evaluation and update its internal state.
+At the end of each control loop iteration, the node can send messages to its neighbourhood. This message may contain raw sensor data or aggregated information over time or from other neighbours.
+% about its current evaluation and update its internal state.
 %
-This general model can be adapted to several different scenarios, 
- such as a network of sensors, a network of robots (e.g., swarm robotics), or a network of mobile phones.
+%This general model can be adapted to several different scenarios, such as a network of sensors, a network of robots (e.g., swarm robotics), or a network of mobile phones.
+Examples of such large scale system can be a network of robots (e.g., swarm robotics), camera and IoT networks, or a network of mobile phones.
 %
@@
  the system to achieve a certain \emph{collective} requirements through \emph{cooperation}, 
- such as spatial area coverage, phenomena tracking, and robot aggregation.
+ such as spatial area coverage, phenomena tracking, and node aggregation.
 %
@@
 %
-\subsection{Field Coordination Approaches}
+\lukas{not really a system definition - maybe we rename this to \emph{Swarm Systems}?}
+\subsection{Field-based Coordination}
 \ga{discussion about field used as a coordination mechanism in multi-agent systems. Introduction to AC}
-The field-based coordination approaches utilize a concept of \emph{computational fields} 
+The field-based coordination approaches utilise a concept of \emph{computational fields} 
  (or simply \emph{fields}), 
@@
 %
-A modern field-coordination approach called \emph{aggregate computing}~\cite{Beal2015Computer}
- is rooted in earlier work on \emph{artificial potential fields}~\cite{DBLP:conf/icra/Warren89} and \emph{co-fields}~\cite{DBLP:journals/pervasive/MameiZL04}. 
+\emph{Aggregate computing}~\cite{Beal2015Computer}, a modern field-coordination approach, is rooted in earlier work on \emph{artificial potential fields}~\cite{DBLP:conf/icra/Warren89} and \emph{co-fields}~\cite{DBLP:journals/pervasive/MameiZL04}.
 % 
- In this \emph{macro-programming}~\cite{DBLP:journals/corr/abs-2201-03473} paradigm is possible to devise collective and self-organizing behaviour through a composition of functions operating on fields.
+Within aggregate computing, the  \emph{macro-programming}~\cite{DBLP:journals/corr/abs-2201-03473} paradigm can be realised to devise collective and self-organizing behaviour through a composition of functions operating on fields.
 %
@@
  and the \emph{programming model} known as field calculus~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}.
+\lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
 \subsection{Graph Neural Networks}
-Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
-  simple \ac{MLP}~\cite{werbos1975beyond} to \ac{CNN}~\cite{oshea2015introduction}.
-%
-Whereas \ac{CNN} is designed to be used with image and spatial-like data (e.g., audio),
- \ac{GNN} is a novel model used to process graph-structured data. 
+%Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
+%  simple \ac{MLP}~\cite{werbos1975beyond} to \ac{CNN}~\cite{oshea2015introduction}.
+%  \lukas{more examples? }
+%%
+%Whereas \ac{CNN} is designed to be used with image and spatial-like data (e.g., audio), 
+ \ac{GNN} is a novel neural network model used to process graph-structured data with deep learning approaches.
 %
@@
 where $E$ defines the neighbourhood relations for each participating node, 
-and $V$ identifies the nodes present in the graph at that time. 
+and $V$ identifies the nodes present in the graph. %at that time.  \lukas{no time in the definition}
 % 
-Each node $v$ is associated with an observation (or feature set), $o_v$, 
+Each node $v \in V$ is associated with an observation (or feature set) $o_v$.
 %
-Given $o_v$ a feature vector associated with each node $v \in V$, 
+Given $o_v$, %a feature vector associated with each node $v \in V$, 
  the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$.
+The node embedding $h_v$ describes the node in the network and summarise the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
  In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $N(v)$,
   and then combining it with the node's current embedding $h_v$.
+  \lukas{maybe add:``The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased.''}
 % 
@@
 \end{equation}
-where $h_{v}^{k}$ is the embedding of node $v$ at the $k$-th layer ($h_v^0 = o_v$), 
- $\mathcal{N}(v)$ is the set of neighbors of node $v$ computed from $E$.
+where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer ($h_v^0 = o_v$), 
+ $\mathcal{N}(v)$ is the set of neighbours of node $v$ computed from $E$.
 %
-The differential part come into play in the $\operatorname{COMBINE}$ function, 
+The differential part comes into play in the $\operatorname{COMBINE}$ function, 
  which is usually a differentiable function such as a neural network layer.
@@
 $\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
-  and it could be a simple sum or a more complex function such as a neural network layer.
+  and it could be a simple sum, max or sum of products. More complex aggregation functions are available~\cite{pellegrini2020learning}.
 %
@@
 
-\acp{GNN} are used in several fields such as social network analysis, chemistry, and physics.
+\acp{GNN} are used in several application areas such as social network analysis, chemistry, and physics.
 In this paper, we use GNNs to learn a local behaviour for each agent in a multi-agent system (more details in Section~\ref{sec:approach}).
@@
  performs actions in an environment (i.e., everything outside the agent) according to a policy, 
- to maximize long-term rewards.
+ to maximise long-term rewards (or minimise a regret).
 
@@
  have started to consider the use of deep learning approaches to try 
- to synthesize robust controllers capable of generalizing to new tasks. 
+ to synthesise robust controllers capable of generalizing to new tasks. 
 In this context, mean-field reinforcement learning~\cite{pmlr-v80-yang18d} is certainly noteworthy. 
@@
 These approaches have shown promising results in various domains, such as multi-agent coordination 
- and decentralized control, and are actively being researched and developed for further applications
+ and decentralised control, and are actively being researched and developed for further applications
 \subsection{Problem formalisation}
@@
 % 
-A SwarMDP is characterized by a \emph{swarming agent} ($\mathbb{A}$) and the dynamics of the environment ($\mathbb{E}$).
+A SwarMDP is characterised by a \emph{swarming agent} ($\mathbb{A}$) and the dynamics of the environment ($\mathbb{E}$).
 Specifically, $\mathbb{A}$ is a tuple ($\mathcal{S}, \mathcal{O}, \mathcal{A}, \mathcal{R}, \pi$) where:
@@
  \emph{manual design}, where the application is conceived and described through a programming language, 
- and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesize distributed controllers. 
+ and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesise distributed controllers. 
 %
@@
 %  
-We present a subsequent approach where agent intelligence is synthesized through \ac{MAARL} and the use of \ac{GNN} to learn a local representation from the neighbourhood.
+We present a subsequent approach where agent intelligence is synthesised through \ac{MAARL} and the use of \ac{GNN} to learn a local representation from the neighbourhood.
 %
@@
 \end{figure*}
-In this section, we aim to clarify the components involved in our proposed solution (i.e., architecture), how these components interact with each other to bring the system to perform the collective behavior (i.e dynamics), and finally, we will detail the learning algorithm designed and used to synthesize the policy (learning algorithm).
+In this section, we aim to clarify the components involved in our proposed solution (i.e., architecture), how these components interact with each other to bring the system to perform the collective behavior (i.e dynamics), and finally, we will detail the learning algorithm designed and used to synthesise the policy (learning algorithm).
 \subsection{Architecture and System dynamics}
@@
 %
-Despite appearing centralized, this system proposes a completely decentralized execution. 
+Despite appearing centralised, this system proposes a completely decentralised execution. 
 %
 In fact, the program $\Gamma$ is proactively executed at every node, 
-and the \ac{GNN} can be locally evaluated using only neighborhood information. 
+and the \ac{GNN} can be locally evaluated using only neighbourhood information. 
 %
-We want to emphasize that, in this case, the \acp{GNN} must be 1-hop; 
+We want to emphasise that, in this case, the \acp{GNN} must be 1-hop; 
 otherwise, they could not have a local interpretation for each node, according to our system model.
@@
 
-In our case, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighborhood. 
+In our case, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighbourhood. 
 %
@@
 For the second point,
- the use of \acp{GNN} allows us to define policies on a variable neighborhood  that is essential in such systems as this can vary based on the neighborhood policy,
- and it is known that \acp{GNN} have a certain ability to generalize to new structures and scale with different nodes. 
+ the use of \acp{GNN} allows us to define policies on a variable neighbourhood  that is essential in such systems as this can vary based on the neighbourhood policy,
+ and it is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes. 
 % 
@@
 
-  Initialize $\mathcal{D}$ with random initial transitions;
+  Initialise $\mathcal{D}$ with random initial transitions;
 
-  Initialize $\theta$ with random weights;
+  Initialise $\theta$ with random weights;
 
@@
  we decided to experiment with a case study related to swarm robotics, 
- specifically the tracking and coverage of a spatiotemporal phenomenon, cf. tracking a wildfire 
+ specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
  or monitoring the water levels in a canal.
@@
  that describes the area it can monitor. 
- Each drone can only communicate directly with its own neighborhood $\mathcal{N}$, 
+ Each drone can only communicate directly with its own neighbourhood $\mathcal{N}$, 
  which in this case depends on a $\mathbb{O}$ range,
@@
 \begin{itemize}
-\item Maximize the number of drones within the phenomena
-\item Minimize the number of drones without a neighbourhood
-\item Maximize the coverage of the system
+\item Maximise the number of drones within the phenomena
+\item Minimise the number of drones without a neighbourhood
+\item Maximise the coverage of the system
 \end{itemize}
@@
 
- Finally, to maximize the coverage, 
+ Finally, to maximise the coverage, 
   we decided to calculate the maximum and minimum distance of the neighbourhood and give more value in case these two values are similar. 

### 2023-05-10 paper.tex
feedback on section 2;

@@
 In particular, in this work, we consider homogenous \ac{MAARL}~\cite{yang2021many}, 
- where $N \gg 2$ and each agent is interchangeable and indistinguishable.
+ where the set of agents $N \gg 2$ and each agent is interchangeable and indistinguishable.
 %
@@
  which is found in collective adaptive systems such as swarm robotics.
+ \lukas{i am not sure i understand the previous sentence.}
 
-In fact, the first works in this direction were discussed precisely in the last field, 
+In fact, the first works in this direction were discussed precisely in the last field, \lukas{what does this mean `in the last field'?}
  exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating 
- a policy equal to the entire system. 
-Modern approaches, however, 
- have started to consider the use of deep learning approaches to try 
+ a policy equal to \lukas{maybe `valid to' or `representing' instead of `equal to'?}  the entire system. 
+%Modern approaches, however, 
+% have started to consider the use of 
+ Recently, deep learning approaches have been considered 
  to synthesise robust controllers capable of generalizing to new tasks. 
-In this context, mean-field reinforcement learning~\cite{pmlr-v80-yang18d} is certainly noteworthy. 
- Mean-field RL is a technique for learning approximate inference in large-scale systems, 
- where the interactions between the agents can be decoupled into a single conditional probability distribution function for each agent, 
+%In this context, mean-field reinforcement learning~\cite{pmlr-v80-yang18d} is certainly noteworthy. 
+ Mean-field RL~\cite{pmlr-v80-yang18d}  is a technique for learning approximate inference in large-scale systems.
+Here, the interactions between the agents can be decoupled into a single conditional probability distribution function for each agent,
  thus simplifying the learning process and reducing the computational cost.
@@
 Some known approaches using mean-field reinforcement learning include Q-mean, 
- which is an extension of Q-learning to mean-field settings, 
- and actor-critic mean-field, which combines actor-critic algorithms with mean-field approximations. 
+ which is an extension of Q-learning to mean-field settings~\cite{}, 
+ and actor-critic mean-field~\cite{}, which combines actor-critic algorithms with mean-field approximations. 
 %
 These approaches have shown promising results in various domains, such as multi-agent coordination 
- and decentralised control, and are actively being researched and developed for further applications
+ and decentralised control, and are actively being researched and developed for further applications.
 \subsection{Problem formalisation}
-Given the homogeneity and large scale considered and the \emph{locality} (i,e., each agent can only observe its neighbours), 
+Given the homogeneity, large system scale, and the \emph{locality} (i,e., each agent can only observe its neighbours), 
  the problem can be modelled through the SwarMDP model~\cite{DBLP:conf/atal/SosicKZK17} -- 
- an extension of the \ac{DecPOMDP}~\cite{Decpomdp2000} model to swarm-like systems.
+ an extension of the \ac{DecPOMDP}~\cite{Decpomdp2000} model for swarm-like systems.
 % 
@@
   \item $\mathcal{P}$ is the total number of agents in the systems (the agent population), which is assumed to be fixed;
-  \item $\mathbb{A}$ is the above defined agent prototype that rules each agent $v \in P$;
+  \item $\mathbb{A}$ is the defined agent prototype that rules each agent $v \in P$;
   \item $\mathcal{T}: \mathcal{S}^P \times \mathcal{A}^P \times \mathcal{S}^P \rightarrow \mathbb{R}$ is the transition  global function, which is influenced by the actions of the agents and returns a collective reward -- this is typically not known by the swarming agents;
@@
 In swarMDP, the neighbourhood is not directly defined, but it is implicitly defined by the observation model $\xi$.
-Since in our case, the agents can only interact with 1-hop neighbourhoods and they aren't directly influenced by other 
-  agent observations, we restrict the observation model as follows:
+In our specific case, the agents can only interact with 1-hop neighbourhoods and are not directly influenced by other agent observations. We can therefore restrict the observation model as follows:
 \begin{equation}

### 2023-05-11 paper.tex
on lukas question: prefer swarm systems instead of system definition

@@
 We demonstrate the effectiveness of this new approach in a simulated use case 
- where coordination tasks are successfully solved. \lukas{do we really have multiple case studies?} 
+ where coordination tasks are successfully solved. %\lukas{do we really have multiple case studies?} 
 \end{abstract}
@@
 In this section, we aim to provide a concise overview of the system we intend to address, 
- starting with a system definition. 
+ starting with a high-level system model. 
 %
@@
 and the need for a scalable approach. In doing this, we will discuss some of the existing approaches (declarative, RL and )}
-\subsection{System definition}
+\subsection{Swarm systems}
 This article focuses on the definition of \emph{decentralised collective intelligence} within the context of \emph{large-scale} distributed systems.
@@
 %
-\lukas{not really a system definition - maybe we rename this to \emph{Swarm Systems}?}
+%\lukas{not really a system definition - maybe we rename this to \emph{Swarm Systems}?}
 \subsection{Field-based Coordination}

### 2023-05-11 paper.tex
on lukas question: prefer to discuss about collective behaviour instead of distributed (or decentralised) collective intellingence

@@
 \subsection{Swarm systems}
-This article focuses on the definition of \emph{decentralised collective intelligence} within the context of \emph{large-scale} distributed systems.
+This article aims in defining intelligent \emph{collective behaviours} within the context of \emph{large-scale} distributed systems.
  Specifically, we focus on \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}, i.e.,  systems consisting of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator} able to interact with the real world.
  Each of these nodes can interact with its direct neighbours, either based on physical (e.g. communication range) or logical distance (e.g. 1-hop neighbourhood).
- \lukas{does the article focus on the definition of decentralised collective intelligence?}
-  \lukas{distributed or decentralised?}
+ %\lukas{does the article focus on the definition of decentralised collective intelligence?}
+ % \lukas{distributed or decentralised?}
 

### 2023-05-11 paper.tex
on lukas question: clarify the sentence in many agent RL

@@
 This research area is relevant in the context of large-scale systems 
- when the incentive of the individual is reduced compared to the collective behaviour of the system, 
- which is found in collective adaptive systems such as swarm robotics.
- \lukas{i am not sure i understand the previous sentence.}
+ where collective intelligence emerges from local and repeated interaction of simple entities, like in swarm robotics.
+%\lukas{i am not sure i understand the previous sentence.}
 

### 2023-05-11 paper.tex
on lukas question: clarify the field in which is many rl is explored

@@
 
-In fact, the first works in this direction were discussed precisely in the last field, \lukas{what does this mean `in the last field'?}
+In fact, the first works in this direction were discussed precisely swarm robotics, %\lukas{what does this mean `in the last field'?}
  exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating 

### 2023-05-11 paper.tex
on lukas question: prefer representing instead of equal to

@@
  exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating 
- a policy equal to \lukas{maybe `valid to' or `representing' instead of `equal to'?}  the entire system. 
+ a policy representing 
+ %\lukas{maybe `valid to' or `representing' instead of `equal to'?} 
+ the entire system. 
 %Modern approaches, however, 

### 2023-05-11 paper.tex
minor

@@
  the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$.
-The node embedding $h_v$ describes the node in the network and summarise the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
+The node embedding $h_v$ describes the node in the network and summarises the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
  In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $N(v)$,

### 2023-05-11 paper.tex
minor: add hypen in large-scale

@@
 %
-\title{Field-informed Reinforcement Learning for Learning Large Scale Collective Tasks} %% Todo improve
+\title{Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks} %% Todo improve
 % or: Field-Informed Reinforcement Learning: A Scalable and Effective Method for Collective Intelligence
@@
 %This general model can be adapted to several different scenarios, such as a network of sensors, a network of robots (e.g., swarm robotics), or a network of mobile phones.
-Examples of such large scale system can be a network of robots (e.g., swarm robotics), camera and IoT networks, or a network of mobile phones.
+Examples of such large-scale system can be a network of robots (e.g., swarm robotics), camera and IoT networks, or a network of mobile phones.
 %

### 2023-05-11 paper.tex
minor on intro

@@
 \begin{IEEEkeywords}
-aggregate computing, Graph Neural Network, Cyber-Physical Swarms.
+aggregate computing, Graph Neural Networks, and Cyber-Physical Swarms.
 \end{IEEEkeywords}
@@
 %
-This follow a novel trends in which high-level declarative programming language are mixed with machine learning technique in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
+This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
 %Specifically, we will show... \lukas{more info...}

### 2023-05-11 paper.tex
chore: add lukas sentence on GNN

@@
   and then combining it with the node's current embedding $h_v$.
-  \lukas{maybe add:``The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased.''}
+ The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased
+%\lukas{maybe add:``The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased.''}
 % 

### 2023-05-11 paper.tex
more comments

@@
 We demonstrate the effectiveness of this new approach in a simulated use case 
- where coordination tasks are successfully solved. %\lukas{do we really have multiple case studies?} 
+ where coordination tasks are successfully solved. 
 \end{abstract}
@@
 This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
-%Specifically, we will show... \lukas{more info...}
 Specifically, 
@@
 \subsection{Swarm systems}
-This article aims in defining intelligent \emph{collective behaviours} within the context of \emph{large-scale} distributed systems.
+This article studies intelligent \emph{collective behaviours} within the context of \emph{large-scale} distributed systems.
  Specifically, we focus on \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}, i.e.,  systems consisting of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator} able to interact with the real world.
  Each of these nodes can interact with its direct neighbours, either based on physical (e.g. communication range) or logical distance (e.g. 1-hop neighbourhood).
- %\lukas{does the article focus on the definition of decentralised collective intelligence?}
- % \lukas{distributed or decentralised?}
 
@@
 %
-%\lukas{not really a system definition - maybe we rename this to \emph{Swarm Systems}?}
 \subsection{Field-based Coordination}
@@
 Each node $v \in V$ is associated with an observation (or feature set) $o_v$.
+\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
 %
@@
  The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased
-%\lukas{maybe add:``The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased.''}
-% 
 Formally, a \ac{GNN} can be defined by two phases:
@@
 This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
+\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}
 
@@
  where collective intelligence emerges from local and repeated interaction of simple entities, like in swarm robotics.
-%\lukas{i am not sure i understand the previous sentence.}
 
 In fact, the first works in this direction were discussed precisely swarm robotics, %\lukas{what does this mean `in the last field'?}
- exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating 
- a policy representing 
- %\lukas{maybe `valid to' or `representing' instead of `equal to'?} 
- the entire system. 
+ exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating a policy representing  the entire system.
+\lukas{This sentence still needs rework.}
 %Modern approaches, however, 
@@
 Some known approaches using mean-field reinforcement learning include Q-mean, 
- which is an extension of Q-learning to mean-field settings~\cite{}, 
- and actor-critic mean-field~\cite{}, which combines actor-critic algorithms with mean-field approximations. 
+ which is an extension of Q-learning to mean-field settings~\cite{yang2018mean}, 
+ and actor-critic mean-field~\cite{frikha2023actor}, which combines actor-critic algorithms with mean-field approximations. 
 %
@@
 \begin{figure*}
-  \includegraphics[width=\textwidth]{imgs/architecture.pdf}
+  \includegraphics[width=\linewidth]{imgs/architecture.pdf}
   \label{fig:architecture}
@@
   Finally, utilizing the GNN, actions are computed for each node in the system to be performed against the environment, enabling advancement in the simulations according to swarMDP rules.
+  \lukas{should we have numbers at each step in the figure (and then referring in the caption and text)?}
+  \lukas{also, i think we can have the AC and GNN box a bit smaller but increase the size of the networks slightly - and put it on column width ot save space}
 }
@@
 \emph{i)} the aggregate program used to create part of the observation and 
-\emph{ii)} the policy $\pi_{gnn}$ learned through \ac{GNN}-based learning. 
+\emph{ii)} the policy $\pi_{gnn}$ learned through \ac{GNN}-based approach. 
 %
-Let $\Gamma$ be the aggregate program that runs against $G_t$. 
+Let $\Gamma$ be the aggregate program that runs against $G_t$.  \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
 %

### 2023-05-11 paper.tex
chore: comment for @LukasEsterle

@@
 %
-Let $\Gamma$ be the aggregate program that runs against $G_t$.  \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
+Let $\Gamma$ be the aggregate program that runs take $G_t$ as input.  \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
+\ga{In the problem formalization, I defined $G_t$, but it may be more appropriate to define it directly in the GNN, as suggested.
+I made this choice since GNNs typically only receive a graph as input, without regard for time, I included the definition of $G_t$ in the problem formalization as it pertains to the dynamics of the overall system.
+Specifically, when discussing the system's dynamics at a given time t, it's necessary to define the corresponding graph.
+I hope this clarifies my reasoning.
+Let me know!}
 %
-The evaluation of $\Gamma$  produce a field value $\theta^v_t$ for each node $v$ in $G_t$. 
+The evaluation of $\Gamma$ produces a field value $\theta^v_t$ for each node $v$ in $G_t$. 
 The feature vector $o_v$ of which is composed the graph $G'_t$ used by the policy $\pi_{gnn}$ is described by:

### 2023-05-11 paper.tex
chore: change the picture following lukas' suggestionsù

@@
   \caption{High-level description of Field-Informed reinforcement learning dynamics. 
-  For each time step, a graph is constructed from the environment, and it will associate each node with a local feature (hexagons in the picture). 
+  For each time step $t$, a graph is constructed from the environment, and it will associate each node with a local feature (hexagons in the picture). 
   Using this feature, an aggregate program computes spatio-temporal information that enhances the feature set of each node, depicted as colours in the middle graph. 
   Finally, utilizing the GNN, actions are computed for each node in the system to be performed against the environment, enabling advancement in the simulations according to swarMDP rules.
-  \lukas{should we have numbers at each step in the figure (and then referring in the caption and text)?}
+  %\lukas{should we have numbers at each step in the figure (and then referring in the caption and text)?}
   \lukas{also, i think we can have the AC and GNN box a bit smaller but increase the size of the networks slightly - and put it on column width ot save space}

### 2023-05-11 paper.tex
chore: reduce the image size for the high level description of field-infor rl
typo

@@
 but it could be seen as a global evolution, therefore we can use global information to inform the local controller.}
-\begin{figure*}
+\begin{figure}
   \includegraphics[width=\linewidth]{imgs/architecture.pdf}
@@
 }
-\end{figure*}
-In this section, we aim to clarify the components involved in our proposed solution (i.e., architecture), how these components interact with each other to bring the system to perform the collective behavior (i.e dynamics), and finally, we will detail the learning algorithm designed and used to synthesise the policy (learning algorithm).
+\end{figure}
+In this section, we aim to clarify the components involved in our proposed solution (i.e., architecture), how these components interact with each other to bring the system to perform the collective behaviour (i.e dynamics), and finally, we will detail the learning algorithm designed and used to synthesise the policy (learning algorithm).
 \subsection{Architecture and System dynamics}

### 2023-05-11 paper.tex
prefer to use regret instead of reward in the discussion
add a brief description of the system model and the dynamics of the node

@@
 % 
-Within aggregate computing, the  \emph{macro-programming}~\cite{DBLP:journals/corr/abs-2201-03473} paradigm can be realised to devise collective and self-organizing behaviour through a composition of functions operating on fields.
+This \emph{macro-programming}~\cite{DBLP:journals/corr/abs-2201-03473} paradigm can be realised to devise collective and self-organizing behaviour through a composition of functions operating on fields.
 %
@@
 Fields are computed locally but subject to a global viewpoint, 
- enabling emergent collective behaviour through the interplay of the system model (similar to the above described) 
- and the \emph{programming model} known as field calculus~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}.
+ enabling emergent collective behaviour through the interplay of the \emph{system model} (i.e., structure and dynamics of the systems)
+ and the \emph{programming model} (i.e., how collective behaviours are expressed) known as field calculus~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}.
+%
+The foster is structurally similar to \emph{swarm-like systems}, 
+ as it is composed of a set of \emph{agents} 
+ that interact with each other through a \emph{network} of \emph{neighbourhood relations}.
+%
+The dynamics of each agent follow a local control loop called \emph{round}
+ that comprises the following steps:
+\begin{enumerate}
+  \item \emph{Sense}: the agent acquires information from the environment and collects the messages of its neighbourhood, building a \emph{context};
+  \item \emph{Compute}: the agent executes a \emph{program} that computes an \emph{export}, that is a set of values to be sent to the neighbourhood;
+  \item \emph{Act}: the agent sends the export to its neighbourhood and updates its internal state acting on the environment.
+\end{enumerate}
+The proactive and iterative execution of the round by each agent leads to the emergence of collective behaviour specified by the program.
+
+The programming model is ruled by the \emph{field calculus}~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}, 
+ a core language that allows the expression of collective behaviours through the composition of \emph{functions} operating on fields.
+
 \lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
@@
 %
-Therefore, we relied on the value-based approach \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications: 
+Therefore, we relied on the value-based approach \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications (see \Cref{alg:dqn}): 
 \begin{enumerate}
@@
 \end{enumerate}
-The first point is a natural extension due to the fact that we work on graphs rather than simple values. 
+The first point is a natural extension because we work on graphs rather than simple values. 
 %
@@
 %
-\ga{todo: describe the algorithm with the modifications}
 \begin{algorithm}
@@
   \caption{Deep Q-Network (DQN) with GNN and Graph Replay Buffer}
+  \label{alg:dqn}
   \end{algorithm}
@@
 \begin{equation*}
-R_{v} = R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}
+R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
 \end{equation*}
+Specifically, we decided to express the signal as a \emph{regret}
+  as it is a more general measure of the quality of the action taken by the agent.
+It is simplify evaluated as:
+\begin{equation*}
+R^e_{v} = 1 - R_v
+\end{equation*}
+%
 

### 2023-05-12 scala.tex
add other files

\lstdefinelanguage{scala}{
  keywords={abstract,case,catch,class,def,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,lazy,%
    type,val,var,while,with,yield,forSome},
  otherkeywords={=>,<-,<\%,<:,>:,\#},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]""",
  basicstyle=\lst@ifdisplaystyle\small\fi\ttfamily,
  emphstyle=\bfseries
}
\definecolor{ddarkgreen}{rgb}{0,0.5,0}
\lstdefinelanguage{scafi}{frame=single,basewidth=0.5em,language={scala},
keywordstyle=\color{blue}\textbf, commentstyle=\color{ddarkgreen},
keywordstyle=[2]\color{red}\textbf, keywords=[2]{rep,nbr,foldhood,foldhoodPlus,aggregate,branch,spawn,mux,mid},
keywordstyle=[3]\color{gray}, keywords=[3]{Me,AroundMe,Everywhere,Forever}, %,@@,@@@
keywordstyle=[4]\color{red}\textbf, keywords=[4]{in,out,rd},
keywordstyle=[5]\color{violet}, keywords=[5]{evolve,when,andNext,workflow,G,C,broadcast,gradient,gossip},
keywordstyle=[6]\color{orange}, keywords=[6]{Available,Serving,Done,Waiting,Removing,None,Set}}

### 2023-05-12 paper.tex
chore: add root in scala def

@@
 %
-
+\input{scala.tex}
+\lstset{language={scafi}}
 %%% Comment command, to be removed before submission
@@
 \ga{Page budget: 1/2 page \\}
-\ga{Plan: discussion about the problem of coordination in multi-agent systems 
-and the need for a scalable approach. In doing this, we will discuss some of the existing approaches (declarative, RL and )}
+%\ga{Plan: discussion about the problem of coordination in multi-agent systems and the need for a scalable approach.
+In doing this, we will discuss some of the existing approaches (declarative, RL and )}
 \subsection{Swarm systems}
@@
 \subsection{Field-based Coordination}
-\ga{discussion about field used as a coordination mechanism in multi-agent systems. Introduction to AC}
+%\ga{discussion about field used as a coordination mechanism in multi-agent systems. Introduction to AC}
 The field-based coordination approaches utilise a concept of \emph{computational fields} 
@@
  a core language that allows the expression of collective behaviours through the composition of \emph{functions} operating on fields.
+%
+Field calculus includes main operators for expressing \emph{spatio-temporal} compotations that allow for 
+\emph{i)} the progression of values over time, achieved by transforming a value computed in a previous round into a new value; 
+\emph{ii)} the exchange of data with neighbouring fields, where received data is validated by adjacent fields; and 
+\emph{iii)} the conditional division of computation into distinct domains of collective computation. 
+%
+We avoid presenting here the full details of these operators since they are not essential to the contribution. 
+For further elaboration on the actual field calculus, please refer to the work of~\cite{DBLP:journals/jlap/ViroliBDACP19}.
 
+On these minimal operators, 
+ it is then possible to build \emph{self-organizing} coordination blocks.
+ One of the founding blocks is the \emph{gradient} essential for 
+ \emph{information} flows~\cite{DBLP:conf/saso/WolfH07}.
+%
+This operator generates a numerical field that represents the minimum distance from a source zone. 
+%
+In other words, it maps a Boolean field (where ``true'' indicates the presence of a source node and ``false'' indicates its absence) to a distance field that indicates the proximity to the nearest source.
+In ScaFi (a scala implementation of aggregate computing) it is simply expressed as:
+\begin{lstlisting}
+def gradient(source: Boolean): Double
+\end{lstlisting}
+Along the gradient, it is possible to cast information within the system, accumulating value during the programming. This pattern is called \emph{gradient-cast} (or \emph{G}) 
+\begin{lstlisting}
+def G[V](source: Boolean, value: V, acc: V => V): V
+\end{lstlisting}
+\ga{more description here}
+this is used in the use case to accumulate the distance from the phenomena that is tracked
 \lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
@@
 \end{equation}
+
 \begin{equation}
@@
 \subsection{Many-Agent Reinforcement Learning}
-\ga{plan: we could discuss briefly about standard approach (few agents), and mean-field approach (many agents). 
-This will lead to our approach, which is a combination of DQN (or any Deep RL approach) and GNN.}
+%\ga{plan: we could discuss briefly about standard approach (few agents), and mean-field approach (many agents). 
+%This will lead to our approach, which is a combination of DQN (or any Deep RL approach) and GNN.}
 
@@
 \ga{Page budget: 2/3 pages \\}
-\ga{Plan: we should discuss the approach in detail, including the system model, and the learning dynamics.
-Particularly, I would like to highlight how each node in the graph is a local controller,
-but it could be seen as a global evolution, therefore we can use global information to inform the local controller.}
+%\ga{Plan: we should discuss the approach in detail, including the system model, and the learning dynamics.
+Particularly, I would like to highlight how each node in the graph is a local controller, but it could be seen as a global evolution, therefore we can use global information to inform the local controller.}
 \begin{figure}
@@
 \subsection{Architecture and System dynamics}
-\ga{Plan: 
-  discussion about the typical AC evaluation model, specifying how GNN (with 1-hop information diffusion)
-  can be easily integrated with this distributed model
-}
+%\ga{Plan: 
+%  discussion about the typical AC evaluation model, specifying how GNN (with 1-hop information diffusion)
+%  can be easily integrated with this distributed model
+%}
 The proposed solution consists mainly of two parts, 
@@
 \subsection{Learning algorithm}
-\ga{
-  Plan: discuss the chosen approach, that is Centralised Training and Distributed Execution.
-  Particularly, we use DQN in which the neural network is a GNN. So, the dataset consists in a set of graphs,
-  where each graph is a snapshot of the system. 
-  The target is the Q-value of the action taken in the current state.
-  Explain how this can be then used for distributed controllers.
-}
+%\ga{
+%  Plan: discuss the chosen approach, that is Centralised Training and Distributed Execution.
+%  Particularly, we use DQN in which the neural network is a GNN. So, the dataset consists in a set of graphs,
+%  where each graph is a snapshot of the system. 
+%  The target is the Q-value of the action taken in the current state.
+%  Explain how this can be then used for distributed controllers.
+%}
 In swarm-like systems, 
@@
 \ga{Page budget: 2 pages \\}
-\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
-I dunno if it makes sense to discuss the variants (i.e., fixed area, two areas and moving area). We will see.
-}
+%\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
+%I dunno if it makes sense to discuss the variants (i.e., fixed area, two areas and moving area). We will see.}
 To test the effectiveness of the proposed approach, 

### 2023-05-12 paper.tex
chore: update affiliation

@@
 \textit{%Alma Mater Studiorum--
-University of Bologna}\\
+Alma Mater Studiorum — Università di Bologna}\\
 Cesena, Italy\\
@@
 \begin{abstract}
-Coordinatinating a group of intelligent agents in multi-agent systems 
- is a research problem that has been addressed for a long time, 
- due to the challenges posed by distribution and the definition of distributed intelligence. 
+Coordinating a multi-agent system of intelligent situated agents is a traditional research problem, 
+impacted by the challenges posed by the very notion of distributed intelligence. 
 %
-The problem is even more evident in collective adaptive systems, 
- where the scale of the systems considered makes the definition of collective behaviours even more challenging. 
+The problem is even more evident in large-scale collective adaptive systems, 
+ where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder.
 % 
-Thus far, two main macro approaches have been proposed for creating distributed controllers: 
+In this context, two main approaches have been proposed for creating distributed controllers out of macro-level task/goal descriptions: 
  \emph{manual design}, in which programmers build the controllers directly, and 
@@
 In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning}, 
-  where we use computational \emph{computational fields} to manage the interaction between agents 
-  and \emph{Graph Neural Networks} to learn a local behaviour necessary to solve collective tasks. 
+  which uses manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents 
+  and \emph{Graph Neural Networks} to learn the local behaviour necessary to solve collective tasks. 
 %
 This allows us to create distributed controllers informed by a collective knowledge 
- distilled during learning but that use only local information at runtime.
+ that has been distilled during learning, but that uses only local information at runtime.
 % 
-We demonstrate the effectiveness of this new approach in a simulated use case 
- where coordination tasks are successfully solved. 
+We demonstrate the effectiveness of this new approach in simulated use cases 
+ where tracking and covering tasks for swarm robotics are successfully solved. 
 \end{abstract}
@@
 
-\Ac{RL} has gained a lot of interest recently, 
+ \Ac{RL} learning has gained a lot of interest recently, 
  thanks to its successful application in various scenarios, 

### 2023-05-12 paper.tex
abstract

--- 2023-05-12 paper.tex
@@
 \begin{abstract}
-Coordinatinating a group of intelligent agents in multi-agent systems 
- is a research problem that has been addressed for a long time, 
- due to the challenges posed by distribution and the definition of distributed intelligence. 
+Coordinating a group of autonomous intelligent agents in multi-agent systems to achieve a common, global goal requires them to acquire information locally, share their knowledge, and act accordingly on their environment.
+This shared knowledge leads to distributed intelligence, enabling the collective of agents to operate together.
+With the rising number of such autonomous agents within these systems, sharing knowledge becomes a dedicated challenge.
+Here, the individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
+
+% is a research problem that has been addressed for a long time, 
+% due to the challenges posed by distribution and the definition of distributed intelligence. 
 %
-The problem is even more evident in collective adaptive systems, 
- where the scale of the systems considered makes the definition of collective behaviours even more challenging. 
+%The problem is even more evident in collective adaptive systems, 
+% where the scale of the systems considered makes the definition of collective behaviours even more challenging. 
 % 
-Thus far, two main macro approaches have been proposed for creating distributed controllers: 
+Thus far, two main approaches have been proposed for creating such distributed controllers: 
  \emph{manual design}, in which programmers build the controllers directly, and 
@@
   and \emph{Graph Neural Networks} to learn a local behaviour necessary to solve collective tasks.
+   \lukas{are you considering computational fields as `manual design'} 
 %
@@
 \end{equation}
-
 \begin{equation}

### 2023-05-12 paper.tex
minor

--- 2023-05-12 paper.tex
@@
 \begin{abstract}
+Coordinating a group of autonomous intelligent agents in multi-agent systems to achieve a common, global goal requires them to acquire information locally, share their knowledge, and act accordingly on their environment.
+%
+From this shared knowledge distributed intelligence emerges, enabling the collective of agents to operate together.
+With the rising number of such autonomous agents within these systems, 
+ sharing knowledge becomes a dedicated challenge. 
+ Here, the individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
+% is a research problem that has been addressed for a long time, 
+% due to the challenges posed by distribution and the definition of distributed intelligence. 
+%
+%The problem is even more evident in collective adaptive systems, 
+% where the scale of the systems considered makes the definition of collective behaviours even more challenging.
+
 Coordinating a multi-agent system of intelligent situated agents is a traditional research problem, 
-impacted by the challenges posed by the very notion of distributed intelligence. 
+impacted by the challenges posed by the very notion of distributed intelligence.  \lukas{what are these challenges?}
 %
 The problem is even more evident in large-scale collective adaptive systems, 
- where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder.
+where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder. \lukas{what does it make harder?}
 % 
@@
 
- \Ac{RL} learning has gained a lot of interest recently, 
+ \Ac{RL} has gained a lot of interest recently, 
  thanks to its successful application in various scenarios, 

### 2023-05-12 paper.tex
removed comments

@@
 In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning}, 
-  which uses manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents 
-  and \emph{Graph Neural Networks} to learn the local behaviour necessary to solve collective tasks. 
+  which uses manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents and a \emph{reinforcement learning technique} in combination with \emph{Graph Neural Networks} to learn the local behaviour necessary to solve collective tasks.
 %
@@
 Each node $v \in V$ is associated with an observation (or feature set) $o_v$.
-\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
+%\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
 %

### 2023-05-12 paper.tex
rewrote one sentence in introduction

@@
 %
-In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning}, 
-  which uses manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents and a \emph{reinforcement learning technique} in combination with \emph{Graph Neural Networks} to learn the local behaviour necessary to solve collective tasks.
+In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning}.
+We utilise manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents.
+Using \emph{Deep Q-learning} in combination with \emph{Graph Neural Networks} we enable the agents to learn the necessary local behaviour automatically to solve collective tasks.
 %
@@
 %
-Let $\Gamma$ be the aggregate program that runs take $G_t$ as input.  \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
+Let $\Gamma$ be the aggregate program that takes $G_t$ as input. 
+% \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
 \ga{In the problem formalization, I defined $G_t$, but it may be more appropriate to define it directly in the GNN, as suggested.

### 2023-05-12 paper.tex
integrate lukas and mirko introduction

@@
 \begin{abstract}
-Coordinating a group of autonomous intelligent agents in multi-agent systems to achieve a common, global goal requires them to acquire information locally, share their knowledge, and act accordingly on their environment.
+%Coordinating a group of autonomous intelligent agents in multi-agent systems to achieve a common, global goal requires them to acquire information locally, share their knowledge, and act accordingly on their environment.
 %
-From this shared knowledge distributed intelligence emerges, enabling the collective of agents to operate together.
-With the rising number of such autonomous agents within these systems, 
- sharing knowledge becomes a dedicated challenge. 
- Here, the individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
+%From this shared knowledge distributed intelligence emerges, enabling the collective of agents to operate together.
+%With the rising number of such autonomous agents within these systems, 
+% sharing knowledge becomes a dedicated challenge. 
+% Here, the individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
 % is a research problem that has been addressed for a long time, 
@@
 % where the scale of the systems considered makes the definition of collective behaviours even more challenging. 
-
+%
 Coordinating a multi-agent system of intelligent situated agents is a traditional research problem, 
-impacted by the challenges posed by the very notion of distributed intelligence.  \lukas{what are these challenges?}
+impacted by the challenges posed by the very notion of distributed intelligence,
+ that concern in achieving a common, global goal of acquiring information locally, sharing their knowledge, and acting accordingly in their environment. 
 %
 The problem is even more evident in large-scale collective adaptive systems, 
-where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder. \lukas{what does it make harder?}
+where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder, since the individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
+%\lukas{what does it make harder?}
 % 

### 2023-05-12 paper.tex
chore: remove lukas comment to fig.1

@@
 %\lukas{what does it make harder?}
-% 
+
 In this context, two main approaches have been proposed for creating distributed controllers out of macro-level task/goal descriptions: 
@@
 % 
-In \ac{RL}, an agent (i.e., a smart entity capable of making decisions) 
- performs actions in an environment (i.e., everything outside the agent) according to a policy, 
- to maximise long-term rewards (or minimise a regret).
+In \ac{RL}, an \emph{agent} (i.e., a smart entity capable of making decisions) 
+ performs \emph{actions} in an \emph{environment} (i.e., everything outside the agent) according to a \emph{policy}, 
+ to maximise long-term \emph{reward signal} (or minimise a \emph{regret}).
 
@@
   %\lukas{should we have numbers at each step in the figure (and then referring in the caption and text)?}
-  \lukas{also, i think we can have the AC and GNN box a bit smaller but increase the size of the networks slightly - and put it on column width ot save space}
+  %\lukas{also, i think we can have the AC and GNN box a bit smaller but increase the size of the networks slightly - and put it on column width ot save space}
 }
@@
 % \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
-\ga{In the problem formalization, I defined $G_t$, but it may be more appropriate to define it directly in the GNN, as suggested.
+%\ga{In the problem formalization, I defined $G_t$, but it may be more appropriate to define it directly in the GNN, as suggested.
 I made this choice since GNNs typically only receive a graph as input, without regard for time, I included the definition of $G_t$ in the problem formalization as it pertains to the dynamics of the overall system.

### 2023-05-12 paper.tex
chore: leave mean field but move CTDE in background

@@
  where collective intelligence emerges from local and repeated interaction of simple entities, like in swarm robotics.
-
-In fact, the first works in this direction were discussed precisely swarm robotics, %\lukas{what does this mean `in the last field'?}
- exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating a policy representing the entire system.
+%
+In such many-agent scenarios, 
+ the implementation of fully decentralized learning is often unfeasible due to the large number of learning agents, 
+ which renders the system non-stationary and difficult to manage. 
+%
+Conversely, a centralized controller capable of coordinating 
+ the entire system may not be a viable solution due to scalability concerns. 
+ To address this challenge, a practical solution is the adoption of \ac{CTDE}.
+%
+The idea is to learn a policy at simulation time where there is a collective view of the system, 
+ and then at runtime use the policy found with global information. 
+This approach allows policies that are influenced by global information but only require local information to function at runtime. 
+%
+The typical approach in such cases is based on actor-critic systems~\cite{DBLP:conf/nips/LoweWTHAM17,wu2022more,song2022ctds,song2022centralized},
+  where the \emph{actor} is the distributed policy (with only local information) and the critic is a neural network that takes the overall system state.
+%
+%In fact, the first works in this direction were discussed precisely swarm robotics, %\lukas{what does this mean `in the last field'?}
+% exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating a policy representing the entire system.
 \lukas{This sentence still needs rework.}
@@
 % have started to consider the use of 
- Recently, deep learning approaches have been considered 
- to synthesise robust controllers capable of generalizing to new tasks. 
+% Recently, deep learning approaches have been considered 
+% to synthesise robust controllers capable of generalizing to new tasks. 
 %In this context, mean-field reinforcement learning~\cite{pmlr-v80-yang18d} is certainly noteworthy. 
- Mean-field RL~\cite{pmlr-v80-yang18d}  is a technique for learning approximate inference in large-scale systems.
-Here, the interactions between the agents can be decoupled into a single conditional probability distribution function for each agent,
- thus simplifying the learning process and reducing the computational cost.
+Mean-field RL~\cite{pmlr-v80-yang18d} is one of such concrete applications of \ac{CTDE} 
+ where the interactions among the population of agents are estimated by considering either the effect of a single agent and the average impact of the entire population or the influence of neighbouring agents.
 %
@@
 %}
-In swarm-like systems, 
- the typical approach for learning follows a strategy known as \ac{CTDE}. \ga{perhaps it should be introduced in many RL} 
-The idea is to learn a policy at simulation time where there is a collective view of the system, 
- and then at runtime use the policy found with global information. 
-This approach allows policies that are influenced by global information but only require local information to function at runtime. 
+Since we consider swarm-like systems, that are an example of many-agent systems, 
+ the proposed approach follows a \ac{CDTE} learning pattern,
+ but differently from mean-field approaches and actor-critic solutions,
+ we use a value-based approach combined with a \ac{GNN} as a function approximator.
+%In swarm-like systems, 
+% the typical approach for learning follows a strategy known as \ac{CTDE}. \ga{perhaps it should be introduced in many RL} 
+%The idea is to learn a policy at simulation time where there is a collective view of the system, 
+% and then at runtime use the policy found with global information. 
+%This approach allows policies that are influenced by global information but only require local information to function at runtime. 
 %
-The typical approach in such cases is based on actor-critic systems~\cite{DBLP:conf/nips/LoweWTHAM17,wu2022more,song2022ctds,song2022centralized},
-  where the \emph{actor} is the distributed policy (with only local information) and the critic is a neural network that takes the overall system state.
-
-In our case, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighbourhood. 
+%The typical approach in such cases is based on actor-critic systems~\cite{DBLP:conf/nips/LoweWTHAM17,wu2022more,song2022ctds,song2022centralized},
+%  where the \emph{actor} is the distributed policy (with only local information) and the critic is a neural network that takes the overall system state.
 %
-Therefore, we relied on the value-based approach \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications (see \Cref{alg:dqn}): 
+Specifically, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighbourhood. 
+%
+As value-based algorithm we relied on \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications (see \Cref{alg:dqn}): 
 \begin{enumerate}

### 2023-05-12 paper.tex
another try

--- 2023-05-12 paper.tex
@@
 Coordinating a multi-agent system of intelligent situated agents is a traditional research problem, 
-impacted by the challenges posed by the very notion of distributed intelligence,
- that concern in achieving a common, global goal of acquiring information locally, sharing their knowledge, and acting accordingly in their environment. 
+impacted by the challenges posed by the very notion of distributed intelligence.
+These problems arise from agents acquiring information locally, sharing their knowledge, and acting accordingly in their environment in order to achieve a common, global goal.
 %
-The problem is even more evident in large-scale collective adaptive systems, 
-where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder, since the individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
-%\lukas{what does it make harder?}
+Individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
+These problems are even more evident in large-scale collective adaptive systems, where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder.
 

### 2023-05-15 paper.tex
chore: fix minors
remove first comment on paper levels
chore: add plots for reward and coerage in time
some comments

@@
 \usepackage{amsmath}
+\usepackage{subcaption}
 \usepackage[ruled]{algorithm2e}
@@
 %
-\ga{Paper levels:
-\begin{itemize}
-  \item Distributed collective intelligence
-  \item GNN and field coordination
-  \item Spatial tracking
-\end{itemize}
-}
+%\ga{Paper levels:
+%\begin{itemize}
+%  \item Distributed collective intelligence
+%  \item GNN and field coordination
+%  \item Spatial tracking
+%\end{itemize}
+%}
 \section{Introduction}
@@
 \end{figure}
-In this section, we aim to clarify the components involved in our proposed solution (i.e., architecture), how these components interact with each other to bring the system to perform the collective behaviour (i.e dynamics), and finally, we will detail the learning algorithm designed and used to synthesise the policy (learning algorithm).
+In this section, we discuss the components involved in our proposed solution (i.e., architecture) and how these components interact with each other to bring the system to perform the collective behaviour (i.e dynamics).
+Finally, we will detail the learning algorithm designed and used to synthesise the policy (i.e., learning algorithm).
 \subsection{Architecture and System dynamics}
@@
 %
-Let $\Gamma$ be the aggregate program that takes $G_t$ as input. 
+Let $\Gamma$ be the aggregate program that takes a graph $G_t$, representing the participating nodes and their neighbourhood relations at time $t$, as input. 
 % \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
@@
 The feature vector $o_v$ of which is composed the graph $G'_t$ used by the policy $\pi_{gnn}$ is described by:
- $o_v = (\theta^v_t, \omega)$.
+ $o_v = (\theta^v_t, \omega)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
  where $\omega$ are other local information gather from sensors (like temperature, distances from neighbourhood, etc.). 
@@
 %}
-Since we consider swarm-like systems, that are an example of many-agent systems, 
- the proposed approach follows a \ac{CDTE} learning pattern,
+Since we consider swarm-like systems, which are an example of many-agent systems, 
+ the proposed approach follows a \ac{CTDE} learning pattern,
  but differently from mean-field approaches and actor-critic solutions,
@@
 \begin{equation*}
-R^e_{v} = 1 - R_v
+R^e_{v} = R_v - 1
 \end{equation*}
@@
 \subsection{Discussion and Results}
+%% Subfigures of coverage-in-time and reward-in-time
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
+    \caption{Coverage in time}
+    \label{fig:coverage}
+  \end{subfigure}
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\linewidth]{imgs/reward-in-time}
+    \caption{Reward in time}
+    \label{fig:reward}
+  \end{subfigure}
+  \caption{Coverage and reward in time}
+  \label{fig:coverage-reward}
+\end{figure}
 \section{Conclusion and Future Work}

### 2023-05-15 paper.tex
some comments
minors
minor

@@
 Coordinating and organising this information exchange within the network of agents is required for improve the collaboration among them further~\cite{jennings1996coordination}.
-Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complimentary competences of the individual agents, or (iii) the collective resources of multiple agents.
+Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complementary competences of the individual agents, or (iii) the collective resources of multiple agents.
 
@@
  $o_v = (\theta^v_t, \omega)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
+ \lukas{we have not introduced $G'_t$ yet}
  where $\omega$ are other local information gather from sensors (like temperature, distances from neighbourhood, etc.). 
@@
 
-  Compute the current expetected value for each node in the graph:
+  Compute the current expected value for each node in the graph:
   $y^* = Q(s'i,a';\theta)$;

### 2023-05-15 paper.tex
try to answer to lukas questions

@@
 The node embedding $h_v$ describes the node in the network and summarises the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
- In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $N(v)$,
+ In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $\mathcal{N}_G(v)$,
   and then combining it with the node's current embedding $h_v$.
@@
 \begin{equation}
-a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}(v)\right\}\right)   
+a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}_G(v)\right\}\right)   
 \end{equation}
@@
 \end{equation}
-where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer ($h_v^0 = o_v$), 
- $\mathcal{N}(v)$ is the set of neighbours of node $v$ computed from $E$.
+
+where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
+ $\mathcal{N}_G(v)$ is the set of neighbours of node $v$ computed from $E$.
+%
+$h_v{v}^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
 %
@@
   where $E_t$ is built from $\mathcal{N}$.
-  Each node is then decorated with the local observation $o_v$. 
+  Each node is then decorated with the local observation $f_v = o_v$. 
   Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022}.
@@
 %
-Let $\Gamma$ be the aggregate program that takes a graph $G_t$, representing the participating nodes and their neighbourhood relations at time $t$, as input. 
+Let $\Gamma$ be the aggregate program that takes a graph $G_t$ decorated by $o_v$ (i.e., the local observation)), 
+ representing the participating nodes and their neighbourhood relations at time $t$, as input. 
 % \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
@@
 The evaluation of $\Gamma$ produces a field value $\theta^v_t$ for each node $v$ in $G_t$. 
-The feature vector $o_v$ of which is composed the graph $G'_t$ used by the policy $\pi_{gnn}$ is described by:
- $o_v = (\theta^v_t, \omega)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
+From this field, we construct the feature vector $f_v$ for each node $v$ in $G_t$ as follows:
+ $f_v = (\theta^v_t, o_t)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
  \lukas{we have not introduced $G'_t$ yet}
- where $\omega$ are other local information gather from sensors (like temperature, distances from neighbourhood, etc.). 
+ %where $\omega$ are other local information gather from sensors (like temperature, distances from neighbourhood, etc.). 
 %

### 2023-05-15 paper.tex
try to answer to lukas questions

@@
 % 
-Each node $v \in V$ is associated with an observation (or feature set) $o_v$.
+Each node $v \in V$ is associated with an observation (or feature set) $f_v$.
 %\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
 %
-Given $o_v$, %a feature vector associated with each node $v \in V$, 
+Given $f_v$, %a feature vector associated with each node $v \in V$, 
  the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$.
@@
   where $E_t$ is built from $\mathcal{N}$.
-  Each node is then decorated with the local observation $f_v = o_v$. 
-  Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022}.
+  Each node is then decorated with the local observation perceived at the time $t$: $o^v_t \in O$. 
+  Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022} if we set $f_v = o^v_t$.
 \subsection{Motivation}
@@
 %
-Let $\Gamma$ be the aggregate program that takes a graph $G_t$ decorated by $o_v$ (i.e., the local observation)), 
+Let $\Gamma$ be the aggregate program that takes a graph $G_t$ decorated by $o^t_v$, 
  representing the participating nodes and their neighbourhood relations at time $t$, as input. 
@@
 The evaluation of $\Gamma$ produces a field value $\theta^v_t$ for each node $v$ in $G_t$. 
-From this field, we construct the feature vector $f_v$ for each node $v$ in $G_t$ as follows:
- $f_v = (\theta^v_t, o_t)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
+The feature vector $o_v$ of which is composed the graph $G'_t$ used by the policy $\pi_{gnn}$ is described by:
+ $o_v = (\theta^v_t, \omega)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
  \lukas{we have not introduced $G'_t$ yet}

### 2023-05-16 paper.tex
chore: put simulations image
chore: improve approach description

@@
 The evaluation of $\Gamma$ produces a field value $\theta^v_t$ for each node $v$ in $G_t$. 
-The feature vector $o_v$ of which is composed the graph $G'_t$ used by the policy $\pi_{gnn}$ is described by:
- $o_v = (\theta^v_t, \omega)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
+From this field, we construct the feature vector $f_v$ for each node $v$ in $G_t$ as follows:
+ $f_v = (\theta^v_t, o^t_v)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
  \lukas{we have not introduced $G'_t$ yet}
@@
 %
-The policy $\pi_{gnn}$ is then evaluated for each node, 
+The policy $\pi_{gnn}$ is then evaluated for each node using $f_v$ as input, 
  producing an action $a_t$ that will then modify the global state of the system.
@@
  but differently from mean-field approaches and actor-critic solutions,
- we use a value-based approach combined with a \ac{GNN} as a function approximator.
+ we use a \emph{value-based} approach combined with a \ac{GNN} as a function approximator.
 %In swarm-like systems, 
@@
 %
-Specifically, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighbourhood. 
+Specifically, we leveraged the property of \ac{GNN}s to have a dual interpretation, 
+ i.e., to function globally over the entire graph and locally only over the neighbourhood. 
 %
@@
 \begin{enumerate}
-  \item experience replay stores experiences in the form of \emph{graphs},
+  \item experience replay stores experiences in the form of \emph{graphs} decorated with features (e.g., observations, actions, rewards, etc.),
   \item the neural network used to compute the Q function is based on a \ac{GNN} with an \ac{MLP} downstream.
@@
 For the second point,
- the use of \acp{GNN} allows us to define policies on a variable neighbourhood  that is essential in such systems as this can vary based on the neighbourhood policy,
+ the use of \acp{GNN} allows us to define policies on a variable neighbourhood 
+ that is essential in such systems as this can vary based on the neighbourhood policy,
  and it is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes. 
@@
 \label{sec:eval}
+
+%% Put three subfigures of simulations, start, after, end
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/start.png}
+      \caption{Initial state}
+      \label{fig:initial}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/after.png}
+      \caption{After 50 steps}
+      \label{fig:after}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/end.png}
+      \caption{After 200 steps}
+      \label{fig:end}
+  \end{subfigure}
+  \caption{Simulation of the proposed approach. 
+  The dots represent the robots, 
+  the circle area on the top right represent the phenomenon to be monitored, 
+  and the gray area represent the area covered by the robots.}
+  \label{fig:simulations}
+\end{figure}
 \ga{Page budget: 2 pages \\}

### 2023-05-16 paper.tex
chore: more on notation

@@
 Each node $v \in V$ is associated with an observation (or feature set) $f_v$.
+For the sake of simplicity, 
+ we thereafter describe $G_f$ as a graph that contains the feature set $f_v$ for each node $v \in V$.
 %\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
@@
 \end{equation}
-
 where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
@@
 %
+Thereafter, we express the application of a \ac{GNN} to a graph $G_f$ as:
+\begin{equation}
+    GNN(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
+\end{equation} 
 This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.

### 2023-05-16 paper.tex
some comments

--- 2023-05-15 paper.tex
@@
 While manual deployment will mitigate the problem, not all phenomena are known exactly \emph{a priori} and would require maintenance and adjustment during runtime.
-Even worse, with phenomena able to change their location, size, and distribution---as it is the case with crowds, wild life, or wild fires---an adaptation of the collective sensors is required in order to keep the required information.
+Even worse, with phenomena able to change their location, size, and shape---as it is the case with crowds, wild life, or wild fires---an adaptation of the collective sensors is required in order to ensure all necessary information is acquired.
 
@@
 
-In this paper, we propose a novel approach combining aggregate computing~\cite{Beal2015Computer} with Graph Neural Networks~\cite{Zhou2020AIOpen} to focus on the relevant areas.
+In this paper, we propose a novel approach utilising aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
+\lukas{can we have a very short summary of how it works here.
+something along the lines of ``First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent.
+Second, we utilise a \ac{GNN} within the \ac{DQN}.
+Here the \ac{GNN} provides node embeddings for each agent, serving as input for the \ac{DQN}.
+The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.''}
 We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
@@
 Despite appearing centralised, this system proposes a completely decentralised execution. 
+\lukas{maybe: ``While the graph, containing the aggregate information, might appear as global knowledge, this is not the case as the information is never aggregated globally.
+The individual nodes only combine information from their local neighbourhood.''}
 %
@@
 Specifically, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighbourhood. 
+ \lukas{Global information is not available to the agent, right? The agent only provide locally aggregated information (resulting from the AC part)?}
 %
@@
 For the second point,
- the use of \acp{GNN} allows us to define policies on a variable neighbourhood  that is essential in such systems as this can vary based on the neighbourhood policy,
- and it is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes. 
+ the use of \acp{GNN} allows us to define policies on a variable neighbourhood, which is essential in such systems as this can change due to the applied neighbourhood policy.
+ It is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes~\cite{??}. 
 % 
@@
  During learning, it is as if we have a \emph{single} agent collectively choosing which action to take (i.e., resulting action graph).
+\lukas{this makes it sound like a singe agent has all information, selecting the actions for all other agents...}
 %
@@
   \While{not done}{
-  Observe current graph observations $G_o$;
+  Observe current graph observations $G_o$; \lukas{whats $G_o$?}
 
+%  $r \leftarrow $ randomNumber();
+%  
+%	\If{$r \leq \epsilon$}{select a random graph action $G_a$}
+%	\Else{select $G_a =$\\
+%		$\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;}
+%  
   With probability $\epsilon$ select a random graph action $G_a$, 
@@
   Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G^i_o)$ 
-   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$;
+   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$; \lukas{how does the merge work? do we explain this?}
 
@@
  specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
- or monitoring the water levels in a canal.
+ or monitoring the water levels in a canal. The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
+ \lukas{``Initially we consider a stationary phenomenon before expanding towards a moving phenomenon.
+Additionally, the phenomenon can change its shape during runtime.
+Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function.
+Figure~\ref{fig:simulations} illustrates a set of 25 agents and a single, stationary phenomenon with a with a 3D Gaussian distribution.''}
 

### 2023-05-16 paper.tex

@@
 %
-Specifically, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighbourhood. 
+Specifically, we leveraged the property of \ac{GNN}s to have a dual interpretation, 
+i.e., to function globally over the entire graph and locally only over the neighbourhood. 
  \lukas{Global information is not available to the agent, right? The agent only provide locally aggregated information (resulting from the AC part)?}
@@
  which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
+ \lukas{drone, robot, agent, node...}
 \subsection{Goal}

### 2023-05-16 paper.tex
minor changes
chore: fix time

--- 2023-05-16 paper.tex
@@
 %
-Let $\Gamma$ be the aggregate program that takes a graph $G_t$ decorated by $o^t_v$, 
+Let $\Gamma$ be the aggregate program that takes a graph $G_t$ decorated by $o^v_t$, 
  representing the participating nodes and their neighbourhood relations at time $t$, as input. 
@@
 From this field, we construct the feature vector $f_v$ for each node $v$ in $G_t$ as follows:
- $f_v = (\theta^v_t, o^t_v)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
+ $f_v = (\theta^v_t, o^v_t)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
  \lukas{we have not introduced $G'_t$ yet}
@@
 The policy $\pi_{gnn}$ is then evaluated for each node using $f_v$ as input, 
- producing an action $a_t$ that will then modify the global state of the system.
+ producing an action $a^v_t$ that will then modify the global state of the system.
  \ga{todo: put pseudo code to express this dynamics}

### 2023-05-16 paper.tex
chore: more on notation

@@
  we thereafter describe $G_f$ as a graph that contains the feature set $f_v$ for each node $v \in V$.
+ Note that, when we refer to $G_f$ and $G_o$ we are referring to the same graph $G$ but with different node features.
 %\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
@@
 
-  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G^i_o)$ 
-   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$;
+  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G'^i_o)$ 
+   from $\mathcal{D}$ and merge them in $(G^b_o,G^b_a,G^b_r,G'^b_o)$;
 
-  Compute the target Q-value for each node in the graph: 
-  $y = G_r + \gamma * \max{a'} Q(s'i,a';\theta^{-})$;
+  Compute the target Q-value for each node $v$ in the graph $G$: 
+  $y = G^b_r + \gamma * \max{a'} Q(o'i,a';\theta^{-})$;
 

### 2023-05-16 paper.tex
more comments

--- 2023-05-16 paper.tex
@@
  Each drone can only communicate directly with its own neighbourhood $\mathcal{N}$, 
- which in this case depends on a $\mathbb{O}$ range,
+ which in this case depends on a $\mathbb{O}$ range. \lukas{$\mathbb{O}$  not defined.}
 From the neighbourhood, 
@@
 \subsection{Goal}
- The objective of this scenario is to:
-\begin{itemize}
-\item Maximise the number of drones within the phenomena
+ The objective of this scenario is threefold:
+\begin{enumerate}
+\item Maximise the number of drones within the phenomena ($ndp$)
 \item Minimise the number of drones without a neighbourhood
 \item Maximise the coverage of the system
-\end{itemize}
+\end{enumerate}
 
@@
 
- Regarding the first point, the reward function is simple and is defined as:
-
+% Regarding the first point, the reward function is simple and is defined as:
+Formally, we define the reward of a drone being within the phenomena as
  \begin{equation*}
@@
   1 & \text{if } \zeta_v > 0 \\
-  0 & \text{otherwise} 
+  0 & \text{otherwise.} 
  \end{cases}
@@
 
-This will lead the system to prefer a configuration in which every node is present in the system. 
+ I.e., a drone is considered within the phenomena as soon as the drone can sense the phenomenon.
+This will lead the system to prefer a configuration in which every node is present in the system. \lukas{do you mean `... is present within the phenomenon'?}
 %
-The second point is important because if the system breaks into many scattered drones, 
- they are not able to move in the environment due to the reduced observability of the system. 
+The second element in the objective function ensures cohesion among the agents.
+This is important because if the system breaks into many scattered drones, the observability of the phenomenon is reduced, limiting the ability of the agents to move appropriately in the environment.
+% they are not able to move in the environment due to the reduced observability of the system. 
  In this case, the reward is defined as:
-
  \begin{equation*}
@@
   we decided to calculate the maximum and minimum distance of the neighbourhood and give more value in case these two values are similar.
+\lukas{minimum / maximum distance to what?
+the next neighbour?
+measuring the phenomena?}
 %
@@
 \end{equation*}
-The final reward function is defined as:
+where 
+$d_{max}$ and $d_{min}$ describes the maximum and minimum distance of the node to \lukas{??}, respectively. This distance is normalized by the maximum communication range of each agent $\mathbb{O}$.
+\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
+The final reward function for an agent $v$ is defined as:
 \begin{equation*}

### 2023-05-16 paper.tex
chore: more on notation

@@
 
-  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G^i_o)$ 
-   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$; \lukas{how does the merge work? do we explain this?}
+  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G'^i_o)$ 
+   from $\mathcal{D}$ and merge them in $(G^b_o,G^b_a,G^b_r,G'^b_o)$; \lukas{how does the merge work? do we explain this?}
 
-  Compute the target Q-value for each node in the graph: 
-  $y = G_r + \gamma * \max{a'} Q(s'i,a';\theta^{-})$;
+  Compute the target Q-value for each node $v$ in the graph $G$: 
+  $y = G^b_r + \gamma * \max{a'} Q(o'i,a';\theta^{-})$;
 

### 2023-05-16 paper.tex
minor
chore: improve the algorithm description

@@
  Note that, when we refer to $G_f$ and $G_o$ we are referring to the same graph $G$ but with different node features.
+ Also, to access the feature set $f_v$ of a node $v \in V$ we use the notation $f_v$ or $G_f[v]$.
 %\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
@@
   \While{not done}{
-  Observe current graph observations $G_o$; \lukas{whats $G_o$?}
+  Observe current graph observations $G_o$;
 
-%  $r \leftarrow $ randomNumber();
-%  
-%	\If{$r \leq \epsilon$}{select a random graph action $G_a$}
-%	\Else{select $G_a =$\\
-%		$\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;}
-%  
-  With probability $\epsilon$ select a random graph action $G_a$, 
-  otherwise select $G_a =$ 
-  $\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;
+  \uIf{random $< \epsilon$}{
+    select a random action $a$;
+  }
+  \Else{
+    $G_q = Q(G_o, \theta)$;
+    $a = \{ v \in G_q | a_v \in argmax_{a_v} G_q[v](a_v) \}$;
 
-  Execute the collective action $G_a$ in the environment $\mathbb{E}$ and observe a graph-level reward $G_r$ and the next state $G_o'$;
+  }
+
+  Execute the collective action $G_a$ in the environment $\mathbb{E}$ and observe a graph-level reward $G_r$ and the next observation $G_o'$;
 
@@
 
-  Compute the target Q-value for each node $v$ in the graph $G$: 
-  $y = G^b_r + \gamma * \max{a'} Q(o'i,a';\theta^{-})$;
+  Compute the target Q-value for each node $v$ in the graph $G^b$: 
+  $y_v = G^b_r[v] + \gamma * \max{a'} Q(G'^b_o[v],G^b_a[a'];\theta^{-})$;
 
-  Compute the current expected value for each node in the graph:
-  $y^* = Q(s'i,a';\theta)$;
+  Compute the current expected value for each node $v$ in the graph $G^b$:
+  $y_v^* = Q(G^b_o[v],G^b_a[v];\theta)$;
 
-  Update the current network weights using gradient descent: $\theta \leftarrow \theta - \alpha \nabla{\theta} \frac{1}{|G_o|} (y - y^*)^2$;
+  Update the current network weights using gradient descent: $\theta \leftarrow \theta - \alpha \nabla{\theta} \frac{1}{|G^b|} (y - y^*)^2$;
 
@@
 \begin{enumerate}
-\item Maximise the number of drones within the phenomena ($ndp$)
+\item Maximise the number of drones within the phenomena
 \item Minimise the number of drones without a neighbourhood

### 2023-05-16 paper.tex
chore: minors

@@
   The dots represent the robots, 
-  the circle area on the top right represent the phenomenon to be monitored, 
-  and the gray area represent the area covered by the robots.}
+  the circle area on the top right represents the phenomenon to be monitored, 
+  and the grey area represents the area covered by the robots.}
   \label{fig:simulations}

### 2023-05-16 paper.tex
chore: try to solve lukas concerns about single agent

@@
 %
-Despite appearing centralised, this system proposes a completely decentralised execution. 
-\lukas{maybe: ``While the graph, containing the aggregate information, might appear as global knowledge, this is not the case as the information is never aggregated globally.
-The individual nodes only combine information from their local neighbourhood.''}
+While the graph, containing the aggregate information, 
+ might appear as global knowledge, 
+ this is not the case as the information is never aggregated globally. 
+The individual nodes only combine information from their local neighbourhood.
 %
@@
  the use of \acp{GNN} allows us to define policies on a variable neighbourhood, which is essential in such systems as this can change due to the applied neighbourhood policy.
- It is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes~\cite{??}. 
+ It is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes~\cite{DBLP:journals/aiopen/ZhouCHZYLWLS20,DBLP:conf/nips/KnyazevTA19}. 
 % 
@@
 % 
- During learning, it is as if we have a \emph{single} agent collectively choosing which action to take (i.e., resulting action graph).
-\lukas{this makes it sound like a singe agent has all information, selecting the actions for all other agents...}
+This is because, even though the actions are produced using only local and neighbourhood information, 
+ during the learning phase, we have access to the internal graph, 
+ which will influence the policy through non-local information 
+ during the backpropagation.
 %

### 2023-05-16 paper.tex
chore; add images for the testing phase

@@
 In this paper, we propose a novel approach utilising aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
-\lukas{can we have a very short summary of how it works here.
-something along the lines of ``First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent.
+First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent.
 Second, we utilise a \ac{GNN} within the \ac{DQN}.
 Here the \ac{GNN} provides node embeddings for each agent, serving as input for the \ac{DQN}.
-The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.''}
-We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
+The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.'
+%We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
 %
 This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
-Specifically, 
- our approach utilises aggregate computing to disseminate information by manipulating a computational field, 
- which is a distributed data structure that maps devices to information. 
- This field functions as a dynamic environment in which information is continuously updated and diffused throughout the system. 
- By leveraging this layer, we can build collective distributed intelligence using \ac{GNN},
-  which uses the information in the field to determine the best action for a given task.
-\ga{first draft, need a refinement}
+%Specifically, 
+% our approach utilises aggregate computing to disseminate information by manipulating a computational field, 
+% which is a distributed data structure that maps devices to information. 
+% This field functions as a dynamic environment in which information is continuously updated and diffused throughout the system. 
+% By leveraging this layer, we can build collective distributed intelligence using \ac{GNN},
+%  which uses the information in the field to determine the best action for a given task.
+%\ga{first draft, need a refinement}
 The remainder of this paper is structured as follows.
@@
 The first point is a natural extension because we work on graphs rather than simple values.
+This also influences how we create a batch of experiences to train the network.
+ In fact, we sample a batch of graphs from the replay buffer,
+ and then we merge them into a single graph,
+ which is then used to train the network.
+ This process is called \emph{graph mini-batching}~\cite{DBLP:journals/corr/abs-1903-02428,wang2019deep}
+ and its main purpose is to pass an entire batch of graphs to the same \ac{GNN} for improved performance.
 %
@@
 The proposed simulation consists 
- of a set of robots placed in a 2D space capable of perceiving the intensity 
+ of a set of agents placed in a 2D space capable of perceiving the intensity 
  of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
@@
 %
-Additionally, each node has a coverage range $\omega$ 
+Additionally, each agent has a coverage range $\omega$ 
  that describes the area it can monitor. 
- Each drone can only communicate directly with its own neighbourhood $\mathcal{N}$, 
- which in this case depends on a $\mathbb{O}$ range. \lukas{$\mathbb{O}$  not defined.}
+ Each agent can only communicate directly with its own neighbourhood $\mathcal{N}$, 
+ which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. \lukas{$\mathbb{O}$  not defined.}
 From the neighbourhood, 
- each node can also perceive the direction of the other nodes.
+ each agent can also perceive the direction of the other agents.
 %
-Each node moves following a certain action composed of two components ($\alpha$, $\theta$) 
+Each agent moves following a certain action composed of two components ($\alpha$, $\theta$) 
  which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
@@
  these three components must be encoded in a reward function 
- that provides an estimate of the current action taken by a given robot.
+ that provides an estimate of the current action taken by a given agent.
 
 % Regarding the first point, the reward function is simple and is defined as:
-Formally, we define the reward of a drone being within the phenomena as
+Formally, we define the reward of a drone being within the phenomenon as
  \begin{equation*}
@@
  I.e., a drone is considered within the phenomena as soon as the drone can sense the phenomenon.
-This will lead the system to prefer a configuration in which every node is present in the system. \lukas{do you mean `... is present within the phenomenon'?}
+This will lead the system to prefer a configuration in which every agent is present within the phenomenon.
 %
 The second element in the objective function ensures cohesion among the agents. 
-This is important because if the system breaks into many scattered drones, the observability of the phenomenon is reduced, limiting the ability of the agents to move appropriately in the environment.
+ This is important because if the system breaks into many scattered agents, 
+ the observability of the phenomenon is reduced, 
+ limiting the ability of the agents to move appropriately in the environment.
 % they are not able to move in the environment due to the reduced observability of the system. 
@@
   we decided to calculate the maximum and minimum distance of the neighbourhood and give more value in case these two values are similar. 
-\lukas{minimum / maximum distance to what?
-the next neighbour?
-measuring the phenomena?}
+  \lukas{minimum / maximum distance to what? the next neighbour? measuring the phenomena?}
 %
@@
 where 
-$d_{max}$ and $d_{min}$ describes the maximum and minimum distance of the node to \lukas{??}, respectively. This distance is normalized by the maximum communication range of each agent $\mathbb{O}$.
+$d_{max}$ and $d_{min}$ describes the maximum and minimum distance of the agent to \lukas{??}, respectively. 
+This distance is normalized by the maximum communication range of each agent $\mathbb{O}$.
 \lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
@@
   \centering
-  \begin{subfigure}[b]{\linewidth}
+  \begin{subfigure}[b]{0.49\linewidth}
     \includegraphics[width=\linewidth]{imgs/coverage-in-time}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{\linewidth}
+  \begin{subfigure}[b]{0.49\linewidth}
     \includegraphics[width=\linewidth]{imgs/reward-in-time}
@@
 \end{figure}
+%% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test
+
+\begin{figure*}
+  \centering
+  \begin{subfigure}[b]{0.49\linewidth}
+    \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
+    \caption{Coverage test}
+    \label{fig:coverage-test}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\linewidth}
+    \includegraphics[width=\linewidth]{imgs/inside-test}
+    \caption{Inside test}
+    \label{fig:inside-test}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\linewidth}
+    \includegraphics[width=\linewidth]{imgs/inside-two-test}
+    \caption{Inside two test}
+    \label{fig:inside-two-test}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\linewidth}
+    \includegraphics[width=\linewidth]{imgs/coverage-two-test}
+    \caption{Coverage two test}
+    \label{fig:coverage-two-test}
+  \end{subfigure}
+  \caption{Test results, \ga{todo}}
+  \label{fig:test}
+\end{figure*}
 \section{Conclusion and Future Work}

### 2023-05-16 paper.tex
remove old comment
handle some question from lukas in evaluation

@@
   Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G'^i_o)$ 
-   from $\mathcal{D}$ and merge them in $(G^b_o,G^b_a,G^b_r,G'^b_o)$; \lukas{how does the merge work? do we explain this?}
+   from $\mathcal{D}$ and merge them in $(G^b_o,G^b_a,G^b_r,G'^b_o)$;
 
@@
 \label{sec:eval}
-
-%% Put three subfigures of simulations, start, after, end
-\begin{figure}[t]
+\begin{figure}
   \centering
-  \begin{subfigure}[b]{0.32\linewidth}
-      \includegraphics[width=\textwidth]{imgs/start.png}
-      \caption{Initial state}
-      \label{fig:initial}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.32\linewidth}
-      \includegraphics[width=\textwidth]{imgs/after.png}
-      \caption{After 50 steps}
-      \label{fig:after}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.32\linewidth}
-      \includegraphics[width=\textwidth]{imgs/end.png}
-      \caption{After 200 steps}
-      \label{fig:end}
-  \end{subfigure}
-  \caption{Simulation of the proposed approach. 
+  \includegraphics[width=0.3\textwidth]{imgs/start.png}
+  \label{fig:example}
+  \caption{Simulation of the case study scenario in Alchemist. 
   The dots represent the robots, 
@@
   and the grey area represents the area covered by the robots.}
-  \label{fig:simulations}
 \end{figure}
@@
  or monitoring the water levels in a canal. The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
- \lukas{``Initially we consider a stationary phenomenon before expanding towards a moving phenomenon.
-Additionally, the phenomenon can change its shape during runtime.
+ %
+ Initially, 
+  we perform the training phase using a stationary phenomenon before expanding towards a moving phenomenon in the test phase. 
+  %Additionally, the phenomenon can change its shape during runtime. 
   Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function. 
-Figure~\ref{fig:simulations} illustrates a set of 25 agents and a single, stationary phenomenon with a with a 3D Gaussian distribution.''}
+  Figure~\ref{fig:example} illustrates a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian distribution.''
 
@@
 \begin{enumerate}
-\item Maximise the number of drones within the phenomena
-\item Minimise the number of drones without a neighbourhood
+\item Maximise the number of agents within the phenomena
+\item Minimise the number of agents without a neighbourhood
 \item Maximise the coverage of the system
@@
 % Regarding the first point, the reward function is simple and is defined as:
-Formally, we define the reward of a drone being within the phenomenon as
+Formally, we define the reward of an agent being within the phenomenon as
  \begin{equation*}
@@
  \end{equation*}
-
- I.e., a drone is considered within the phenomena as soon as the drone can sense the phenomenon.
+Namely, an agent is considered within the phenomena as soon as the drone can sense the phenomenon.
+%
 This will lead the system to prefer a configuration in which every agent is present within the phenomenon.
@@
 Finally, to maximise the coverage, 
-  we decided to calculate the maximum and minimum distance of the neighbourhood and give more value in case these two values are similar. 
-  \lukas{minimum / maximum distance to what? the next neighbour? measuring the phenomena?}
+  we define a reward function that favours the maximum distance between the agents
+  equal to the coverage range $\omega$.
 %
@@
 \begin{equation*}
-R^{C}_{v} = \frac{d_{max} - d_{min}}{\mathbb{O}}
+R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
 \end{equation*}
-where 
-$d_{max}$ and $d_{min}$ describes the maximum and minimum distance of the agent to \lukas{??}, respectively. 
-This distance is normalized by the maximum communication range of each agent $\mathbb{O}$.
-\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
+where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
+%
+%\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
 The final reward function for an agent $v$ is defined as:
@@
 \subsection{Discussion and Results}
-%% Subfigures of coverage-in-time and reward-in-time
-
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/start.png}
+      \caption{Initial configuration}
+      \label{fig:initial}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/after.png}
+      \caption{After 50 steps}
+      \label{fig:after}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/end.png}
+      \caption{After 200 steps}
+      \label{fig:end}
+  \end{subfigure}
+  \caption{Test results of the proposed approach. 
+  The agents go towards the phenomenon and try to cover it without collapsing into a single point.}
+  \label{fig:simulations}
+\end{figure}
 \begin{figure}

### 2023-05-16 paper.tex
minor in evaluation

@@
   \centering
-  \includegraphics[width=0.3\textwidth]{imgs/start.png}
+  \includegraphics[width=0.3\textwidth]{imgs/example.png}
   \label{fig:example}
@@
   %Additionally, the phenomenon can change its shape during runtime. 
-  Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function. 
-  Figure~\ref{fig:example} illustrates a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian distribution.''
+  Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function. '
 
 \subsection{Scenario}
-The proposed simulation consists 
- of a set of agents placed in a 2D space capable of perceiving the intensity 
+Figure~\ref{fig:example} illustrates the scenario used during the learning, 
+which consists of a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian 
+Each agent can perceive the presence
  of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
@@
 %
-Additionally, each agent has a coverage range $\omega$ 
+Additionally, each agent has a coverage range $\omega$ (fixed to 7.5 meters) 
  that describes the area it can monitor. 

### 2023-05-16 paper.tex
chore: more in the evaluation

@@
   \includegraphics[width=0.3\textwidth]{imgs/example.png}
-  \label{fig:example}
+
   \caption{Simulation of the case study scenario in Alchemist. 
@@
   and the grey area represents the area covered by the robots.}
+  \label{fig:example}
 \end{figure}
@@
   %Additionally, the phenomenon can change its shape during runtime. 
-  Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function. '
+  Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function.
+As a simulation environment, we use Alchemist~\cite{alchemist}, 
+ a simulator for multi-agent systems that allows us to simulate the swarm behaviour of the agents 
+ and the phenomenon of interest.
+For the \ac{GNN}, we use the implementation provided by PyTorch Geometric~\cite{Fey/Lenssen/2019}, 
+ which is a library for deep learning on graphs built on top of PyTorch~\cite{torch}.
+Finally, we use ScaFi~\cite{casadei2022scafi} as the aggregate programming language to support 
+ our field-informed approach.
 
@@
  Each agent can only communicate directly with its own neighbourhood $\mathcal{N}$, 
- which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. \lukas{$\mathbb{O}$  not defined.}
+ which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. %\lukas{$\mathbb{O}$  not defined.}
 From the neighbourhood, 
@@
  which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
- \lukas{drone, robot, agent, node...}
+ %\lukas{drone, robot, agent, node...}
 \subsection{Goal}

### 2023-05-16 paper.tex
more comments and details on scenarios

@@
 From this field, we construct the feature vector $f_v$ for each node $v$ in $G_t$ as follows:
- $f_v = (\theta^v_t, o^v_t)$. \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
- \lukas{we have not introduced $G'_t$ yet}
+ $f_v = (\theta^v_t, o^v_t)$.
+%  \lukas{we used $o_v$ to describe the observation of node $v$ in section II.C - so this would not include field information, right?}
+% \lukas{we have not introduced $G'_t$ yet}
  %where $\omega$ are other local information gather from sensors (like temperature, distances from neighbourhood, etc.). 
@@
 i.e., to function globally over the entire graph and locally only over the neighbourhood. 
- \lukas{Global information is not available to the agent, right? The agent only provide locally aggregated information (resulting from the AC part)?}
+Importantly, each agent only has local information from itself and its neighbourhood to utilise in the \ac{GNN}.
+% \lukas{Global information is not available to the agent, right? The agent only provide locally aggregated information (resulting from the AC part)?}
 %
@@
 \subsection{Scenario}
-Figure~\ref{fig:example} illustrates the scenario used during the learning, 
+Figure~\ref{fig:example} illustrates the scenario used during the learning \lukas{is this training the \ac{GNN}? }, 
 which consists of a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian 
@@
  which respectively describe the angle and intensity of the movement (i.e., the velocity vector). 
+ \lukas{$\theta$ used for target weights...} 
+ \lukas{not sure if this paragraph should be its own subsection. or at least it needs  different title.}
+ For the evaluation, we explore three different types of scenarios.
+The first type considers only a single, stationary phenomenon.
+The second type considers multiple phenomena, requiring the agents to make a decision which phenomena to cover.
+Finally, the third type considers a single moving phenomena which the agents have to follow throughout the simulation.
+In all three types of scenarios, the shape and underlying distribution is initially unknown to the agents but remains constant for the entire duration of the simulation \lukas{correct?}.
+We generated random scenarios for each type of scenario.
+Examples of such types are shown in Figure~\ref{??} \lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?} We compare our field-informed reinforcement learning approach against a baseline approach utilising a \ac{MLP} as well as an approach only relying on \acp{GNN}.
+In all approaches, the underlying neural network (i.e., the \ac{MLP} and the \ac{GNN}) are trained with \lukas{xyz} samples of a single, stationary phenomena.
+\lukas{we should describe the training a bit.}
+ \lukas{do we need more information on the MLP?}
+
  %\lukas{drone, robot, agent, node...}
@@
   we define a reward function that favours the maximum distance between the agents
-  equal to the coverage range $\omega$.
+  equal to the coverage range $\omega$. This will minimise multiple agents covering a common area.
 %
@@
   \centering
-  \begin{subfigure}[b]{0.49\linewidth}
+  \hfill
+  \begin{subfigure}[b]{0.96\linewidth}
     \includegraphics[width=\linewidth]{imgs/coverage-in-time}
@@
     \label{fig:coverage}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\linewidth}
+  \end{subfigure}\\
+  \hfill
+  \begin{subfigure}[b]{0.99\linewidth}
     \includegraphics[width=\linewidth]{imgs/reward-in-time}
@@
   \centering
-  \begin{subfigure}[b]{0.49\linewidth}
+  \begin{subfigure}[b]{0.9\linewidth}
     \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\linewidth}
+  \begin{subfigure}[b]{0.9\linewidth}
     \includegraphics[width=\linewidth]{imgs/inside-test}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\linewidth}
+  \begin{subfigure}[b]{0.85\linewidth}
     \includegraphics[width=\linewidth]{imgs/inside-two-test}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\linewidth}
+  \begin{subfigure}[b]{0.85\linewidth}
     \includegraphics[width=\linewidth]{imgs/coverage-two-test}

### 2023-05-16 paper.tex
feat: more structure

--- 2023-05-16 paper.tex
@@
   \centering
-  \includegraphics[width=0.3\textwidth]{imgs/example.png}
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/example.png}
+      \caption{}
+      \label{fig:static}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/two-zones.png}
+      \caption{}
+      \label{fig:two}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+      \includegraphics[width=\textwidth]{imgs/example-moving.png}
+      \caption{}
+      \label{fig:moving}
+  \end{subfigure}
 
-  \caption{Simulation of the case study scenario in Alchemist. 
-  The dots represent the robots, 
-  the circle area on the top right represents the phenomenon to be monitored, 
-  and the grey area represents the area covered by the robots.}
-  \label{fig:example}
+  %\includegraphics[width=0.3\textwidth]{imgs/example.png}
+
+  \caption{Simulations of the case study scenario in Alchemist. 
+  The dots represent the agent, 
+  the circle area on represents the phenomenon to be monitored.
+  \Cref{fig:static} represents the scenario used during training. 
+  The others are used in the test phase to evaluate the policy found. }
+  \label{fig:scenarios}
 \end{figure}
@@
 \subsection{Scenario}
-Figure~\ref{fig:example} illustrates the scenario used during the learning, 
-which consists of a set of 25 agents and a single, stationary phenomenon with a 3D Gaussian 
+Figure~\ref{fig:static} illustrates the scenario used during the learning \lukas{is this training the \ac{GNN}? }, 
+which consists of a set of 25 agents placed in a 2D grid large 1000x1000 meters and a single, stationary phenomenon with a 3D Gaussian 
 Each agent can perceive the presence
@@
 %
-Additionally, each agent has a coverage range $\omega$ (fixed to 7.5 meters) 
+Additionally, each agent has a coverage range $\omega$ (fixed to 75 meters) 
  that describes the area it can monitor. 
  Each agent can only communicate directly with its own neighbourhood $\mathcal{N}$, 
- which in this case depends on a $\mathbb{O}$ range fixed to 30 meters. %\lukas{$\mathbb{O}$  not defined.}
+ which in this case depends on a $\mathbb{O}$ range fixed to 300 meters. %\lukas{$\mathbb{O}$  not defined.}
 From the neighbourhood, 
@@
 %
-Each agent moves following a certain action composed of two components ($\alpha$, $\theta$) 
+Each agent moves following a certain action composed of two components (r, i) 
  which respectively describe the angle and intensity of the movement (i.e., the velocity vector).
+Since we used a value-based approach, 
+ the action space is discrete and composed of 18 possible angles and 3 possible intensities.
+In particular, for the angles, we have quantized to 20 degrees, and for the velocities, we have selected [0, 5, 10] m/s.
+%
+Each agent will produce a computation field with which they will try to approximate the direction of the phenomenon of interest. 
+%
+The program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
+This can be expressed in Scafi as follows:
+\begin{lstlisting}[mathescape=true]
+val source = maxHood(nbr(sense($\zeta$))) == sense($\zeta$)
+G(source, Point3D.Zero, _ + nbrVector())
+\end{lstlisting}
+where \texttt{maxHood} is a function that returns the maximum value of the neighbourhood, 
+ \texttt{nbr} is a function that returns the value of the neighbourhood, 
+\texttt{sense} is a function that returns the value of the sensor, and
+\texttt{nbrRange} is a function that returns an approximate direction for each agent in the neighbourhood.
+%
+this value will then be fed into the \ac{GNN} to compute the action to be performed.
+ %\lukas{not sure if this paragraph should be its own subsection. or at least it needs  different title.}
+
  %\lukas{drone, robot, agent, node...}
@@
 %
+\subsection{Training Phase}
+The learning process was divided into 100 episodes, each consisting of 200 steps, resulting in a total of 20,000 experiences.
+%
+For each episode, 
+ the 25 agents are semi-randomly positioned on a grid without knowing the correct position of the phenomenon, 
+ but that is fixed in the top right corner. 
+The feature set used by the \ac{GNN} created for each node consists of the vector computed by the aggregated program and the value of the local sensor $\zeta$.
+%
+In this case, we chose to use an exponential epsilon decay, defined as:
+\begin{equation*}
+\epsilon = \epsilon_{min} + (\epsilon_{max} - \epsilon_{min}) \cdot e^{-\lambda \cdot e}
+\end{equation*}
+where e is the current episode number. 
+This leads to a high number of random actions at the beginning and gradually shifts towards exploitation in the later episodes. 
+%
+In our training process, 
+ we set $\epsilon_{min} = 0.02$, $\epsilon_{max} = 0.99$, and $\lambda = 0.1$.
+$\gamma$ was set to 0.99, 
+ as we want to give more value to future returns, 
+ aiming to achieve good coverage by continuously tracking the phenomenon. 
+The neural network structure used consists of a layer of SuperGAT~\cite{DBLP:journals/corr/abs-2204-04879} -- A \ac{GNN} based on attention mechanisms -- and a layer of MLP. 
+The hidden size was set to 256.
+%
+As the reward function is defined as a regret, 
+ we decided to use the Huber loss function, 
+ which is a combination of the $L_1$ and $L_2$ loss functions.
+%
+The loss function is defined as:
+\begin{equation*}
+L = \begin{cases}
+  \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
+  \delta (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise.} 
+ \end{cases}
+\end{equation*}
+This function is used to penalise the agent if the action taken is too far from the optimal action.
+%
+We use the RMSprop optimiser with a learning rate of 0.0001.
+%
+\subsection{Test phase}
+For the evaluation, 
+ we explore three different types of scenarios.
+%
+The first type considers only a single, stationary phenomenon like the one used in the training phase. 
+%
+The second type considers multiple phenomena, requiring the agents to make a decision about which phenomena to cover. 
+%
+Finally, the third type considers a single moving phenomenon which the agents have to follow throughout the simulation. 
+%
+In all three types of scenarios, the shape and underlying distribution are initially unknown to the agents but remain constant for the entire duration of the simulation. 
+%
+We generated random scenarios for each type of scenario. 
+%
+Examples of such types are shown in~\Cref{fig:scenarios}. %\lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?}
 
-%\paragraph*{Variant A}
-%\paragraph*{Variant B}
-%\paragraph*{Variant C}
+\subsection{Baseline}
+We compare our field-informed reinforcement learning approach against a baseline approach utilising a \ac{MLP} as well as an approach only relying on \acp{GNN}.
+In all approaches, the underlying neural network (i.e., the \ac{MLP} and the \ac{GNN}) are trained with the same parameters of the proposed approach with a single, stationary phenomenon.
+
+\subsection{Metrics}
 
@@
       \includegraphics[width=\textwidth]{imgs/start.png}
-      \caption{Initial configuration}
+      \caption{Start}
       \label{fig:initial}
@@
   \caption{Test results of the proposed approach. 
-  The agents go towards the phenomenon and try to cover it without collapsing into a single point.}
+  The agents go towards the phenomenon and try to cover it without collapsing into a single point.
+  the grey area represents the area covered by the robots (last figure on the right).}
   \label{fig:simulations}

### 2023-05-16 paper.tex
draft: on metrics

@@
 \usepackage{subcaption}
+\usepackage{bbm}
 \usepackage[ruled]{algorithm2e}
@@
 \subsection{Metrics}
+We evaluate the performance of the different approaches by measuring the coverage of the phenomenon over time.
+%
+The coverage is defined as the percentage of the phenomenon covered by the agents.
+%
+Specifically, we can measure the coverage as the intersection over the union of the phenomenon and the agents' view range.
+%
+Formally, first, we define the overall area covered by the agents as:
+\begin{equation*}
+\Omega = \bigcup_{v \in V} \omega_v
+\end{equation*}
+
+Finally, we define the coverage as:
+\begin{equation*}
+C = \frac{|\Omega \cap \mathcal{P}|}{|\mathcal{P}|}
+\end{equation*}
+where $\omega$ is the viewing range of the agents and $\mathcal{P}$ is the phenomenon.
+This is computed for each step of the simulation.
+
+We also measure the number of agents that are within the phenomenon at each step of the simulation.
+This will be a measure of how well the agents are tracking the phenomenon.
+\ga{todo improve the description of the coverage}
 

### 2023-05-17 paper.tex
draft: on metrics
chore: fix mlp images

@@
 Figure~\ref{fig:static} illustrates the scenario used during the learning \lukas{is this training the \ac{GNN}? }, 
-which consists of a set of 25 agents placed in a 2D grid large 1000x1000 meters and a single, stationary phenomenon with a 3D Gaussian 
+which consists of a set of $\mathcal{P} = 25$ agents placed in a 2D grid large 1000x1000 meters and a single, stationary phenomenon with a 3D Gaussian 
 Each agent can perceive the presence
@@
 Since we used a value-based approach, 
- the action space is discrete and composed of 18 possible angles and 3 possible intensities.
+ the action space $\mathcal{A}$ is discrete and composed of 18 possible angles and 3 possible intensities.
 In particular, for the angles, we have quantized to 20 degrees, and for the velocities, we have selected [0, 5, 10] m/s.
@@
 
-\subsection{Baseline}
+\subsection{Baselines}
 We compare our field-informed reinforcement learning approach against a baseline approach utilising a \ac{MLP} as well as an approach only relying on \acp{GNN}. 
+%
 In all approaches, the underlying neural network (i.e., the \ac{MLP} and the \ac{GNN}) are trained with the same parameters of the proposed approach with a single, stationary phenomenon.
-
+%
+The \ac{MLP} uses the same feature set as the \ac{GNN} in the proposed approach, but differently, it uses a standard DQN approach without leveraging the graph structure. 
+%
+The \ac{GNN} instead of using the field information, use directly the position of nodes and the local sensor value as input features.
+%
+These baselines are used to verify the effectiveness of the components used in the field-informed \ac{RL}. 
+%
+Indeed the \ac{MLP} baseline is used to verify the effectiveness of the \ac{GNN} in the proposed approach, while the \ac{GNN} baseline is used to verify the effectiveness of the field information in the proposed approach.
 \subsection{Metrics}
@@
 \end{equation*}
-where $\omega$ is the viewing range of the agents and $\mathcal{P}$ is the phenomenon.
+where $\mathcal{P}$ is the area of the phenomenon.
 This is computed for each step of the simulation. 
-
+In the training phases, 
+ we measure the average coverage in each episode, and the total reward obtained by the agents at each episode of the simulation. 
+%
 We also measure the number of agents that are within the phenomenon at each step of the simulation.
@@
     \includegraphics[width=\linewidth]{imgs/coverage-in-time}
-    \caption{Coverage in time}
+    \caption{Average coverage for each episode in training }
     \label{fig:coverage}
@@
     \includegraphics[width=\linewidth]{imgs/reward-in-time}
-    \caption{Reward in time}
+    \caption{Reward during training}
     \label{fig:reward}
   \end{subfigure}
-  \caption{Coverage and reward in time}
+  \caption{Training results of field-informed \ac{RL}. 
+  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
+  }
   \label{fig:coverage-reward}
@@
     \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
-    \caption{Coverage test}
+    \caption{Coverage in the three types of scenarios}
     \label{fig:coverage-test}
@@
     \includegraphics[width=\linewidth]{imgs/inside-test}
-    \caption{Inside test}
+    \caption{Number of agents inside the phenomenon in the three types of scenarios}
     \label{fig:inside-test}
@@
     \includegraphics[width=\linewidth]{imgs/inside-two-test}
-    \caption{Inside two test}
+    \caption{Two zones scenario: agent inside each phenomenon}
     \label{fig:inside-two-test}
@@
     \includegraphics[width=\linewidth]{imgs/coverage-two-test}
-    \caption{Coverage two test}
+    \caption{Two zones scenario: coverage of each area}
     \label{fig:coverage-two-test}

### 2023-05-17 paper.tex
chore: more on discussion
chore: minors

@@
 %
+Finally, we use a replay buffer of size 1000 to store the graph experiences and a batch size of 32.
 \subsection{Test phase}
@@
 The \ac{MLP} uses the same feature set as the \ac{GNN} in the proposed approach, but differently, it uses a standard DQN approach without leveraging the graph structure.
+Moreover, we increase the batch size to 512 and the replay buffer to 10000 since we record 25 agents' experiences for each step instead of one graph experience.
 %
@@
   \end{subfigure}
-  \caption{Test results of the proposed approach. 
+  \caption{Qualitative test results of the proposed approach. 
   The agents go towards the phenomenon and try to cover it without collapsing into a single point.
@@
   \end{subfigure} 
-  \caption{Test results, \ga{todo}}
+  \caption{Quantative test results in the three types of scenarios.
+  We can see that the proposed approach is able to cover and track the phenomenon better then the baselines.
+  }
   \label{fig:test}

### 2023-05-17 paper.tex
draft discussion

@@
 \subsection{Discussion and Results}
+The results of the training process are summarized in \Cref{fig:training}. 
+%
+Specifically, we observe that the proposed version achieves higher coverage and total reward compared to other approaches. 
+%
+Interestingly, despite the global information available in \acp{GNN} without fields, they fail to converge to a good result like the one obtained with the field. 
+%
+This outcome was expected, as the computed field helps agents encode the necessary information to navigate towards the phenomenon.
+%
+Furthermore, we note that \ac{GNN} combined with \ac{DQN} and graph replay buffer outperforms the simple \ac{MLP} informed field computation. 
+%
+This is because relying solely on \ac{MLP} and basic deep learning leads to non-stationary and unstable learning, 
+ as evident from the wider confidence interval of the reward over training time.
+
+Focusing now on the test phase, 
+ we observe that the field-informed version achieves higher coverage than the other approaches in all scenarios since it shows the ability of our solution to generalize to situations--see \Cref{fig:simulations}.
+% 
+We observe that the field-informed version successfully moves the nodes closer to the target phenomenon, 
+ distributing them evenly without collapsing into a single central point. 
+ \Cref{fig:test} quantitatively presents the results across various previously described scenarios.
+
+The \ac{GNN} versions demonstrate the capability to transfer the learned experience to the test phase whereas the MLP version fails to generalize. 
+%
+It is worth noting that, in the static configuration, 
+ once the desired configuration is achieved in the static case, 
+ the nodes cease to move, maintaining the found configuration.
+%
+Interesting observations arise when we use different scenarios from the training phase. 
+ Starting from a division into two zones, 
+ we notice that even in this case, the informed version finds a better configuration than the GNN counterpart. 
+%
+It exhibits both higher overall coverage and manages to divide the system into two equally covered parts. 
+%
+In contrast, the uninformed version also maintains a fair division but with significantly different coverage between the two parts, 
+ indicating an imbalance among the nodes in one of the zones.
+%
+Finally, the moving scenario emphasizes how the informed version generates a more robust policy for new scenarios. 
+ Indeed, we observe that the informed version maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
+%
+The uninformed GNN version, however, 
+ fails to generalize its movement behaviour, as evidenced by the simulations where the agents, once reaching the target zone, stop moving due to tracking issues.
+
+In conclusion, the results demonstrate how the proposed idea can generate more robust controllers. 
+ By guiding information flow in \acp{GNN}, 
+ we improve learning efficiency and alleviate the challenge of encoding relevant information. 
+%
+Nevertheless, we acknowledge the crucial role of \acp{GNN}. 
+ Our modified version of DQN, combined with \acp{GNN}, 
+ enables the discovery of robust behaviours in a few episodes, 
+ which is challenging to capture with \acp{MLP} combined with \ac{DQN}, even if we use field information.
+\begin{figure}
+  \centering
+  \hfill
+  \begin{subfigure}[b]{0.96\linewidth}
+    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
+    \caption{Average coverage for each episode in training }
+    \label{fig:coverage}
+  \end{subfigure}\\
+  \hfill
+  \begin{subfigure}[b]{0.99\linewidth}
+    \includegraphics[width=\linewidth]{imgs/reward-in-time}
+    \caption{Reward during training}
+    \label{fig:reward}
+  \end{subfigure}
+  \caption{Training results of field-informed \ac{RL}. 
+  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
+  }
+  \label{fig:training}
+\end{figure}
 \begin{figure}[t]
@@
 \end{figure}
-\begin{figure}
-  \centering
-  \hfill
-  \begin{subfigure}[b]{0.96\linewidth}
-    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
-    \caption{Average coverage for each episode in training }
-    \label{fig:coverage}
-  \end{subfigure}\\
-  \hfill
-  \begin{subfigure}[b]{0.99\linewidth}
-    \includegraphics[width=\linewidth]{imgs/reward-in-time}
-    \caption{Reward during training}
-    \label{fig:reward}
-  \end{subfigure}
-  \caption{Training results of field-informed \ac{RL}. 
-  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
-  }
-  \label{fig:coverage-reward}
-\end{figure}
+
 %% Create a subfigure of coverage-test, inside-test, inside-two-test and coverage-two-test
@@
   \caption{Quantative test results in the three types of scenarios.
-  We can see that the proposed approach is able to cover and track the phenomenon better then the baselines.
+  We can see that the proposed approach is able to cover and track the phenomenon better than the baselines.
   }

### 2023-05-17 paper.tex
some comments on evaluation

--- 2023-05-17 paper.tex
@@
 Figure~\ref{fig:static} illustrates the scenario used during the learning \lukas{is this training the \ac{GNN}? }, 
-which consists of a set of $\mathcal{P} = 25$ agents placed in a 2D grid large 1000x1000 meters and a single, stationary phenomenon with a 3D Gaussian 
+which consists of a set of 25 agents placed in a 2D grid large 1000x1000 meters and a single, stationary phenomenon with a 3D Gaussian 
 Each agent can perceive the presence
@@
 Since we used a value-based approach, 
- the action space $\mathcal{A}$ is discrete and composed of 18 possible angles and 3 possible intensities.
+ the action space is discrete and composed of 18 possible angles and 3 possible intensities.
 In particular, for the angles, we have quantized to 20 degrees, and for the velocities, we have selected [0, 5, 10] m/s.
@@
 For each episode, 
- the 25 agents are semi-randomly positioned on a grid without knowing the correct position of the phenomenon, 
- but that is fixed in the top right corner. 
+ the 25 agents are semi-randomly \lukas{what is semi-random?
+are the position fixed but the agents random?
+or is it in a specific layout (e.g.
+lattice) but with some randomness of the actual position?} positioned on a grid without knowing the correct position of the phenomenon,
+ but that is fixed in the top right corner.  \lukas{so for all 20,000 experiences, the phenomena is exactly as shown in fig 2a during training?}
 The feature set used by the \ac{GNN} created for each node consists of the vector computed by the aggregated program and the value of the local sensor $\zeta$.
@@
 As the reward function is defined as a regret, 
- we decided to use the Huber loss function, 
+ we decided to use the Huber loss function~\cite{??}, 
  which is a combination of the $L_1$ and $L_2$ loss functions.
@@
 \end{equation*}
-This function is used to penalise the agent if the action taken is too far from the optimal action.
+This function is used to penalise the agent if the action taken is too far from the optimal action. \lukas{whats $y, \hat{y}$ and $\delta$?}
 %
@@
 
-\subsection{Baselines}
+\subsection{Baseline}
 We compare our field-informed reinforcement learning approach against a baseline approach utilising a \ac{MLP} as well as an approach only relying on \acp{GNN}.
-%
 In all approaches, the underlying neural network (i.e., the \ac{MLP} and the \ac{GNN}) are trained with the same parameters of the proposed approach with a single, stationary phenomenon.
-%
-The \ac{MLP} uses the same feature set as the \ac{GNN} in the proposed approach, but differently, it uses a standard DQN approach without leveraging the graph structure. 
-%
-The \ac{GNN} instead of using the field information, use directly the position of nodes and the local sensor value as input features.
-%
-These baselines are used to verify the effectiveness of the components used in the field-informed \ac{RL}. 
-%
-Indeed the \ac{MLP} baseline is used to verify the effectiveness of the \ac{GNN} in the proposed approach, while the \ac{GNN} baseline is used to verify the effectiveness of the field information in the proposed approach.
+
 \subsection{Metrics}
@@
 \end{equation*}
-where $\mathcal{P}$ is the area of the phenomenon.
+where $\omega$ is the viewing range of the agents and $\mathcal{P}$ is the phenomenon.
 This is computed for each step of the simulation.
-In the training phases, 
- we measure the average coverage in each episode, and the total reward obtained by the agents at each episode of the simulation. 
-%
+
 We also measure the number of agents that are within the phenomenon at each step of the simulation.
@@
     \includegraphics[width=\linewidth]{imgs/coverage-in-time}
-    \caption{Average coverage for each episode in training }
+    \caption{Coverage in time}
     \label{fig:coverage}
@@
     \includegraphics[width=\linewidth]{imgs/reward-in-time}
-    \caption{Reward during training}
+    \caption{Reward in time}
     \label{fig:reward}
   \end{subfigure}
-  \caption{Training results of field-informed \ac{RL}. 
-  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
-  }
+  \caption{Coverage and reward in time}
   \label{fig:coverage-reward}
@@
     \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
-    \caption{Coverage in the three types of scenarios}
+    \caption{Coverage test}
     \label{fig:coverage-test}
@@
     \includegraphics[width=\linewidth]{imgs/inside-test}
-    \caption{Number of agents inside the phenomenon in the three types of scenarios}
+    \caption{Inside test}
     \label{fig:inside-test}
@@
     \includegraphics[width=\linewidth]{imgs/inside-two-test}
-    \caption{Two zones scenario: agent inside each phenomenon}
+    \caption{Inside two test}
     \label{fig:inside-two-test}
@@
     \includegraphics[width=\linewidth]{imgs/coverage-two-test}
-    \caption{Two zones scenario: coverage of each area}
+    \caption{Coverage two test}
     \label{fig:coverage-two-test}

### 2023-05-17 paper.tex
conclusion draft

--- 2023-05-17 paper.tex
@@
 \section{Conclusion and Future Work}
+In this paper, 
+ we have introduced a novel approach for constructing distributed controllers by leveraging aggregate computing to encode agent interactions, 
+ along with the combination of \ac{DQN} and \ac{GNN} for synthesizing distributed intelligence.
+The proposed \emph{Field-Informed reinforcement learning} approach offers a promising solution to the challenges faced in coordinating multi-agent systems. 
+By combining manual design and machine learning techniques, 
+ the approach enables agents to autonomously learn and adapt their behaviour while leveraging locally available information. 
+%
+The demonstrated success in the proposed case study in solving collective tasks underscores the potential impact of this approach in advancing the field of multi-agent systems and swarm robotics.
+
+Future research could explore its application in diverse domains and evaluate its scalability and robustness in increasingly complex scenarios. 
+%
+In addition, we also plan to take the approach to modern actor-critical solutions, which are better suited to modern swarm robotics problems because of the continuous action space.
 \ga{Page budget: 0.5}

### 2023-05-17 paper.tex
more comments

--- 2023-05-17 paper.tex
@@
  specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
- or monitoring the water levels in a canal. The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
+ or monitoring the water levels in a canal with multiple autonomous agents embodied in embedded devices (e.g., drones or IoT devices).
+The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
  %
@@
   Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function.
+\lukas{are we utiliising this in any way?
+i.e.
+do the agents respond to intensity of their measurement.}
 As a simulation environment, we use Alchemist~\cite{alchemist}, 
@@
  the action space $\mathcal{A}$ is discrete and composed of 18 possible angles and 3 possible intensities.
-In particular, for the angles, we have quantized to 20 degrees, and for the velocities, we have selected [0, 5, 10] m/s.
+In particular, the angles are quantized to 20 degrees, and for the velocities, we have selected [0, 5, 10] m/s.
 %
+\lukas{should the following go into section III?}
 Each agent will produce a computation field with which they will try to approximate the direction of the phenomenon of interest. 
@@
 
+\lukas{I would move this up into IV.A}
+
 \subsection{Baselines}
@@
 %
-In all approaches, the underlying neural network (i.e., the \ac{MLP} and the \ac{GNN}) are trained with the same parameters of the proposed approach with a single, stationary phenomenon.
+In all approaches, the underlying neural network (i.e., the \ac{MLP} and the \ac{GNN}) are trained %with the same parameters of the proposed approach 
+with a single, stationary phenomenon.
 %
@@
 Moreover, we increase the batch size to 512 and the replay buffer to 10000 since we record 25 agents' experiences for each step instead of one graph experience.
+\lukas{to make sure: in all three cases we use DQN but with (i) MLP, (ii) standard GNN, (iii) our approach - a field-informed GNN, right?}
 %
@@
 
-Focusing now on the test phase, 
+Focusing now on the test phase, % in~\Cref{fig:test}
  we observe that the field-informed version achieves higher coverage than the other approaches in all scenarios since it shows the ability of our solution to generalize to situations--see \Cref{fig:simulations}.

### 2023-05-17 paper.tex
split fig 5 into two figures
comments from meeting

@@
 i.e.
-do the agents respond to intensity of their measurement.}
+do the agents respond to intensity of their measurement.
+-- YES.
+It is part of the feature set.
+While we only use Guassian distributions but we can use any other distribution and shape.
+}
 As a simulation environment, we use Alchemist~\cite{alchemist}, 
@@
 \subsection{Scenario}
-Figure~\ref{fig:static} illustrates the scenario used during the learning \lukas{is this training the \ac{GNN}? }, 
+Figure~\ref{fig:static} illustrates the scenario used during the training % \lukas{is this training the \ac{GNN}? }, 
 which consists of a set of $\mathcal{P} = 25$ agents placed in a 2D grid large 1000x1000 meters and a single, stationary phenomenon with a 3D Gaussian 
@@
 %
-\lukas{should the following go into section III?}
+
+\lukas{should stay here but this is the explicit implementation of the field:}
 Each agent will produce a computation field with which they will try to approximate the direction of the phenomenon of interest. 
@@
 or is it in a specific layout (e.g.
-lattice) but with some randomness of the actual position?} positioned on a grid without knowing the correct position of the phenomenon,
- but that is fixed in the top right corner.  \lukas{so for all 20,000 experiences, the phenomena is exactly as shown in fig 2a during training?}
+lattice) but with some randomness of the actual position?
+lattice layout but random variation around a fixed point.} positioned on a grid without knowing the correct position of the phenomenon,
+ but that is fixed in the top right corner.
+\lukas{so for all 20,000 experiences, the phenomena is exactly as shown in fig 2a during training?
+--> WE ONLY USE THE SINGLE PHENOMA IN A SINGLE FIXED LOCATION BUT WITH VARYING POSITIONS OF AGENTS!!}
  The feature set used by the \ac{GNN} created for each node consists of the vector computed by the aggregated program and the value of the local sensor $\zeta$.
@@
 As the reward function is defined as a regret, 
- we decided to use the Huber loss function~\cite{??}, 
+ we decided to use the Huber loss function, %~\cite{??}, 
  which is a combination of the $L_1$ and $L_2$ loss functions.
@@
 We generated random scenarios for each type of scenario. 
+\lukas{64 experiments per mode with random initial positions of the phenomena. Moving: starting at random position and move in a straight line moving with 5m/s in an unbounded environment.}
 %
@@
 
-\lukas{I would move this up into IV.A}
+\lukas{I would move this up into IV.A --> emphasise that we only used a single static phenomena in the training therefore this section is improtant!}
 
@@
  which is challenging to capture with \acp{MLP} combined with \ac{DQN}, even if we use field information.
+ \lukas{add short discussion on 'new' figure 6}
 \begin{figure}
@@
   \end{subfigure}
+	\caption{Quantative test results in the three types of scenarios.
+	 	We can see that the proposed approach is able to cover and track the phenomenon better than the baselines.
+	}
+	\label{fig:test}
+\end{figure*}
+
+\begin{figure*}
+	\centering
   	\begin{subfigure}[b]{0.85\linewidth}
@@
   \end{subfigure} 
-  \caption{Quantative test results in the three types of scenarios.
-  We can see that the proposed approach is able to cover and track the phenomenon better than the baselines.
-  }
-  \label{fig:test}
+  \caption{Coverage of two zones using the different modes of the controller.}
+  \label{fig:resCoverage}
 \end{figure*}
+
 \section{Conclusion and Future Work}

### 2023-05-17 paper.tex
chore: minor
chore: some comments on fig 6

@@
 %
+Indeed, observing the \Cref{fig:resCoverage}, we notice that the informed version maintains a balanced coverage between the two zones, 
+ with a difference of less than 5\% between the two parts, maintaining a fair division of the phenomenon.
+%
 In contrast, the uninformed version also maintains a fair division but with significantly different coverage between the two parts, 
- indicating an imbalance among the nodes in one of the zones.
+ indicating a wrong placement among the nodes in one of the zones.
+%
+This is a consequence of the uninformed version's inability to encode the necessary information to divide the agents into two zones, therefore it is not able to generalise.
 %
@@
   The agents go towards the phenomenon and try to cover it without collapsing into a single point.
-  the grey area represents the area covered by the robots (last figure on the right).}
+  the grey area represents the area covered by the agents (last figure on the right).}
   \label{fig:simulations}

### 2023-05-17 paper.tex
partially handle fig 6

@@
 The uninformed GNN version, however, 
- fails to generalize its movement behaviour, as evidenced by the simulations where the agents, once reaching the target zone, stop moving due to tracking issues.
+ fails again to generalize its movement behaviour, as evidenced by the simulations where the agents, once reaching the target zone, stop moving due to tracking issues.
 

### 2023-05-18 paper.tex
chore: minor on intro

@@
 Coordinating and organising this information exchange within the network of agents is required for improve the collaboration among them further~\cite{jennings1996coordination}.
-Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complementary competences of the individual agents, or (iii) the collective resources of multiple agents.
+Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complementary competencies of the individual agents, or (iii) the collective resources of multiple agents.
 

### 2023-05-18 paper.tex
remove page limit descriptions

--- 2023-05-17 paper.tex
@@
 \acrodef{DQN}[DQN]{Deep Q-Learning}
-\ga{Page limit: 10 pages (include refs)}
+%\ga{Page limit: 10 pages (include refs)}
 
@@
 Finally, we present a formalization of the problem and a discussion of the motivations behind our proposed work.
-\ga{Page budget: 1/2 page \\}
+%\ga{Page budget: 1/2 page \\}
 %\ga{Plan: discussion about the problem of coordination in multi-agent systems and the need for a scalable approach.
@@
 this is used in the use case to accumulate the distance from the phenomena that is tracked
-\lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
+\lukas{Lukas to review this section again.}
 \subsection{Graph Neural Networks}
@@
 \label{sec:approach}
-\ga{Page budget: 2/3 pages \\}
+%\ga{Page budget: 2/3 pages \\}
 %\ga{Plan: we should discuss the approach in detail, including the system model, and the learning dynamics.
@@
 \end{figure}
-\ga{Page budget: 2 pages \\}
+%\ga{Page budget: 2 pages \\}
 %\ga{Plan: we should discuss the robot aggregation scenario, specifying the state space, the action space, and the reward function.
@@
   Finally, the phenomenon may have varying areas of interest, defined by an underlying distribution function.
-\lukas{are we utiliising this in any way?
-i.e.
-do the agents respond to intensity of their measurement.
--- YES.
-It is part of the feature set.
+This underlying distribution is utilised in the feature set of each agent's observation and guides the agents to rally over the phenomenon.
 While we only use Guassian distributions but we can use any other distribution and shape.
-}
 As a simulation environment, we use Alchemist~\cite{alchemist}, 
@@
  which is a library for deep learning on graphs built on top of PyTorch~\cite{torch}.
-Finally, we use ScaFi~\cite{casadei2022scafi} as the aggregate programming language to support 
- our field-informed approach.
+Finally, we use ScaFi~\cite{casadei2022scafi} as the aggregate programming language to support our field-informed approach.
+The evaluation is performed in two stages, first the neural networks are trained in an explicit training phase before extensively evaluated in the testing stage.
 
 \subsection{Scenario}
-Figure~\ref{fig:static} illustrates the scenario used during the training % \lukas{is this training the \ac{GNN}? }, 
-which consists of a set of $\mathcal{P} = 25$ agents placed in a 2D grid large 1000x1000 meters and a single, stationary phenomenon with a 3D Gaussian 
+Figure~\ref{fig:scenarios} presents the three different types of scenarios utilised within the evaluation.
+The first type of experiments considers a single phenomenon at a static location (i.e., \emph{Zone Fixed}), the second type of experiment considers two phenomena in two independent but static locations (i.e., \emph{Two Zones}), and the third type of experiment considers a moving phenomenon (i.e., \emph{Moving}).
+All phenomena are modelled as a Gaussian distribution.
+Importantly, only the left scenario illustrated in Figure~\ref{fig:static} was used for training the neural networks.
+Furthermore, all three types of scenarios contain a set of $\mathcal{P} = 25$ agents placed in a 2D grid large 1000x1000 meters. % and a single, stationary phenomenon with a  
 Each agent can perceive the presence
  of the phenomenon of interest through an installed sensor $\zeta_v$ with $v \in \mathbb{V}$. 
- (e.g., camera, temperature sensor, etc.). 
+ (e.g., camera, temperature sensor, etc.) if it is within range.
 %
@@
  which in this case depends on a $\mathbb{O}$ range fixed to 300 meters. %\lukas{$\mathbb{O}$  not defined.}
-From the neighbourhood, 
- each agent can also perceive the direction of the other agents.
+Through this communication channel, agents can exchange information. 
+%From the neighbourhood, each agent can also perceive the direction of the other agents.
 %
@@
 In particular, the angles are quantized to 20 degrees, and for the velocities, we have selected [0, 5, 10] m/s.
-%
-
-\lukas{should stay here but this is the explicit implementation of the field:}
-Each agent will produce a computation field with which they will try to approximate the direction of the phenomenon of interest. 
-%
-The program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
+%\lukas{is this training the \ac{GNN}? }, 
+%
+
+For the aggregated information, each agent will produce a computation field with which they will try to approximate the direction of the phenomenon of interest. 
+%
+The specific program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
 This can be expressed in Scafi as follows:
@@
 \subsection{Training Phase}
-The learning process was divided into 100 episodes, each consisting of 200 steps, resulting in a total of 20,000 experiences.
+Before we can evaluate our approach, the underlying neural networks have to be fine-tuned in a dedicated training phase.
+The training process for each neural network was divided into 100 episodes, each consisting of 200 steps, resulting in a total of 20,000 experiences.
 %
 For each episode, 
- the 25 agents are semi-randomly \lukas{what is semi-random?
-are the position fixed but the agents random?
-or is it in a specific layout (e.g.
-lattice) but with some randomness of the actual position?
-lattice layout but random variation around a fixed point.} positioned on a grid without knowing the correct position of the phenomenon,
- but that is fixed in the top right corner.
-\lukas{so for all 20,000 experiences, the phenomena is exactly as shown in fig 2a during training?
+ the 25 agents are semi-randomly positioned on a grid (i.e., in a lattice layout with a random variation in their position) without knowing the correct position of the phenomenon, 
+ but that is fixed in the top right corner. The position of the phenomenon with an example of positioned agents can be seen in Figure~\ref{fig:static}. 
+% \lukas{so for all 20,000 experiences, the phenomena is exactly as shown in fig 2a during training?
 --> WE ONLY USE THE SINGLE PHENOMA IN A SINGLE FIXED LOCATION BUT WITH VARYING POSITIONS OF AGENTS!!}
@@
   \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
-  \delta (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise.} 
+  \delta (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise,} 
  \end{cases}
 \end{equation*}
-This function is used to penalise the agent if the action taken is too far from the optimal action. \lukas{whats $y, \hat{y}$ and $\delta$?}
+where $y$ is the expected location, $\hat{y}$ represents the ground truth, and $\delta$ is a dedicated threshold. \lukas{what is the value of $\delta$ in the training?}
+\lukas{also: $\delta$ is used as a function and as a value?}
 %
@@
 For the evaluation, 
- we explore three different types of scenarios.
-%
-The first type considers only a single, stationary phenomenon like the one used in the training phase. 
-%
-The second type considers multiple phenomena, requiring the agents to make a decision about which phenomena to cover. 
-%
-Finally, the third type considers a single moving phenomenon which the agents have to follow throughout the simulation. 
-%
-In all three types of scenarios, the shape and underlying distribution are initially unknown to the agents but remain constant for the entire duration of the simulation. 
-%
-We generated random scenarios for each type of scenario. 
-\lukas{64 experiments per mode with random initial positions of the phenomena. Moving: starting at random position and move in a straight line moving with 5m/s in an unbounded environment.}
-%
-Examples of such types are shown in~\Cref{fig:scenarios}. %\lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?}
-
-\lukas{I would move this up into IV.A --> emphasise that we only used a single static phenomena in the training therefore this section is improtant!}
+ we explore the previously discussed three different types of experiments.
+%%
+%The first type considers only a single, stationary phenomenon like the one used in the training phase. 
+%%
+%The second type considers multiple phenomena, requiring the agents to make a decision about which phenomena to cover. 
+%%
+%Finally, the third type considers a single moving phenomenon which the agents have to follow throughout the simulation. 
+%%
+%In all three types of scenarios, the shape and underlying distribution are initially unknown to the agents but remain constant for the entire duration of the simulation. 
+%
+We generated 64 random scenarios for each type of experiment. Additionally, the placement of the agents were randomised as it has been done during training.
+For the first type, considering a single static the phenomena is randomly placed in the environment. This is in contrast to the training where the phenomena was always placed in the same location.  
+For the second type, we placed two distinct phenomena within the area.
+There location is kept constant on all 64 experiments.
+As the training only contained a single phenomena, this setup represents a challenge for the agents.
+Finally, the third type contained a moving phenomena.
+In each scenario, the starting position as well as their direction of movement is randomly sampled from a uniform distribution.
+The movement is in a straight line with a constant speed of 5m/s within an unbounded environment.
+%to ensure the agents can keep up with the phenomenon. 
+%\lukas{64 experiments per mode with random initial positions of the phenomena. Moving: starting at random position and move in a straight line moving with 5m/s in an unbounded environment.}
+%
+Examples of all three types of experiment are shown in~\Cref{fig:scenarios}.
+%\lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?}
+
+%\lukas{I would move this up into IV.A --> emphasise that we only used a single static phenomena in the training therefore this section is improtant!}
 
 \subsection{Baselines}
-We compare our field-informed reinforcement learning approach against a baseline approach utilising a \ac{MLP} as well as an approach only relying on \acp{GNN}. 
+We compare our field-informed reinforcement learning approach against a baseline approaches where the \ac{DQN} utilises a \ac{MLP} as well as an approach only relying on \acp{GNN}, without additional field information.
 %
@@
 %
-The \ac{MLP} uses the same feature set as the \ac{GNN} in the proposed approach, but differently, it uses a standard DQN approach without leveraging the graph structure.
-Moreover, we increase the batch size to 512 and the replay buffer to 10000 since we record 25 agents' experiences for each step instead of one graph experience.
-\lukas{to make sure: in all three cases we use DQN but with (i) MLP, (ii) standard GNN, (iii) our approach - a field-informed GNN, right?}
-%
-The \ac{GNN} instead of using the field information, use directly the position of nodes and the local sensor value as input features.
+
+The \ac{MLP} uses the same feature set as the \ac{GNN} but applies it in the \ac{DQN} but 
+%in the proposed approach, but differently, it uses a standard DQN approach 
+without leveraging the graph structure. Moreover, we increase the batch size to 512 and the replay buffer to 10000 since we record 25 agents' experiences for each step instead of one graph experience.
+%\lukas{to make sure: in all three cases we use DQN but with (i) MLP, (ii) standard GNN, (iii) our approach - a field-informed GNN, right?}
+%
+%The \ac{GNN} instead of using the field information, use directly the position of nodes and the local sensor value as input features.
+The \ac{GNN} alone, without using the field information, apply the position of nodes and the local sensor value directly as input features within the \ac{DQN}.
 %
@@
 %
-Indeed, observing the \Cref{fig:resCoverage}, we notice that the informed version maintains a balanced coverage between the two zones, 
- with a difference of less than 5\% between the two parts, maintaining a fair division of the phenomenon.
-%
 In contrast, the uninformed version also maintains a fair division but with significantly different coverage between the two parts, 
- indicating a wrong placement among the nodes in one of the zones.
-%
-This is a consequence of the uninformed version's inability to encode the necessary information to divide the agents into two zones, therefore it is not able to generalise.
+ indicating an imbalance among the nodes in one of the zones.
 %
@@
   The agents go towards the phenomenon and try to cover it without collapsing into a single point.
-  the grey area represents the area covered by the agents (last figure on the right).}
+  the grey area represents the area covered by the robots (last figure on the right).}
   \label{fig:simulations}
@@
 In addition, we also plan to take the approach to modern actor-critical solutions, which are better suited to modern swarm robotics problems because of the continuous action space.
-\ga{Page budget: 0.5}
+%\ga{Page budget: 0.5}
 \label{sec:conclusion}

### 2023-05-18 paper.tex
chore: some minors

--- 2023-05-18 paper.tex
@@
 Autonomous agents interacting and operating towards a common task, rely on shared information and consequently emerging knowledge in order to achieve the underlying goals efficiently. 
+%
 Coordinating and organising this information exchange within the network of agents is required for improve the collaboration among them further~\cite{jennings1996coordination}.
+%
 Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complementary competencies of the individual agents, or (iii) the collective resources of multiple agents.
@@
 
-In this paper, we propose a novel approach utilising aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
-First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent.
-Second, we utilise a \ac{GNN} within the \ac{DQN}.
+In this paper, 
+we propose a novel approach utilising aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
+ First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent. Second, we utilise a \ac{GNN} within the \ac{DQN}. 
  Here the \ac{GNN} provides node embeddings for each agent, serving as input for the \ac{DQN}. 
-The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.'
+ The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.
 %We further combine this with reinforcement learning techniques to respond to changing conditions in the environment. 
@@
 This article studies intelligent \emph{collective behaviours} within the context of \emph{large-scale} distributed systems.
- Specifically, we focus on \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}, i.e.,  systems consisting of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator} able to interact with the real world.
+ Specifically, we focus on \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}, 
+ i.e.,  systems consisting of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator} able to interact with the real world.
  Each of these nodes can interact with its direct neighbours, either based on physical (e.g. communication range) or logical distance (e.g. 1-hop neighbourhood).
@@
 %
-Each node is governed by an autonomous software agent, executing a \emph{local} control loop.
+Each node is governed by an autonomous software agent, 
+ executing a \emph{local} control loop. 
  In each iteration of this loop, the agent can access the information available within its own \emph{context}. 
-This context is comprised by information acquired by the node directly through its sensors and the information received from its neighbours.
-%
-At the end of each control loop iteration, the node can send messages to its neighbourhood. This message may contain raw sensor data or aggregated information over time or from other neighbours.
+ This context is comprised of information acquired by the node directly through its sensors and the information received from its neighbours. 
+%
+At the end of each control loop iteration, 
+ the node can send messages to its neighbourhood. 
+ This message may contain raw sensor data or aggregated information over time or from other neighbours.
 % about its current evaluation and update its internal state.
@@
 %This general model can be adapted to several different scenarios, such as a network of sensors, a network of robots (e.g., swarm robotics), or a network of mobile phones.
-Examples of such large-scale system can be a network of robots (e.g., swarm robotics), camera and IoT networks, or a network of mobile phones.
+Examples of such large-scale systems can be a network of robots (e.g., swarm robotics), camera and IoT networks, or a network of mobile phones.
 %
@@
  the system to achieve a certain \emph{collective} requirements through \emph{cooperation}, 
- such as spatial area coverage, phenomena tracking, and node aggregation.
+ such as spatial area coverage, phenomena tracking, and robot aggregation~\cite{DBLP:journals/firai/SchranzUSE20}.
 %
@@
 %
-The foster is structurally similar to \emph{swarm-like systems}, 
+The system model is structurally similar to \emph{swarm-like systems}, 
  as it is composed of a set of \emph{agents} 
@@
 Field calculus includes main operators for expressing \emph{spatio-temporal} compotations that allow for 
-\emph{i)} the progression of values over time, achieved by transforming a value computed in a previous round into a new value; 
-\emph{ii)} the exchange of data with neighbouring fields, where received data is validated by adjacent fields; and 
+\emph{i)} the progression of values over time, achieved by transforming a field computed in a previous round into a new field; 
+\emph{ii)} the exchange of data with neighbouring fields, where received data is accessed by adjacent fields; and 
 \emph{iii)} the conditional division of computation into distinct domains of collective computation. 
 %
-We avoid presenting here the full details of these operators since they are not essential to the contribution. 
+%We avoid presenting here the full details of these operators since they are not essential to the contribution. 
 For further elaboration on the actual field calculus, please refer to the work of~\cite{DBLP:journals/jlap/ViroliBDACP19}.
@@
 In other words, it maps a Boolean field (where ``true'' indicates the presence of a source node and ``false'' indicates its absence) to a distance field that indicates the proximity to the nearest source.
-In ScaFi (a scala implementation of aggregate computing) it is simply expressed as:
+%
+In ScaFi~\cite{casadei2022scafi} (a scala implementation of aggregate computing) it is simply expressed as:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Along the gradient, it is possible to cast information within the system, accumulating value during the programming. This pattern is called \emph{gradient-cast} (or \emph{G}) 
+Along the gradient, it is possible to cast information within the system, accumulating value during the field expansion. 
+This pattern is called \emph{gradient-cast} (or \emph{G}) and is a fundamental building block for the coordination of swarm-like systems.
+It is expressed as: 
 \begin{lstlisting}
@@
 \end{lstlisting}
-\ga{more description here}
-this is used in the use case to accumulate the distance from the phenomena that is tracked
+%\ga{more description here}
+where \texttt{source} is a Boolean field that indicates the presence of a source node, 
+ \texttt{value} is the value to be cast, and 
+ \texttt{acc} is the accumulation function that is applied to the value during the cast.
+
 \lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}

### 2023-05-18 paper.tex
chore: more on evaluation

@@
 \emph{i)} the progression of values over time, achieved by transforming a field computed in a previous round into a new field; 
-\emph{ii)} the exchange of data with neighbouring fields, where received data is accessed by adjacent fields; and 
+\emph{ii)} the exchange of data with neighbouring fields (typically referred as \lstinline|nbr|), where received data is accessed by adjacent fields; and 
 \emph{iii)} the conditional division of computation into distinct domains of collective computation. 
@@
  The node embedding $h_v$ describes the node in the network and summarises the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
+%
 In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $\mathcal{N}_G(v)$,
  and then combining it with the node's current embedding $h_v$.
- The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased
+%
+The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. 
+ With each traversal of an edge $e \in E$ from $v$, the number of layers increased
 Formally, a \ac{GNN} can be defined by two phases:
@@
 $\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
-  and it could be a simple sum, max or sum of products. More complex aggregation functions are available~\cite{pellegrini2020learning}.
+ and it could be a simple sum, max or sum of products. 
+ More complex aggregation functions are available~\cite{pellegrini2020learning}.
 %
@@
 \end{equation} 
-This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
-\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}
+This formulation allows \acp{GNN} to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
+%\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}
 
 \acp{GNN} are used in several application areas such as social network analysis, chemistry, and physics.
-In this paper, we use GNNs to learn a local behaviour for each agent in a multi-agent system (more details in Section~\ref{sec:approach}).
+ In this paper, we use \acp{GNN} to learn a local behaviour for each agent in a multi-agent system (more details in Section~\ref{sec:approach}).
 \subsection{Many-Agent Reinforcement Learning}
@@
 In particular, in this work, we consider homogenous \ac{MAARL}~\cite{yang2021many}, 
- where the set of agents $N \gg 2$ and each agent is interchangeable and indistinguishable.
+ where the set of agents is large ($N \gg 2$) and each agent is interchangeable and indistinguishable.
 %
@@
 In this section, we discuss the components involved in our proposed solution (i.e., architecture) and how these components interact with each other to bring the system to perform the collective behaviour (i.e dynamics).
-Finally, we will detail the learning algorithm designed and used to synthesise the policy (i.e., learning algorithm).
+Finally, we will detail the learning algorithm designed and used to synthesise the policy.
 \subsection{Architecture and System dynamics}
@@
  producing an action $a^v_t$ that will then modify the global state of the system.
- \ga{todo: put pseudo code to express this dynamics}
+ %\ga{todo: put pseudo code to express this dynamics}
 %
@@
 %
-The program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
+The program $\Gamma$ in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
 This can be expressed in Scafi as follows:
@@
 %
-this value will then be fed into the \ac{GNN} to compute the action to be performed.
+This value will then be fed into the $\pi_{GNN}$ to compute the action to be performed.
  %\lukas{not sure if this paragraph should be its own subsection. or at least it needs  different title.}

### 2023-05-18 paper.tex
chore: on huber loss

--- 2023-05-18 paper.tex
@@
 Field calculus includes main operators for expressing \emph{spatio-temporal} compotations that allow for 
-\emph{i)} the progression of values over time, achieved by transforming a value computed in a previous round into a new value; 
-\emph{ii)} the exchange of data with neighbouring fields, where received data is validated by adjacent fields; and 
+\emph{i)} the progression of values over time, achieved by transforming a field computed in a previous round into a new field; 
+\emph{ii)} the exchange of data with neighbouring fields (typically referred as \lstinline|nbr|), where received data is accessed by adjacent fields; and 
 \emph{iii)} the conditional division of computation into distinct domains of collective computation. 
 %
-We avoid presenting here the full details of these operators since they are not essential to the contribution. 
+%We avoid presenting here the full details of these operators since they are not essential to the contribution. 
 For further elaboration on the actual field calculus, please refer to the work of~\cite{DBLP:journals/jlap/ViroliBDACP19}.
@@
 \end{lstlisting}
-\ga{more description here}
-this is used in the use case to accumulate the distance from the phenomena that is tracked
+%\ga{more description here}
+where \texttt{source} is a Boolean field that indicates the presence of a source node, 
+ \texttt{value} is the value to be cast, and 
+ \texttt{acc} is the accumulation function that is applied to the value during the cast.
+
 \lukas{Lukas to review this section again.}
@@
 %
-The specific program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
+
+%The specific program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
+The program $\Gamma$ in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
 This can be expressed in Scafi as follows:

### 2023-05-18 paper.tex

@@
 \begin{equation*}
-L = \begin{cases}
+L_\delta = \begin{cases}
   \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
@@
 \end{equation*}
-where $y$ is the expected location, $\hat{y}$ represents the ground truth, and $\delta$ is a dedicated threshold. \lukas{what is the value of $\delta$ in the training?}
+where $\delta$ is a hyperparameter that determines the threshold between the two loss functions and $y$ and $\hat{y}$ are the target and predicted values, respectively.
+This function is used to penalise the agent if the action taken is too far from the optimal action. 
+\lukas{what is the value of $\delta$ in the training?}
 \lukas{also: $\delta$ is used as a function and as a value?}

### 2023-05-18 paper.tex
chore: add bib

@@
 %
-Formally, first, we define the overall area covered by the agents as:
+Formally, first, we define the overall coverage for a certain time step as:
 \begin{equation*}
 \Omega = \bigcup_{v \in V} \omega_v
-\end{equation*}
-
-Finally, we define the coverage as:
-\begin{equation*}
+\quad 
 C = \frac{|\Omega \cap \mathcal{P}|}{|\mathcal{P}|}
 \end{equation*}
-where $\mathcal{P}$ is the area of the phenomenon.
-This is computed for each step of the simulation. 
+where $\mathcal{P}$ is the area of the phenomenon and $\Omega$ is the area covered by the agents.
+%This is computed for each step of the simulation. 
 In the training phases, 

### 2023-05-18 paper.tex
chore: minor
reply to lukas

@@
   \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
-  \delta (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise,} 
+  \delta \cdot (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise,} 
  \end{cases}
 \end{equation*}
-where $\delta$ is a hyperparameter that determines the threshold between the two loss functions and $y$ and $\hat{y}$ are the target and predicted values, respectively.
+where $\delta$ is a hyperparameter that determines the threshold between the two loss functions (fixed to 1 in our experiemnt) and $y$ and $\hat{y}$ are the target and predicted values, respectively.
 This function is used to penalise the agent if the action taken is too far from the optimal action. 
-\lukas{what is the value of $\delta$ in the training?}
-\lukas{also: $\delta$ is used as a function and as a value?}
+%\lukas{what is the value of $\delta$ in the training?}
+%\lukas{also: $\delta$ is used as a function and as a value?}
 %
@@
 %
-Examples of all three types of experiment are shown in~\Cref{fig:scenarios}.
+Examples of all three types of experiments are shown in~\Cref{fig:scenarios}.
 %\lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?}

### 2023-05-18 paper.tex
feat: try to rewrite the intro

@@
 
-Autonomous agents interacting and operating towards a common task, rely on shared information and consequently emerging knowledge in order to achieve the underlying goals efficiently. 
-%
-Coordinating and organising this information exchange within the network of agents is required for improve the collaboration among them further~\cite{jennings1996coordination}.
-%
-Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complementary competencies of the individual agents, or (iii) the collective resources of multiple agents.
-
-Various phenomena in the real world are not evenly distributed across the environment such as watershed, wild fires, traffic, human crowd movement, or wild life movement from insects, fish, birds, and mammals.
-To acquire appropriate information about such phenomena, we also need to distribute the sensors accordingly to cover all aspects sufficiently.
+%Autonomous agents interacting and operating towards a common task, rely on shared information and consequently emerging knowledge in order to achieve the underlying goals efficiently. 
+%
+%Coordinating and organising this information exchange within the network of agents is required for improve the collaboration among them further~\cite{jennings1996coordination}.
+%
+%Through this coordination, agents may overcome problems requiring (i) efficiency and accuracy only achieved by multiple agents, (ii) the complementary competencies of the individual agents, or (iii) the collective resources of multiple agents.
+
+%Various phenomena in the real world are not evenly distributed across the environment such as watershed, wild fires, traffic, human crowd movement, or wild life movement from insects, fish, birds, and mammals.
+%To acquire appropriate information about such phenomena, we also need to distribute the sensors accordingly to cover all aspects sufficiently.
 While manual deployment will mitigate the problem, not all phenomena are known exactly \emph{a priori} and would require maintenance and adjustment during runtime.
@@
 
-This leads to several questions 
-\begin{enumerate}
-	\item How to distribute sensors to ensure good attainment of information?
-	\item How to maintain knowledge about the phenomena and unlearn irrelevant information?
-	\item How to move sensors when the phenomena moves to ensure good coverage?
-\end{enumerate} \lukas{refine and rework these questions}
-
-In this paper, 
-we propose a novel approach utilising aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
+%This leads to several questions 
+%\begin{enumerate}
+%	\item How to distribute sensors to ensure good attainment of information?
+%	\item How to maintain knowledge about the phenomena and unlearn irrelevant information?
+%	\item How to move sensors when the phenomena moves to ensure good coverage?
+%\end{enumerate} \lukas{refine and rework these questions}
+The coordination of a group of autonomous agents that can perceive and act in their environment is a fundamental problem in artificial intelligence. 
+Such agents need to cooperate and communicate in order to achieve a common goal, 
+ while dealing with the challenges of distributed and situated intelligence. 
+ These challenges include the limited and local nature of the information available to each agent and the emergence of global behaviour from local interactions. 
+%
+One domain where these challenges are particularly evident is Cyber-Physical Swarms (CPSs, or \emph{swarm-like systems}), 
+ where a large number of agents interact with each other and their environment based on spatial proximity. 
+%
+Examples of CPSs include swarm robotics, smart cities, sensor networks, and social systems. 
+ In these systems, agents need to adapt to dynamic and uncertain situations, while ensuring the achievement of global objectives that may not be directly observable or measurable by individual agents.
+
+A key question in designing swarm-lie systems is how to create distributed controllers for the agents that can enable them to perform complex collective tasks. 
+ Two main approaches have been proposed for this problem: manual design and automatic design. 
+%
+In manual design, programmers build the controllers directly, 
+ using domain knowledge and programming languages or frameworks that support distributed computation and communication like macro-programming approaches~\cite{DBLP:journals/corr/abs-2201-03473}. 
+%
+In automatic design, machine learning methods like \ac{MARL} and evolutionary algorithms are used to synthesize programs or policies for the agents from the high-level task or goal descriptions.
+%
+Both approaches have advantages and disadvantages. 
+ Manual design allows programmers to specify desired properties for the system declaratively. However, manual design can be tedious and error-prone.
+%
+Automatic design can overcome these limitations by learning from data and experience. 
+ However, automatic design can also suffer from several challenges, such as learning the right representation for agent communication.
+%
+
+In this paper, we propose a novel hybrid approach that combines manual design and automatic design of distributed controllers.
+This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
+Specifically, we propose a solution called field-informed reinforcement learning (FIRL) that utilizes aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
  First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent. Second, we utilise a \ac{GNN} within the \ac{DQN}. 
@@
 %
-This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
+The main contribution of FIRL is that it allows us to create distributed controllers that are informed by a collective knowledge that has been distilled during learning, but that use only local information at runtime.
+This way, FIRL can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
+%% Perhaps to much detail 
+FIRL can also be seen as a way of bridging the gap between symbolic and sub-symbolic AI methods, by integrating declarative programming with deep learning.
+
 %Specifically, 

### 2023-05-18 paper.tex
minor edits

--- 2023-05-18 paper.tex
@@
 
-Focusing now on the test phase, % in~\Cref{fig:test}
- we observe that the field-informed version achieves higher coverage than the other approaches in all scenarios since it shows the ability of our solution to generalize to situations--see \Cref{fig:simulations}.
+Focusing now on the results of the test phase, highlighted in  \Cref{fig:test},
+ we observe that the field-informed version achieves higher coverage than the other approaches in all scenarios since it shows the ability of our solution to generalize to situations. 
 % 
@@
 
-The \ac{GNN} versions demonstrate the capability to transfer the learned experience to the test phase whereas the MLP version fails to generalize. 
-%
-It is worth noting that, in the static configuration, 
+For all experiments, both \ac{GNN} versions demonstrate the capability to transfer the learned experience to the test phase whereas the MLP version fails to generalize. 
+%
+It is worth noting that, in the \emph{Zone Fixed} experiment, % (left graph in \Cref{fig:test}), 
  once the desired configuration is achieved in the static case, 
@@
 %
-Interesting observations arise when we use different scenarios from the training phase. 
- Starting from a division into two zones, 
- we notice that even in this case, the informed version finds a better configuration than the GNN counterpart. 
+Interesting observations arise when we use scenarios different from the training phase. 
+% Starting from a division into two zones, 
+In the \emph{Two Zones} experiment,
+ we notice that our approach using field-information finds a better configuration than the simple GNN counterpart. 
 %
@@
 Indeed, observing the \Cref{fig:resCoverage}, we notice that the informed version maintains a balanced coverage between the two zones, 
- with a difference of less than 5\% between the two parts, maintaining a fair division of the phenomenon.
+ with a difference of less than 5\% between the two parts, maintaining a fair division of the phenomena.
 %
@@
 %
-Finally, the moving scenario emphasizes how the informed version generates a more robust policy for new scenarios. 
- Indeed, we observe that the informed version maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
+Finally, the \emph{Moving} experiment emphasizes how the informed version generates a more robust policy for new scenarios. 
+ Indeed, we observe that the our approach using field-informed reinforcement learning maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
 %
@@
  which is challenging to capture with \acp{MLP} combined with \ac{DQN}, even if we use field information.
- \lukas{add short discussion on 'new' figure 6}
+% \lukas{add short discussion on 'new' figure 6}
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
-    \caption{Coverage in the three types of scenarios}
+    \caption{Ratio of coverage of the phenomena in the three types of experiments. Our field-informed reinforcement learning approach can outperform other approaches lacking field information.}
     \label{fig:coverage-test}
@@
     \includegraphics[width=\linewidth]{imgs/inside-test}
-    \caption{Number of agents inside the phenomenon in the three types of scenarios}
+    \caption{Number of agents inside the phenomenon in the three types of experiments}
     \label{fig:inside-test}
@@
     \includegraphics[width=\linewidth]{imgs/inside-two-test}
-    \caption{Two zones scenario: agent inside each phenomenon}
+    \caption{\emph{Two Zones} experiment: aggregated number of agent inside each phenomenon}
     \label{fig:inside-two-test}
@@
     \includegraphics[width=\linewidth]{imgs/coverage-two-test}
-    \caption{Two zones scenario: coverage of each area}
+    \caption{\emph{Two Zones} experiment: ratio of covered to uncovered area both phenomena}
     \label{fig:coverage-two-test}

### 2023-05-18 paper.tex
shortend bib, reworked some parts, intro FIRL as ac, moved fig 4

@@
 \acrodef{DQN}[DQN]{Deep Q-Learning}
+\acrodef{FIRL}[FIRL]{Field-Informed Reinforcement Learning}
 %\ga{Page limit: 10 pages (include refs)}
@@
 %
-In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning}.
+In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning} (FIRL).
 We utilise manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents.
@@
  These challenges include the limited and local nature of the information available to each agent and the emergence of global behaviour from local interactions.
-%
+
 One domain where these challenges are particularly evident is Cyber-Physical Swarms (CPSs, or \emph{swarm-like systems}), 
@@
 %
-Examples of CPSs include swarm robotics, smart cities, sensor networks, and social systems. 
+Examples of CPSs include swarm robotics, smart cities~\cite{bajovic2021marvel}, sensor networks~\cite{pianini2022collective}, and social systems. 
  In these systems, agents need to adapt to dynamic and uncertain situations, while ensuring the achievement of global objectives that may not be directly observable or measurable by individual agents.
 
-A key question in designing swarm-lie systems is how to create distributed controllers for the agents that can enable them to perform complex collective tasks. 
+A key question in designing swarm-like systems is how to create distributed controllers for the agents that enable them to perform complex collective tasks. 
  Two main approaches have been proposed for this problem: manual design and automatic design. 
@@
 
-In this paper, we propose a novel hybrid approach that combines manual design and automatic design of distributed controllers.
+In this paper, we propose a novel hybrid approach that combines manual and automatic design of distributed controllers.
 This follows a novel trend in which high-level declarative programming languages are mixed with machine learning techniques in order to synthesise robust collective controllers~\cite{DBLP:conf/acsos/Aguzzi21,DBLP:conf/icdcs/AguzziCV22,DBLP:conf/coordination/AguzziCV22}.
-Specifically, we propose a solution called field-informed reinforcement learning (FIRL) that utilizes aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
- First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent. Second, we utilise a \ac{GNN} within the \ac{DQN}. 
- Here the \ac{GNN} provides node embeddings for each agent, serving as input for the \ac{DQN}. 
+Specifically, we propose a solution called \ac{FIRL} that utilises aggregate computing~\cite{Beal2015Computer} together with a \ac{GNN}~\cite{Zhou2020AIOpen} in combination with a reinforcement learning approach, namely \ac{DQN}~\cite{mnih2015human}.
+% First, aggregate computing is utilised to collect information from neighbours and create a local knowledge base for each agent. Second, we utilise a \ac{GNN} within the \ac{DQN}. 
+ Here the \ac{GNN} is trained on field-information from aggregate computing and provides so-called node embeddings for each agent, serving as input for the \ac{DQN}. 
  The \ac{DQN} provides then the appropriate actions for the agent to achieve their tasks.
@@
 %
-The main contribution of FIRL is that it allows us to create distributed controllers that are informed by a collective knowledge that has been distilled during learning, but that use only local information at runtime.
-This way, FIRL can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
+The main contribution of \ac{FIRL} is that it allows us to create distributed controllers that are informed by a collective knowledge that has been distilled during training, but that use only local information when deployed.
+This way, \ac{FIRL} can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
 %% Perhaps to much detail 
-FIRL can also be seen as a way of bridging the gap between symbolic and sub-symbolic AI methods, by integrating declarative programming with deep learning.
+\ac{FIRL} can also be seen as a way of bridging the gap between symbolic and sub-symbolic AI methods, by integrating declarative programming with deep learning.
+
+We employ this approach in a swarm-like system setting where agents are tasked to cover phenomena detected in their environment.
+Over time, the agents have to converge over the each phenomenon in order to cover it appropriately as illustrated in \Cref{fig:simulations}.
+\begin{figure}[t]
+	\centering
+	\begin{subfigure}[b]{0.32\linewidth}
+		\includegraphics[width=\textwidth]{imgs/start.png}
+		\caption{Start}
+		\label{fig:initial}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.32\linewidth}
+		\includegraphics[width=\textwidth]{imgs/after.png}
+		\caption{After 50 steps}
+		\label{fig:after}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.32\linewidth}
+		\includegraphics[width=\textwidth]{imgs/end.png}
+		\caption{After 200 steps}
+		\label{fig:end}
+	\end{subfigure}
+	\caption{Qualitative test results of the proposed approach. 
+		The agents go towards the phenomenon and try to cover it without collapsing into a single point.
+		the grey area represents the area covered by the agents (last figure on the right).}
+	\label{fig:simulations}
+\end{figure}
 
@@
 \label{sec:background}
-In this section, we aim to provide a concise overview of the system we intend to address, 
- starting with a high-level system model. 
-%
-We then proceed to discuss relevant research areas that inform our approach. 
-%
-Finally, we present a formalization of the problem and a discussion of the motivations behind our proposed work.
+%In this section, we aim to provide a concise overview of the system we intend to address, 
+% starting with a high-level system model. 
+%%
+%We then proceed to discuss relevant research areas that inform our approach. 
+%%
+%Finally, we present a formalization of the problem and a discussion of the motivations behind our proposed work.
 %\ga{Page budget: 1/2 page \\}
@@
 This article studies intelligent \emph{collective behaviours} within the context of \emph{large-scale} distributed systems.
- Specifically, we focus on \emph{Cyber-Physical Swarms} or \emph{swarm-like systems}, 
- i.e.,  systems consisting of \emph{devices} (or nodes) equipped with \emph{sensors} and \emph{actuator} able to interact with the real world.
- Each of these nodes can interact with its direct neighbours, either based on physical (e.g. communication range) or logical distance (e.g. 1-hop neighbourhood).
-
-%
-Each node is governed by an autonomous software agent, 
- executing a \emph{local} control loop. 
+ Specifically, we focus on \emph{Cyber-Physical Swarms} or \emph{swarm-like systems} consisting of \emph{devices} governed by autonomous software \emph{agents} and equipped with \emph{sensors} and \emph{actuator} able to interact with the real world.
+ Each of these agents can interact with its direct neighbours, either based on physical (e.g. communication range) or logical distance (e.g. 1-hop neighbourhood).
+
+%
+Each agent
+ executes a \emph{local} control loop. 
  In each iteration of this loop, the agent can access the information available within its own \emph{context}. 
- This context is comprised of information acquired by the node directly through its sensors and the information received from its neighbours. 
+ This context is comprised of information acquired by the agent directly through its sensors and the information received from its neighbours. 
 %
 At the end of each control loop iteration, 
- the node can send messages to its neighbourhood. 
+ the agent can send messages to its neighbourhood. 
  This message may contain raw sensor data or aggregated information over time or from other neighbours.
@@
 %
-In other words, it maps a Boolean field (where ``true'' indicates the presence of a source node and ``false'' indicates its absence) to a distance field that indicates the proximity to the nearest source.
-%
-In ScaFi~\cite{casadei2022scafi} (a scala implementation of aggregate computing) it is simply expressed as:
+In other words, it maps a Boolean field (where ``true'' indicates the presence of a source agent and ``false'' indicates its absence) to a distance field that indicates the proximity to the nearest source.
+%
+In ScaFi~\cite{casadei2022scafi} %(a scala implementation of aggregate computing)
+ it is simply expressed as:
 \begin{lstlisting}
@@
 %\ga{more description here}
-where \texttt{source} is a Boolean field that indicates the presence of a source node, 
+where \texttt{source} is a Boolean field indicating the presence of a source agent, 
  \texttt{value} is the value to be cast, and 
@@
 
-\lukas{Lukas to review this section again.}
+%\lukas{Lukas to review this section again.}
 \subsection{Graph Neural Networks}
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{imgs/architecture.pdf}
+	\centering
+  \includegraphics[width=.98\linewidth]{imgs/architecture.pdf}
   \label{fig:architecture}
-  \caption{High-level description of Field-Informed reinforcement learning dynamics. 
+  \caption{High-level description of \ac{FIRL} approach. 
   For each time step $t$, a graph is constructed from the environment, and it will associate each node with a local feature (hexagons in the picture). 
-  Using this feature, an aggregate program computes spatio-temporal information that enhances the feature set of each node, depicted as colours in the middle graph. 
-  Finally, utilizing the GNN, actions are computed for each node in the system to be performed against the environment, enabling advancement in the simulations according to swarMDP rules.
+  Using this feature, an aggregate program computes spatio-temporal information that enhances the feature set of each agent, depicted as colours in the middle graph. 
+  Finally, utilizing the GNN, actions are computed for each agent in the system to be performed against the environment, enabling advancement in the simulations according to swarMDP rules.
   %\lukas{should we have numbers at each step in the figure (and then referring in the caption and text)?}
@@
 Let $\Gamma$ be the aggregate program that takes a graph $G_t$ decorated by $o^v_t$, 
- representing the participating nodes and their neighbourhood relations at time $t$, as input. 
+ representing the participating agents and their neighbourhood relations at time $t$, as input. 
 % \lukas{$G_t$ is not defined. also, not sure what `runs against' means - take $G_t$ as input?}
@@
 %
-The policy $\pi_{gnn}$ is then evaluated for each node using $f_v$ as input, 
+The policy $\pi_{gnn}$ is then evaluated for each agent using $f_v$ as input, 
  producing an action $a^v_t$ that will then modify the global state of the system.
@@
  this is not the case as the information is never aggregated globally. 
-The individual nodes only combine information from their local neighbourhood.
-%
-In fact, the program $\Gamma$ is proactively executed at every node, 
+The individual agents only combine information from their local neighbourhood.
+%
+In fact, the program $\Gamma$ is proactively executed at every agent, 
 and the \ac{GNN} can be locally evaluated using only neighbourhood information. 
@@
 We want to emphasise that, in this case, the \acp{GNN} must be 1-hop; 
-otherwise, they could not have a local interpretation for each node, according to our system model.
+otherwise, they could not have a local interpretation for each agent, according to our system model.
 \subsection{Learning algorithm}
@@
  the use of \acp{GNN} allows us to define policies on a variable neighbourhood, which is essential in such systems as this can change due to the applied neighbourhood policy.
- It is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different nodes~\cite{DBLP:journals/aiopen/ZhouCHZYLWLS20,DBLP:conf/nips/KnyazevTA19}. 
+ It is known that \acp{GNN} have a certain ability to generalise to new structures and scale with different agents~\cite{DBLP:journals/aiopen/ZhouCHZYLWLS20,DBLP:conf/nips/KnyazevTA19}. 
 % 
@@
   }
-  \caption{Deep Q-Network (DQN) with GNN and Graph Replay Buffer}
+  \caption{Deep Q-Network (DQN) with GNN and Graph Replay Buffer executed by each agent}
   \label{alg:dqn}
@@
 %
-This means that the average distance will tend towards the one expressed by the viewing range of each node:
+This means that the average distance will tend towards the one expressed by the viewing range of each agent:
 % 
@@
 --> WE ONLY USE THE SINGLE PHENOMA IN A SINGLE FIXED LOCATION BUT WITH VARYING POSITIONS OF AGENTS!!}
- The feature set used by the \ac{GNN} created for each node consists of the vector computed by the aggregated program and the value of the local sensor $\zeta$.
+ The feature set used by the \ac{GNN} created for each agent consists of the vector computed by the aggregated program and the value of the local sensor $\zeta$.
 %
@@
 \subsection{Baselines}
-We compare our field-informed reinforcement learning approach against a baseline approaches where the \ac{DQN} utilises a \ac{MLP} as well as an approach only relying on \acp{GNN}, without additional field information.
+We compare our \ac{FIRL} approach against a baseline approaches where the \ac{DQN} utilises a \ac{MLP} as well as an approach only relying on \acp{GNN}, without additional field information. 
 %
@@
 %The \ac{GNN} instead of using the field information, use directly the position of nodes and the local sensor value as input features.
-The \ac{GNN} alone, without using the field information, apply the position of nodes and the local sensor value directly as input features within the \ac{DQN}.
+The \ac{GNN} alone, without using the field information, apply the position of agents and the local sensor value directly as input features within the \ac{DQN}.
 %
@@
 % 
-We observe that the field-informed version successfully moves the nodes closer to the target phenomenon, 
+We observe that the field-informed version successfully moves the agents closer to the target phenomenon, 
  distributing them evenly without collapsing into a single central point. 
@@
  once the desired configuration is achieved in the static case, 
- the nodes cease to move, maintaining the found configuration.
+ the agents cease to move, maintaining the found configuration.
 %
@@
 In contrast, the uninformed version also maintains a fair division but with significantly different coverage between the two parts, 
- indicating a wrong placement among the nodes in one of the zones.
+ indicating a wrong placement among the agents in one of the zones.
 %
@@
 Finally, the \emph{Moving} experiment emphasizes how the informed version generates a more robust policy for new scenarios. 
- Indeed, we observe that the our approach using field-informed reinforcement learning maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
+ Indeed, we observe that the our approach using \ac{FIRL} maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
 %
@@
   \hfill
-  \begin{subfigure}[b]{0.96\linewidth}
+  \begin{subfigure}[b]{0.87 \linewidth}
     \includegraphics[width=\linewidth]{imgs/coverage-in-time}
@@
   \hfill
-  \begin{subfigure}[b]{0.99\linewidth}
+  \begin{subfigure}[b]{0.9\linewidth}
     \includegraphics[width=\linewidth]{imgs/reward-in-time}
@@
 \end{figure}
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{0.32\linewidth}
-      \includegraphics[width=\textwidth]{imgs/start.png}
-      \caption{Start}
-      \label{fig:initial}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.32\linewidth}
-      \includegraphics[width=\textwidth]{imgs/after.png}
-      \caption{After 50 steps}
-      \label{fig:after}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.32\linewidth}
-      \includegraphics[width=\textwidth]{imgs/end.png}
-      \caption{After 200 steps}
-      \label{fig:end}
-  \end{subfigure}
-  \caption{Qualitative test results of the proposed approach. 
-  The agents go towards the phenomenon and try to cover it without collapsing into a single point.
-  the grey area represents the area covered by the agents (last figure on the right).}
-  \label{fig:simulations}
-\end{figure}
 
@@
     \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
-    \caption{Ratio of coverage of the phenomena in the three types of experiments. Our field-informed reinforcement learning approach can outperform other approaches lacking field information.}
+    \caption{Ratio of coverage of the phenomena in the three types of experiments. Our \ac{FIRL} approach can outperform other approaches lacking field information.}
     \label{fig:coverage-test}
@@
  along with the combination of \ac{DQN} and \ac{GNN} for synthesizing distributed intelligence.
-The proposed \emph{Field-Informed reinforcement learning} approach offers a promising solution to the challenges faced in coordinating multi-agent systems. 
+The proposed \emph{Field-Informed reinforcement learning} (FIRL) approach offers a promising solution to the challenges faced in coordinating multi-agent systems. 
 By combining manual design and machine learning techniques, 

### 2023-05-18 paper.tex
robotic swarms citation

@@
 %
-Examples of CPSs include swarm robotics, smart cities~\cite{bajovic2021marvel}, sensor networks~\cite{pianini2022collective}, and social systems. 
+Examples of CPSs include swarm robotics~\cite{brambilla2013swarm}, smart cities~\cite{bajovic2021marvel}, sensor networks~\cite{pianini2022collective}, and social systems. 
  In these systems, agents need to adapt to dynamic and uncertain situations, while ensuring the achievement of global objectives that may not be directly observable or measurable by individual agents.

### 2023-05-18 paper.tex
cite for social systems

--- 2023-05-18 paper.tex
@@
 %
-Examples of CPSs include swarm robotics, smart cities~\cite{bajovic2021marvel}, sensor networks~\cite{pianini2022collective}, and social systems. 
+Examples of CPSs include swarm robotics~\cite{brambilla2013swarm}, smart cities~\cite{bajovic2021marvel}, sensor networks~\cite{pianini2022collective}, and social systems~\cite{zhou2019cyber}. 
  In these systems, agents need to adapt to dynamic and uncertain situations, while ensuring the achievement of global objectives that may not be directly observable or measurable by individual agents.

### 2023-05-18 paper.tex
abstract and intro ok

@@
 In this paper, we consider a new \emph{hybrid} approach called  \emph{Field-Informed reinforcement learning} (FIRL).
-We utilise manually designed \emph{computational fields} (globally distributed data structures) to manage the interaction between agents.
-Using \emph{Deep Q-learning} in combination with \emph{Graph Neural Networks} we enable the agents to learn the necessary local behaviour automatically to solve collective tasks.
+We utilise manually designed \emph{computational fields} (globally distributed data structures) to manage global agent coordination.
+Then, using \emph{Deep Q-learning} in combination with \emph{Graph Neural Networks} we enable the agents to learn the necessary local behaviour automatically to solve collective tasks, relying on those fields through local perception.
 %
@@
 Both approaches have advantages and disadvantages.
- Manual design allows programmers to specify desired properties for the system declaratively. However, manual design can be tedious and error-prone.
+ Manual design allows programmers to specify desired properties for the system declaratively, however, it can be tedious and error-prone.
 %
@@
 \ac{FIRL} can also be seen as a way of bridging the gap between symbolic and sub-symbolic AI methods, by integrating declarative programming with deep learning.
+%
+Conceptually, \ac{FIRL} leverages the ideas of \emph{behaviour implicit communication} \cite{DBLP:journals/ijaci/CastelfranchiPT10} \cite{DBLP:conf/e4mas/TummoliniCRVO04}, whereby intelligent agents (here trained by DQN) achieve collective goals by learning how to use signs they left in the environment (here made of fields): much like ants collectively rely on pheromones they produce \cite{DBLP:journals/anor/Parunak97}.
 

### 2023-05-19 paper.tex
feat: on motivation incipit

@@
 %
-$h_v{v}^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
+$h_v^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
 %
@@
 \subsection{Motivation}
-The definition of distributed controllers for intelligent systems can be divided into two major approaches: 
- \emph{manual design}, where the application is conceived and described through a programming language, 
- and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesise distributed controllers. 
-%
-Our work lies between these two approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
-
+%The definition of distributed controllers for intelligent systems can be divided into two major approaches: 
+% \emph{manual design}, where the application is conceived and described through a programming language, 
+% and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesise distributed controllers. 
+%
+%Our work lies between these two approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
+As mentioned the above, our work falls between automatic and manual approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
+%
 Compared to earlier works on field-based coordination, 
@@
 %
-As value-based algorithm we relied on \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications (see \Cref{alg:dqn}): 
+As value-based algorithm we relied on \ac{DQN}~\cite{mnih2015playing} %\ga{to introduce in many rl??} 
+with two major modifications (see \Cref{alg:dqn}): 
 \begin{enumerate}

### 2023-05-19 paper.tex
minors

@@
 \usepackage{acronym}
+\usepackage{hyperref}
 \usepackage{amsmath}
@@
   \includegraphics[width=.98\linewidth]{imgs/architecture.pdf}
-  \label{fig:architecture}
+
   \caption{High-level description of \ac{FIRL} approach. 
-  For each time step $t$, a graph is constructed from the environment, and it will associate each node with a local feature (hexagons in the picture). 
-  Using this feature, an aggregate program computes spatio-temporal information that enhances the feature set of each agent, depicted as colours in the middle graph. 
+  For each time step $t$, a graph is constructed from the environment, and it will associate each node with a local feature $o^v_t$ (hexagons in the picture). 
+  Using this feature, an aggregate program computes spatio-temporal information that enhances the feature set of each agent producing $f_v$, depicted as colours in the middle graph. 
   Finally, utilizing the GNN, actions are computed for each agent in the system to be performed against the environment, enabling advancement in the simulations according to swarMDP rules.
@@
   %\lukas{also, i think we can have the AC and GNN box a bit smaller but increase the size of the networks slightly - and put it on column width ot save space}
-}
+}  \label{fig:architecture}
 \end{figure}
@@
 %}
-The proposed solution consists mainly of two parts, 
+The proposed solution, summarised in \Cref{fig:architecture}, consists mainly of two parts, 
 \emph{i)} the aggregate program used to create part of the observation and 
@@
 Finally, we use ScaFi~\cite{casadei2022scafi} as the aggregate programming language to support our field-informed approach.
-The evaluation is performed in two stages, first the neural networks are trained in an explicit training phase before extensively evaluated in the testing stage.
-
+The evaluation is performed in two stages, first the neural networks are trained in an explicit training phase before being extensively evaluated in the testing stage.
+The simulations are publicly available at \url{https://github.com/AggregateComputing/experiment-2023-acsos-field-informed-rl}.
 \subsection{Scenario}
-Figure~\ref{fig:scenarios} presents the three different types of scenarios utilised within the evaluation.
+\Cref{fig:scenarios} presents the three different types of scenarios utilised within the evaluation.
 The first type of experiments considers a single phenomenon at a static location (i.e., \emph{Zone Fixed}), the second type of experiment considers two phenomena in two independent but static locations (i.e., \emph{Two Zones}), and the third type of experiment considers a moving phenomenon (i.e., \emph{Moving}).
@@
 %
-
+%
 %The specific program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
@@
 This will be a measure of how well the agents are tracking the phenomenon.
-\ga{todo improve the description of the coverage}
+%\ga{todo improve the description of the coverage}
 

### 2023-05-19 paper.tex
bib shortened
shrink: merge regret reward in one formula

@@
 The evaluation is performed in two stages, first the neural networks are trained in an explicit training phase before being extensively evaluated in the testing stage.
-The simulations are publicly available at \url{https://github.com/AggregateComputing/experiment-2023-acsos-field-informed-rl}.
+\footnote{The simulations are publicly available at \url{https://github.com/AggregateComputing/experiment-2023-acsos-field-informed-rl}.}
 \subsection{Scenario}
@@
 %\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
-The final reward function for an agent $v$ is defined as:
+The final reward function $R_{v}$ for an agent $v$ is defined as:
 \begin{equation*}
-R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
+R_{v} = (\frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}) - 1
 \end{equation*}
@@
   as it is a more general measure of the quality of the action taken by the agent.
-It is simplify evaluated as:
-\begin{equation*}
-R^e_{v} = R_v - 1
-\end{equation*}
 %

### 2023-05-19 paper.tex
shrink: merge formula
shrink: remove huber loss description

@@
 In our specific case, the agents can only interact with 1-hop neighbourhoods and are not directly influenced by other agent observations. We can therefore restrict the observation model as follows:
-\begin{equation}
+\begin{equation*}
 \xi(v): \{s_j, j \in \mathcal{N}^v\} \rightarrow O
-\end{equation}
-\begin{equation}
+\qquad
 \xi = \{\xi(v), v \in \mathcal{P}\}
-\end{equation}
+\end{equation*}
 where $\mathcal{N}^v$ is the set of neighbours of $v$.
@@
 Specifically, starting from a global state $\mathcal{S}^P_t$, the next state $\mathcal{S}^P_{t+1}$ is defined as:
-\begin{equation}
+\begin{equation*}
 \mathcal{A}^P_t = \pi(\xi(S^P_t))
-\end{equation}
-\begin{equation}
+\qquad
 \mathcal{S}^P_{t+1} = \mathcal{T}(\mathcal{S}^P_t, \mathcal{A}^P_t)
-\end{equation}
+\end{equation*}
 Given a time $t$, the system can be also represented as a graph $G_t = (V_t, E_t)$, 
@@
 As the reward function is defined as a regret, 
- we decided to use the Huber loss function, %~\cite{??}, 
- which is a combination of the $L_1$ and $L_2$ loss functions.
-%
-The loss function is defined as:
-\begin{equation*}
-L_\delta = \begin{cases}
-  \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
-  \delta \cdot (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise,} 
- \end{cases}
-\end{equation*}
-where $\delta$ is a hyperparameter that determines the threshold between the two loss functions (fixed to 1 in our experiemnt) and $y$ and $\hat{y}$ are the target and predicted values, respectively.
+ we decided to use the Huber loss function with $\delta = 1$, %~\cite{??}, 
+ %which is a combination of the $L_1$ and $L_2$ loss functions.
+%
+%The loss function is defined as:
+%\begin{equation*}
+%L_\delta = \begin{cases}
+%  \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
+%  \delta \cdot (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise,} 
+% \end{cases}
+%\end{equation*}
+%where $\delta$ is a hyperparameter that determines the threshold between the two loss functions (fixed to 1 in our experiemnt) and $y$ and $\hat{y}$ are the target and predicted values, respectively.
 This function is used to penalise the agent if the action taken is too far from the optimal action. 

### 2023-05-19 paper.tex
minors

@@
 We employ this approach in a swarm-like system setting where agents are tasked to cover phenomena detected in their environment.
-Over time, the agents have to converge over the each phenomenon in order to cover it appropriately as illustrated in \Cref{fig:simulations}.
+Over time, the agents have to converge over each phenomenon in order to cover it appropriately as illustrated in \Cref{fig:simulations}.
 \begin{figure}[t]
@@
 \emph{i)} the progression of values over time, achieved by transforming a field computed in a previous round into a new field; 
-\emph{ii)} the exchange of data with neighbouring fields (typically referred as \lstinline|nbr|), where received data is accessed by adjacent fields; and 
+\emph{ii)} the exchange of data with neighbouring fields (typically referred to as \lstinline|nbr|), where received data is accessed by adjacent fields; and 
 \emph{iii)} the conditional division of computation into distinct domains of collective computation. 
@@
 %Our work lies between these two approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
-As mentioned the above, our work falls between automatic and manual approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
+As mentioned above, our work falls between automatic and manual approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
 %
@@
 \end{figure}
-In this section, we discuss the components involved in our proposed solution (i.e., architecture) and how these components interact with each other to bring the system to perform the collective behaviour (i.e dynamics).
+In this section, we discuss the components involved in our proposed solution (i.e., architecture) and how these components interact with each other to bring the system to perform the collective behaviour (i.e., dynamics).
 Finally, we will detail the learning algorithm designed and used to synthesise the policy.
@@
 %
-We generated 64 random scenarios for each type of experiment. Additionally, the placement of the agents were randomised as it has been done during training.
-For the first type, considering a single static the phenomena is randomly placed in the environment. This is in contrast to the training where the phenomena was always placed in the same location.  
+We generated 64 random scenarios for each type of experiment. 
+ Additionally, the placement of the agents was randomised as it has been done during training.
+For the first type, consider a single static phenomenon randomly placed in the environment. 
+This is in contrast to the training where the phenomena were always placed in the same location.  
 For the second type, we placed two distinct phenomena within the area. 
-There location is kept constant on all 64 experiments.
-As the training only contained a single phenomena, this setup represents a challenge for the agents.
-Finally, the third type contained a moving phenomena.
-In each scenario, the starting position as well as their direction of movement is randomly sampled from a uniform distribution.
+Their location is kept constant in all 64 experiments. 
+As the training only contained a single phenomenon, 
+this setup represents a challenge for the agents.
+Finally, the third type contained moving phenomena. 
+In each scenario, the starting position as well as the direction of movement is randomly sampled from a uniform distribution. 
 The movement is in a straight line with a constant speed of 5m/s within an unbounded environment. 
@@
 \subsection{Baselines}
-We compare our \ac{FIRL} approach against a baseline approaches where the \ac{DQN} utilises a \ac{MLP} as well as an approach only relying on \acp{GNN}, without additional field information. 
+We compare our \ac{FIRL} approach against baseline approaches where the \ac{DQN} utilises a \ac{MLP} as well as an approach only relying on \acp{GNN}, without additional field information. 
 %
@@
 Finally, the \emph{Moving} experiment emphasizes how the informed version generates a more robust policy for new scenarios. 
- Indeed, we observe that the our approach using \ac{FIRL} maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
+ Indeed, we observe that our approach using \ac{FIRL} maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
 %
@@
     \includegraphics[width=\linewidth]{imgs/coverage-two-test}
-    \caption{\emph{Two Zones} experiment: ratio of covered to uncovered area both phenomena}
+    \caption{\emph{Two Zones} experiment: ratio of covered to the uncovered area both phenomena}
     \label{fig:coverage-two-test}

### 2023-05-19 paper.tex
re-write figure 1 caption

@@
 	\end{subfigure}
-	\caption{Qualitative test results of the proposed approach. 
-		The agents go towards the phenomenon and try to cover it without collapsing into a single point.
-		the grey area represents the area covered by the agents (last figure on the right).}
+	\caption{Agents (coloured dots) are deployed in an area and have to coordinate to cover the phenomenon.
+The phenomenon can have varying areas of importance.
+Over time, the phenomenon will be covered sufficiently without any central controller.
+%		
+%		Qualitative test results of the proposed approach. 
+%		The agents go towards the phenomenon and try to cover it without collapsing into a single point.
+%		the grey area represents the area covered by the agents (last figure on the right).
+		}
 	\label{fig:simulations}

### 2023-05-19 paper.tex
to page 2

@@
 %
-Our goal is to find a \emph{homogeneously} distributed controller $\pi$ (i.e., the same controller for the whole system) that, 
- starting from \emph{only} local configurations, leads 
- the system to achieve a certain \emph{collective} requirements through \emph{cooperation}, 
+Our goal is to find a \emph{homogeneous} distributed controller $\pi$, namely the same controller for all agents of the system:
+ starting from \emph{only} local configurations, it leads 
+ the system to achieve a certain \emph{collective} requirement through \emph{cooperation}, 
  such as spatial area coverage, phenomena tracking, and robot aggregation~\cite{DBLP:journals/firai/SchranzUSE20}.
@@
  (or simply \emph{fields}), 
- which are \emph{distributed} data structures that associate locations with values and evolve over time. 
+ which are \emph{distributed} data structures that associate each location with a computational value that evolves over time. 
 %
@@
  enabling emergent collective behaviour through the interplay of the \emph{system model} (i.e., structure and dynamics of the systems)
- and the \emph{programming model} (i.e., how collective behaviours are expressed) known as field calculus~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}.
-%
-The system model is structurally similar to \emph{swarm-like systems}, 
+ and the \emph{programming model} (i.e., how collective behaviours are expressed).
+
+On the one hand, the system model is structurally similar to \emph{swarm-like systems}, 
  as it is composed of a set of \emph{agents} 
@@
 
-The programming model is ruled by the \emph{field calculus}~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19}, 
+On the other hand, the programming model is ruled by the \emph{field calculus}~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19},  
  a core language that allows the expression of collective behaviours through the composition of \emph{functions} operating on fields.
 %
-Field calculus includes main operators for expressing \emph{spatio-temporal} compotations that allow for 
+Field calculus includes main operators for expressing \emph{spatio-temporal} computations that allow for 
 \emph{i)} the progression of values over time, achieved by transforming a field computed in a previous round into a new field; 

### 2023-05-19 paper.tex
bib ok and sec 2

@@
 
-On the other hand, the programming model is ruled by the \emph{field calculus}~\cite{DBLP:conf/esocc/ViroliDB13,DBLP:journals/corr/ViroliADPB16,DBLP:journals/tocl/AudritoVDPB19},  
+On the other hand, the programming model is ruled by the \emph{field calculus}~\cite{DBLP:conf/saso/ViroliBDP15,DBLP:journals/tocl/AudritoVDPB19},  
  a core language that allows the expression of collective behaviours through the composition of \emph{functions} operating on fields.
@@
 %
-In ScaFi~\cite{casadei2022scafi} %(a scala implementation of aggregate computing)
- it is simply expressed as:
+In the ScaFi~\cite{casadei2022scafi} implementation of field calculus is defined by function:
 \begin{lstlisting}
@@
 Along the gradient, it is possible to cast information within the system, accumulating value during the field expansion. 
-This pattern is called \emph{gradient-cast} (or \emph{G}) and is a fundamental building block for the coordination of swarm-like systems.
-It is expressed as: 
+This pattern is called \emph{gradient-cast} (or \emph{G}):
 \begin{lstlisting}
@@
  \texttt{acc} is the accumulation function that is applied to the value during the cast.
+%
+This is a fundamental building block for the coordination of swarm-like systems: in fact, the resulting system can be used to broadcast information, or route towards selected nodes.
 
@@
 %
-Let $G = (V, E)$ a graph  
-where $E$ defines the neighbourhood relations for each participating node, 
+Let $G = (V, E)$ be a graph  
+where $E\subseteq V\times V$ defines the neighbourhood relations for each participating node, 
 and $V$ identifies the nodes present in the graph. %at that time.  \lukas{no time in the definition}
@@
 The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. 
- With each traversal of an edge $e \in E$ from $v$, the number of layers increased
+ With each traversal of an edge $e \in E$ from $v$, the number of layers increases.
 Formally, a \ac{GNN} can be defined by two phases:
@@
 \begin{equation}
-GNN(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
+\mathit{GNN}(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
 \end{equation} 
@@
  thanks to its successful application in various scenarios, 
- ranging from video games (such as Alpha Go~\cite{Silver2016Go} and Atari~\cite{Atari2016DQN}) 
+ ranging from video games (such as Alpha Go~\cite{Silver2016Go} and Atari~\cite{mnih2015playing}) 
  to chatbots (like ChatGPT~\cite{ChatGPT2023}). 
@@
  the implementation of fully decentralized learning is often unfeasible due to the large number of learning agents, 
- which renders the system non-stationary and difficult to manage. 
+ which makes the system non-stationary and difficult to manage. 
 %
@@
 Given the homogeneity, large system scale, and the \emph{locality} (i,e., each agent can only observe its neighbours), 
- the problem can be modelled through the SwarMDP model~\cite{DBLP:conf/atal/SosicKZK17} -- 
- an extension of the \ac{DecPOMDP}~\cite{Decpomdp2000} model for swarm-like systems.
+ the problem can be modelled through the SwarMDP model~\cite{DBLP:conf/atal/SosicKZK17}---an extension of the \ac{DecPOMDP}~\cite{Decpomdp2000} model for swarm-like systems.
 % 
@@
   \item $\mathcal{R}: \mathcal{S} \rightarrow \mathbb{R}$ is the reward function, which is influenced by the environment;
-  \item $\pi: \mathcal{O} \rightarrow \mathcal{A}$ is the policy function, which maps the observations to the actions. It could be deterministic or stochastic.
+  \item $\pi: \mathcal{O} \rightarrow \mathcal{A}$ is the policy function, which maps the observations to the actions: it could be deterministic or stochastic.
 \end{itemize}
@@
 In swarMDP, the neighbourhood is not directly defined, but it is implicitly defined by the observation model $\xi$.
-In our specific case, the agents can only interact with 1-hop neighbourhoods and are not directly influenced by other agent observations. We can therefore restrict the observation model as follows:
+In our specific case, the agents can only interact with 1-hop neighbours and are not directly influenced by other agent observations. We can therefore restrict the observation model as follows:
 \begin{equation*}
@@
 where $\mathcal{N}^v$ is the set of neighbours of $v$.
-
+%
 This model can be used then to express the evolution of the system in time.
@@
 %Our work lies between these two approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
-As mentioned above, our work falls between automatic and manual approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
+As mentioned above, our work falls in between automatic and manual approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
 %

### 2023-05-19 paper.tex
to sec 4

@@
 To test the effectiveness of the proposed approach, 
- we decided to experiment with a case study related to swarm robotics, 
- specifically the tracking and coverage of a spatio-temporal phenomenon, cf. tracking a wildfire 
+ we experiment with a case study related to swarm robotics, 
+ specifically, tracking and coverage of a spatio-temporal phenomenon---cf. tracking a wildfire 
  or monitoring the water levels in a canal with multiple autonomous agents embodied in embedded devices (e.g., drones or IoT devices).
-The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity,...).
+The individual agents do not have any knowledge of the initial phenomenon itself (i.e., shape, size, location, velocity, and so on).
  %
@@
 \begin{enumerate}
-\item Maximise the number of agents within the phenomena
-\item Minimise the number of agents without a neighbourhood
-\item Maximise the coverage of the system
+\item maximise the number of agents within the phenomena;
+\item minimise the number of agents without neighbours;
+\item maximise the coverage of the system.
 \end{enumerate}
@@
  that provides an estimate of the current action taken by a given agent.
-
+% 
 % Regarding the first point, the reward function is simple and is defined as:
@@
  aiming to achieve good coverage by continuously tracking the phenomenon. 
-The neural network structure used consists of a layer of SuperGAT~\cite{DBLP:journals/corr/abs-2204-04879} -- A \ac{GNN} based on attention mechanisms -- and a layer of MLP. 
+The neural network structure used consists of a layer of SuperGAT~\cite{DBLP:journals/corr/abs-2204-04879} -- a \ac{GNN} based on attention mechanisms -- and a layer of MLP. 
 The hidden size was set to 256.
@@
 As the reward function is defined as a regret, 
- we decided to use the Huber loss function with $\delta = 1$, %~\cite{??}, 
+ we decided to use the Huber loss function with $\delta = 1$. %~\cite{??}, 
  %which is a combination of the $L_1$ and $L_2$ loss functions.

### 2023-05-20 paper.tex
update title (and update intro accordingly)

@@
 %
-\title{Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks} %% Todo improve
+\title{Field-informed Reinforcement Learning Collective Tasks with Graph Neural Networks} %% Todo improve
 % or: Field-Informed Reinforcement Learning: A Scalable and Effective Method for Collective Intelligence
@@
 
-A key question in designing swarm-like systems is how to create distributed controllers for the agents that enable them to perform complex collective tasks. 
+A key question in designing swarm-like systems is how to create distributed controllers for the agents that enable them to perform complex \emph{collective} tasks. 
  Two main approaches have been proposed for this problem: manual design and automatic design. 
@@
 This way, \ac{FIRL} can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
+Moreover, the learned policies have the potential to scale to different network topologies due to the inherent nature of GNNs and aggregate computations.
 %% Perhaps to much detail 

### 2023-05-20 paper.tex
tweaks

@@
 %
-\title{Field-informed Reinforcement Learning Collective Tasks with Graph Neural Networks} %% Todo improve
+\title{Field-informed Reinforcement Learning of Collective Tasks with Graph Neural Networks} %% Todo improve
 % or: Field-Informed Reinforcement Learning: A Scalable and Effective Method for Collective Intelligence
@@
 This way, \ac{FIRL} can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
-Moreover, the learned policies have the potential to scale to different network topologies due to the inherent nature of GNNs and aggregate computations.
+Moreover, the learned policies have the potential to scale with size and adapt to different network topologies due to the inherent nature of GNNs and aggregate computations.
 %% Perhaps to much detail 
@@
  our approach builds on the concept of co-fields~\cite{DBLP:journals/pervasive/MameiZL04}, 
- where agents exploit the field as a stigmergic medium to receive system-wide information and apply reasoning to this data.
+ where agents construct and exploit the field as a ``digital sign'' \cite{DBLP:journals/ijaci/CastelfranchiPT10} to receive system-wide information and apply reasoning to this data.
 %  

### 2023-05-20 paper.tex
3A title
chore: discuss about computational field and gnn

@@
   Each node is then decorated with the local observation perceived at the time $t$: $o^v_t \in O$. 
-  Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022} if we set $f_v = o^v_t$.
+  This graph can be used both to compute computational fields and as input for a GNN, as done in previous work~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022}.
+  % if we set $f_v = o^v_t$.
 \subsection{Motivation}
@@
 Finally, we will detail the learning algorithm designed and used to synthesise the policy.
-\subsection{Architecture and System dynamics}
+\subsection{Architecture, fields and aggregate dynamics}
 %\ga{Plan: 

### 2023-08-01 paper.tex
content: handle simple points raised by the reviewers

@@
 \newcommand{\mv}[1]{\meta{green}{MV}{#1}}
+\newcommand{\review}[1]{\textcolor{red}{#1}}
 %% ACRONYMS
@@
  R^{\mathcal{N}}_{v} = \begin{cases}
-  1 & \text{if } \mathcal{N} > 0 \\
+  1 & \text{if } |\mathcal{N}| > 0 \\
   0 & \text{otherwise} 
@@
 The results of the training process are summarized in \Cref{fig:training}.
+% 
+\review{In the charts, the line represents the average value of the metric, while the shaded area represents the standard deviation.}
 %
@@
     \includegraphics[width=\linewidth]{imgs/coverage-two-test}
-    \caption{\emph{Two Zones} experiment: ratio of covered to the uncovered area both phenomena}
+    \caption{\emph{Two Zones} experiment: ratio of covered to the uncovered zones both phenomena}
     \label{fig:coverage-two-test}

### 2023-08-01 paper.tex
content: on review

@@
   the circle area on represents the phenomenon to be monitored.
-  \Cref{fig:static} represents the scenario used during training. 
-  The others are used in the test phase to evaluate the policy found. }
+  \Cref{fig:static} represents the scenario used during training \review{as well as during test}. 
+  The others \review{instead} are only used in the test phase to evaluate the policy found. }
   \label{fig:scenarios}

### 2023-08-03 paper.tex
chore: prepare for camera ready

@@
 \newcommand{\mv}[1]{\meta{green}{MV}{#1}}
-\newcommand{\review}[1]{\textcolor{red}{#1}}
+\newcommand{\review}[1]{{#1}}
 %% ACRONYMS
@@
 %
-The main contribution of \ac{FIRL} is that it allows us to create distributed controllers that are informed by a collective knowledge that has been distilled during training, but that use only local information when deployed.
+The main contribution of \ac{FIRL} is that it allows us to \emph{create distributed controllers that are informed by a collective knowledge that has been distilled during training, but that use only local information when deployed}.
 This way, \ac{FIRL} can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
@@
 In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $\mathcal{N}_G(v)$,
- and then combining it with the node's current embedding $h_v$.
-%
-The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. 
- With each traversal of an edge $e \in E$ from $v$, the number of layers increases.
-Formally, a \ac{GNN} can be defined by two phases:
+ and then combining it with the node's current embedding $h_v$ in a process called \emph{message passing}~\cite{DBLP:conf/icml/GilmerSRVD17}.
+%
+The \acp{GNN} is partitioned into several message passing layer $k$
+ where each of them is responsible for computing the node embedding $h_v^{(k)}$ for each node $v \in V$.
+ %With each traversal of an edge $e \in E$ from $v$, the number of layers increases.
+Formally, a \ac{GNN} can be defined by three phases:
 \begin{equation}
-a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}_G(v)\right\}\right)   
+m^{(k)}_{uv} = \psi^{(k)}\left(h_u^{(k-1)}, h_v^{(k-1)}, e^{(k-1)}_{uv}\right) 
 \end{equation}
@@
 \begin{equation}
-h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
+a^{(k)}_u = \bigoplus^{(k)}\left(\left\{m^{(k)}_{uv}: v \in \mathcal{N}_G(u)\right\}\right)
+\end{equation}
+
+\begin{equation}
+h_u^{(k)} = \phi^{(k)}\left(h_u^{(k-1)}, a^{(k)}_u\right)
 \end{equation}
@@
 %
-$h_v^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
-%
-The differential part comes into play in the $\operatorname{COMBINE}$ function, 
+$h_v^0$ is the initial embedding of node $v$, and it is usually set to the node's feature vector $f_v$.
+%
+The differential part comes into play in the $\psi$ and $\phi$ functions, 
  which is usually a differentiable function such as a neural network layer.
 %
-$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
- and it could be a simple sum, max or sum of products. 
+The $\psi$ function is called \emph{message function}, and it is responsible for computing the message $m_{uv}^{(k)}$ from node $u$ to node $v$.
+The $\phi$ function is called \emph{update function}, and it is responsible for updating the node embedding $h_v^{(k)}$ of node $v$.
+$bigoplus $ instead is a function that aggregates the information from the neighbours of a node $v$
+ and it could be a simple sum, max or sum of products and it should be permutation invariant. 
  More complex aggregation functions are available~\cite{pellegrini2020learning}.
@@
 % Regarding the first point, the reward function is simple and is defined as:
-Formally, we define the reward of an agent being within the phenomenon as
- \begin{equation*}
- R^a_{v} = \begin{cases}
-  1 & \text{if } \zeta_v > 0 \\
-  0 & \text{otherwise.} 
- \end{cases}
- \end{equation*}
+Formally, we define the reward of an agent being within the phenomenon as:
+$
+  R^a_{v} = 1 \text{ if } \zeta_v > 0 \text{ else } 0
+$.
 Namely, an agent is considered within the phenomena as soon as the drone can sense the phenomenon.
@@
  In this case, the reward is defined as:
- \begin{equation*}
- R^{\mathcal{N}}_{v} = \begin{cases}
-  1 & \text{if } |\mathcal{N}| > 0 \\
-  0 & \text{otherwise} 
- \end{cases}
- \end{equation*}
+$
+R^{\mathcal{N}}_{v} = 1 \text{ if } |\mathcal{N}| > 0 \text{ else } 0
+$.
 
@@
 % 
-\begin{equation*}
+$
 R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
-\end{equation*}
-where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
+$.
+Where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
 %
@@
 The final reward function $R_{v}$ for an agent $v$ is defined as:
-\begin{equation*}
+$
 R_{v} = (\frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}) - 1
-\end{equation*}
+$.
 Specifically, we decided to express the signal as a \emph{regret}
@@
 \end{equation*}
-where e is the current episode number. 
+Where e is the current episode number. 
 This leads to a high number of random actions at the beginning and gradually shifts towards exploitation in the later episodes. 
@@
 %
-Formally, first, we define the overall coverage for a certain time step as:
+Formally, we define the overall coverage for a certain time step as:
 \begin{equation*}
@@
 % 
-\review{In the charts, the line represents the average value of the metric, while the shaded area represents the standard deviation.}
+\review{In the charts, the line represents the average value of a metric of interest, while the shaded area represents its standard deviation.}
 %
@@
   \end{subfigure}
-  \caption{Training results of field-informed \ac{RL}.
-  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
+  \caption{Training results of FIRL.
+  It can cover the phenomenon better than the baselines and it reaches a higher reward. 
   }
@@
     \includegraphics[width=\linewidth]{imgs/coverage-test.pdf}
-    \caption{Ratio of coverage of the phenomena in the three types of experiments. Our \ac{FIRL} approach can outperform other approaches lacking field information.}
+    \caption{Ratio of coverage of the phenomena. Our \ac{FIRL} approach can outperform other approaches lacking field information.}
     \label{fig:coverage-test}
@@
   \end{subfigure}
-	\caption{Quantative test results in the three types of scenarios.
-	 	We can see that the proposed approach is able to cover and track the phenomenon better than the baselines.
+	\caption{Quantitative test results.
+	 	The proposed approach can cover and track the phenomenon better than the baselines.
 	}
@@
 \end{figure*}
-
 \begin{figure*}

### 2023-08-04 paper.tex
chore: finalise for camera ready

@@
 %
-The main contribution of \ac{FIRL} is that it allows us to \emph{create distributed controllers that are informed by a collective knowledge that has been distilled during training, but that use only local information when deployed}.
+The main contribution of \ac{FIRL} is the \emph{definition of distributed controllers that are informed by collective knowledge that has been distilled during training, but that use only local information when deployed}.
 This way, \ac{FIRL} can achieve a balance between manual design and automatic design, combining the benefits of both approaches while mitigating their drawbacks.
@@
 The differential part comes into play in the $\psi$ and $\phi$ functions, 
- which is usually a differentiable function such as a neural network layer.
+ which is usually a differentiable function such as a neural network.
 %
@@
 The $\phi$ function is called \emph{update function}, and it is responsible for updating the node embedding $h_v^{(k)}$ of node $v$.
-$bigoplus $ instead is a function that aggregates the information from the neighbours of a node $v$
+$\bigoplus $ instead is a function that aggregates the information from the neighbours of a node $v$
  and it could be a simple sum, max or sum of products and it should be permutation invariant. 
@@
  performs \emph{actions} in an \emph{environment} (i.e., everything outside the agent) according to a \emph{policy}, 
- to maximise long-term \emph{reward signal} (or minimise a \emph{regret}).
+ to maximise long-term \emph{reward signal}. %(or minimise a \emph{regret}).
 
@@
 %
-In particular, in this work, we consider homogenous \ac{MAARL}~\cite{yang2021many}, 
- where the set of agents is large ($N \gg 2$) and each agent is interchangeable and indistinguishable.
+In particular, in this work, we consider \emph{homogenous \ac{MAARL}}~\cite{yang2021many}, 
+ where the set of agents is large ($N \gg 2$) and each agent is \emph{interchangeable} and \emph{indistinguishable}.
 %
@@
  the entire system may not be a viable solution due to scalability concerns. 
- To address this challenge, a practical solution is the adoption of \ac{CTDE}.
-%
-The idea is to learn a policy at simulation time where there is a collective view of the system, 
- and then at runtime use the policy found with global information. 
-This approach allows policies that are influenced by global information but only require local information to function at runtime. 
+ To address this challenge, a practical solution is the adoption of \emph{\ac{CTDE}} approach.
+%
+The idea is to learn a policy at simulation time when there is a collective view of the system, 
+ and then at runtime use that policy but only with local observations. 
+%This approach creates policies that are influenced by global information but only require local information to function at runtime. 
 %
@@
 \end{equation*}
-Given a time $t$, the system can be also represented as a graph $G_t = (V_t, E_t)$, 
-  where $E_t$ is built from $\mathcal{N}$.
+Given a time $t$, the system can be also represented as a graph $G^t = {V^t, E^t}$, 
+  where $E^t$ is built from $\mathcal{N}$.
   Each node is then decorated with the local observation perceived at the time $t$: $o^v_t \in O$. 
-  This graph can be used both to compute computational fields and as input for a GNN, as done in previous work~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022}.
+%
+This graph can be used both to compute computational fields and, as done in previous work~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022}, can be the input for a GNN.
   % if we set $f_v = o^v_t$.
@@
 %  
-We present a subsequent approach where agent intelligence is synthesised through \ac{MAARL} and the use of \ac{GNN} to learn a local representation from the neighbourhood.
+We present a subsequent approach where agent intelligence is synthesised through \ac{MAARL} and \ac{GNN} is used to learn a local representation from the neighbourhood.
 %
@@
 %
-Let $\Gamma$ be the aggregate program that takes a graph $G_t$ decorated by $o^v_t$, 
+Let $\Gamma$ be the aggregate program that takes a graph $G^t$ decorated by $o^v_t$, 
  representing the participating agents and their neighbourhood relations at time $t$, as input. 
@@
 %
-The evaluation of $\Gamma$ produces a field value $\theta^v_t$ for each node $v$ in $G_t$. 
-From this field, we construct the feature vector $f_v$ for each node $v$ in $G_t$ as follows:
+The evaluation of $\Gamma$ produces a field value $\theta^v_t$ for each node $v$ in $G^t$. 
+From this field, we construct the feature vector $f_v$ for each node $v$ in $G^t$ as follows:
  $f_v = (\theta^v_t, o^v_t)$.
@@
  might appear as global knowledge, 
- this is not the case as the information is never aggregated globally. 
+ this is not the case as the information is \emph{never} aggregated globally. 
 The individual agents only combine information from their local neighbourhood.
@@
 %The specific program in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
-The program $\Gamma$ in question is a simple application of block G, where the source is the maximum value of the neighbourhood. 
-This can be expressed in Scafi as follows:
+The program $\Gamma$ in question is a simple application of block \lstinline|G|, where the source is the maximum value of the neighbourhood. 
+This can be expressed in ScaFi as follows:
 \begin{lstlisting}[mathescape=true]
@@
 where \texttt{maxHood} is a function that returns the maximum value of the neighbourhood, 
- \texttt{nbr} is a function that returns the value of the neighbourhood, 
+ \texttt{nbr} is a function that a neighbourhood field of values (in this case, the sensor value $\zeta$),
 \texttt{sense} is a function that returns the value of the sensor, and
@@
 Formally, we define the reward of an agent being within the phenomenon as:
-$
+\begin{equation}
 R^a_{v} = 1 \text{ if } \zeta_v > 0 \text{ else } 0  
-$.
+\end{equation}
 Namely, an agent is considered within the phenomena as soon as the drone can sense the phenomenon.
@@
  In this case, the reward is defined as:
-$
+\begin{equation*}
 R^{\mathcal{N}}_{v} = 1 \text{ if } |\mathcal{N}| > 0 \text{ else } 0
-$.
+\end{equation*}
 
@@
 % 
-$
+\begin{equation*}
 R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
-$.
+\end{equation*}
 Where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
@@
 The final reward function $R_{v}$ for an agent $v$ is defined as:
-$
+\begin{equation*}
  R_{v} = (\frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}) - 1  
-$.
+\end{equation*}
 Specifically, we decided to express the signal as a \emph{regret}
@@
 In this case, we chose to use an exponential epsilon decay, defined as:
-\begin{equation*}
+$
 \epsilon = \epsilon_{min} + (\epsilon_{max} - \epsilon_{min}) \cdot e^{-\lambda \cdot e}
-\end{equation*}
+$.
 Where e is the current episode number. 
@@
 %
-These baselines are used to verify the effectiveness of the components used in the field-informed \ac{RL}. 
+These baselines are used to verify the effectiveness of the components used in the FIRL. 
 %

### 2023-08-04 paper.tex
chore: minors
shrink abstract for 200 words

@@
 impacted by the challenges posed by the very notion of distributed intelligence.
-These problems arise from agents acquiring information locally, sharing their knowledge, and acting accordingly in their environment in order to achieve a common, global goal.
-%
-Individual agents need to control how they share their knowledge in order to not overburden the network or other agents with unnecessary information.
-These problems are even more evident in large-scale collective adaptive systems, where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder.
+These problems arise from agents acquiring information locally, sharing their knowledge, and acting accordingly in their environment to achieve a common, global goal.
+%
+%Individual agents need to control how they share their knowledge to not overburden the network or other agents with unnecessary information.
+These issues are even more evident in large-scale collective adaptive systems, 
+ where agent interactions are necessarily proximity-based, thus making the emergence of controlled global collective behaviour harder.
 
@@
 We utilise manually designed \emph{computational fields} (globally distributed data structures) to manage global agent coordination.
+%
 Then, using \emph{Deep Q-learning} in combination with \emph{Graph Neural Networks} we enable the agents to learn the necessary local behaviour automatically to solve collective tasks, relying on those fields through local perception.
 %
-This allows us to create distributed controllers informed by a collective knowledge 
- that has been distilled during learning, but that uses only local information at runtime.
+%This allows us to create distributed controllers informed by a collective knowledge 
+% that has been distilled during learning, but that uses only local information at runtime.
 % 
@@
   \hfill
-  \begin{subfigure}[b]{0.87 \linewidth}
+  \begin{subfigure}[b]{0.885\linewidth}
     \includegraphics[width=\linewidth]{imgs/coverage-in-time}
@@
   \hfill
-  \begin{subfigure}[b]{0.9\linewidth}
+  \begin{subfigure}[b]{0.91\linewidth}
     \includegraphics[width=\linewidth]{imgs/reward-in-time}
